### 2021-02-25 document/slides/slides.tex
Slides first draft

\documentclass{beamer}
\usepackage[utf8]{inputenc}

%\usepackage{utopia} %font utopia imported

\usepackage{amsthm, booktabs, setspace, natbib, amsfonts,amssymb,epstopdf,textcomp,listings, xcolor}

\usepackage{adjustbox} % Shrink stuff
\usepackage{pgfpages}

\usepackage{tikz}
\usepackage{verbatim}
\setbeamertemplate{note page}{\pagecolor{yellow!5}\insertnote}
\usetikzlibrary{positioning}
\usetikzlibrary{snakes}
\usetikzlibrary{calc}
\usetikzlibrary{arrows}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{shapes.misc}
\usetikzlibrary{matrix,shapes,arrows,fit,tikzmark}

\usepackage{amsmath}
\usepackage{mathpazo}
\usepackage{hyperref}
\usepackage{lipsum}
\usepackage{multimedia}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{dcolumn}
\usepackage{bbm}
\usepackage[space]{grffile}
\usepackage{tabularx}
\usepackage{subfig}
\usepackage{float}
\usepackage{multirow}

\usepackage{tabularx}
\usepackage{fancyhdr}
\usepackage{lscape}
\usepackage{mdframed}
\usepackage{mathtools}
\usepackage{array}

\usepackage{mwe}
\usepackage[skip=2pt,font=scriptsize]{caption}
\captionsetup{labelfont=bf, format=hang, labelsep=period, labelformat = empty}
%labelformat  =simple
%labelformat = empty

\usepackage{appendixnumberbeamer} %for avoid counting appendix slides add \appendix
\usepackage{colortbl}
\usepackage{xcolor}

%\usetheme{Warsaw}
\usetheme{Madrid}
\usecolortheme{default}

%------------------------------------------------------------
%This block of code defines the information to appear in the
%Title page
\title[Improved Seed Adoption] %optional/
{Improved Seed Adoption and Climate Change: Hybrid Maize in Ethiopia}

\author[et al] % (optional)
{Barriga-Cabanillas,  Chiarella, Correa, Michuda}

% {Oscar Barriga-Cabanillas\inst{1} \\
% (with Travis Lybbert\inst{2})}

\institute[UC Davis] % (optional)
{
%  \inst{1}%
  Agricultural and Resource Economics\\
  UC Davis
 }

\date[Feb 2021] % (optional)
{CIGAT \\
February 2021}

%End of title page configuration block
%------------------------------------------------------------

%------------------------------------------------------------
%The next block of commands puts the table of contents at the 
%beginning of each section and highlights the current section:

\AtBeginSection[]
{
  \begin{frame}
    \frametitle{Table of Contents}
    \tableofcontents[currentsection]
  \end{frame}
}
%------------------------------------------------------------

\begin{document}

\setbeamertemplate{enumerate items}[square]
\setbeamercolor{item projected}{bg=none,fg=black}

%\setbeamercolor{enumerate items}{〈key=value〉 list} to
%\setbeamercolor{item projected}{bg=none,fg=beamer@blendedblue}

\setbeamertemplate{itemize items}[circle]
\setbeamercolor{itemize item}{fg=black}
\setbeamercolor{itemize subitem}{fg=black}

%%% TIKZ STUFF
\tikzset{   
        every picture/.style={remember picture,baseline},
        every node/.style={anchor=base,align=center,outer sep=1.5pt},
        every path/.style={thick},
        }
\newcommand\marktopleft[1]{%
    \tikz[overlay,remember picture] 
        \node (marker-#1-a) at (-.3em,.3em) {};%
}
\newcommand\markbottomright[2]{%
    \tikz[overlay,remember picture] 
        \node (marker-#1-b) at (0em,0em) {};%
}
\tikzstyle{every picture}+=[remember picture] 
\tikzstyle{mybox} =[draw=black, very thick, rectangle, inner sep=10pt, inner ysep=20pt]
\tikzstyle{fancytitle} =[draw=black,fill=red, text=white]
%%%% END TIKZ STUFF

%The next statement creates the title page.
\frame{\titlepage}

% %---------------------------------------------------------
% %This block of code is for the table of contents after
% %the title page
% \begin{frame}
% \frametitle{Table of Contents}
% \tableofcontents
% \end{frame}
% %---------------------------------------------------------

%---------------------------------------------------------
\begin{frame}
\frametitle{Research question}

Food security is a major challenge: Ethiopia is no exception

\begin{itemize}
    \item Struggled to provide an adequate and reliable food supply
    \item Breakthroughs in maize germplasm ($\uparrow$ yields, $\uparrow$ drought tolerance)
\end{itemize}

However adoption of improved seed has remained a challenge

\begin{enumerate}
    \item Market integration
    \item Access to supplementary inputs
    \item Agro-ecological considerations
\end{enumerate}

\textbf{We study how these heterogeneity mask the true impact of adoption}

\begin{enumerate}
    \item Adoption where  net benefits are the highest
    \item How adaptation strategies augment comparative advantage in adoption
    \item Drivers of dis-adoption of improved maize seed
\end{enumerate}

\end{frame}

%---------------------------------------------------------
\begin{frame}
\frametitle{Empirical Strategy I}

Correlated Random Coefficient Model and comparative advantage:

\begin{itemize}
    \item Farmers face unobservable heterogeneous returns to adoption (comparative advantage)
\end{itemize}

\begin{center}
    y_{it}= \delta + \beta h_{it} + \theta_{i} + \phi\theta h_{it} + \tau_{i} + \epsilon_{it}

\end{center}

\begin{itemize}
    \item $y_{it}$: yields or profits for household i at time t
    \item $h$:  indicator for adoption
    \item $\tau$:  household’s absolute advantage
    \item $\theta$: household’s comparative advantage
    \item $\phi$: Gains from adopting, relative to a household’s comparative advantage. 
\end{itemize}

\end{frame}

%---------------------------------------------------------
\begin{frame}
\frametitle{Empirical Strategy II}

Advantages:
\begin{enumerate}
    \item Identification does not rely on the existence of a valid instrumental variable (linear projection of observed adoption history)
    \item Disentangles a household’s absolute advantage from comparative advantage in adopting. 
\end{enumerate}

Extending the model:

\begin{itemize}
    \item Implementing a nonparametric panel identification
    \item Modeling time-varying characteristics to the Group Random Coefficients (GRC) model
\end{itemize}  
\end{frame}

%---------------------------------------------------------

\end{document}

### 2021-11-24 document/slides/slides.tex
Updates from Overleaf

@@
 
-\author[et al] % (optional)
-{Barriga-Cabanillas,  Chiarella, Correa, Michuda}
+\author[et al]{Barriga-Cabanillas,  Chiarella, Correa, Michuda}
 
@@
 \date[Feb 2021] % (optional)
-{CIGAT \\
+{CGIAR-SPIA \\
 February 2021}
@@
 
-\textbf{We study how these heterogeneity mask the true impact of adoption}
+\textbf{We study how heterogeneity masks the true impact of adoption}
 
@@
 
-Correlated Random Coefficient Model and comparative advantage:
+Correlated Random Coefficient Model and comparative advantage (Suri 2011; Tjernstr\"{o}m et al. 2020; Michler et al. 2019; Barriga et al. 2018)
 
@@
 \begin{center}
+$$
     y_{it}= \delta + \beta h_{it} + \theta_{i} + \phi\theta h_{it} + \tau_{i} + \epsilon_{it}
-
+$$  
 \end{center}
@@
 \begin{frame}
-\frametitle{Empirical Strategy II}
+\frametitle{Empirical Strategy II and Extensions}
 
@@
 \begin{itemize}
-    \item Implementing a nonparametric panel identification
-    \item Modeling time-varying characteristics to the Group Random Coefficients (GRC) model
+    \item Implementing a nonparametric panel identification in a novel GRC (Group Random Coefficients)
+    \item Modeling and incorporating time-varying characteristics 
+    \item Complement ESS data with precipitation (CHIRPS) and temperature (CPC) data
 \end{itemize}  

### 2021-11-24 paper.tex
Updates from Overleaf

\documentclass{article}

\title{Estimating Comparative Advantage in Improved Seed Adoption: Hybrid Maize in Ethiopia}

\usepackage{natbib}

\begin{document}

\maketitle

\section{Introduction}

\section{Context}

\section{Data}

\section{Estimation Strategy}

\section{Results}

\section{Conclusion}

\end{document}

### 2021-11-24 paper.tex
Updates from Overleaf

@@
 \usepackage{natbib}
+\bibliographystyle{rusnat}
+
+\usepackage[margin=1in]{geometry}
+\usepackage{setspace}
+\doublespacing
 
@@
 
+Food security is a major challenge in many East African countries-- Ethiopia is no exception, having historically struggled to provide an adequate and reliable food supply \citep{Ramakrishna2002-hv, Jaleta2018-oj}.
+Since the drought of 1984, Ethiopia has become increasingly reliant on maize.
+Despite numerous technological breakthroughs in maize germplasm with increased yields and drought tolerance, adoption of improved seed has remained a challenge.
+There is a broad literature on why sub-Saharan farmers continue to use traditional farming techniques when more modern, higher-return agricultural technologies are available.
+Several answers have been proposed to explain this puzzle.
+These include imperfections in credit markets \citep{Croppenstedt2003-pq}, property rights \citep{Place2000-el}, social learning \citep{Conley2010-ue,Foster1995-bz,Munshi2004-og}, lack of commitment devices \cite{Duflo2009-iv}, and high transportation costs that increase the cost of agricultural inputs \citep{Byerlee2013-qk}.
+This issue is becoming more important as even current varieties of improved seed are over 20 years old and their effectiveness in terms of yield improvement and disease resistance is waning \citep{Abate2015-rj}.
+
+Adoption of a technology involves several dimensions such as market integration, access to supplementary inputs like fertilizer, as well agro-ecological considerations such as climate, altitude and soil quality.
+Ignoring the heterogeneity of these factors can hide the true impact of adoption, as different sub-populations may be impacted by the same technology in different, and even opposite ways.
+One way to tackle this problem, proposed by \citep{Suri2011-oi}, centers around quantifying the comparative advantage of a household to adopting improved seed varieties.
+Even when average returns are high, farmers may face heterogeneous returns based on their own, unobservable, comparative advantage in adopting new technologies.
+Using a correlated random coefficient model, Suri’s seminal paper finds that farmers’ lack of adoption is driven by insufficient net benefits to the technology stemming from poor infrastructure.
+Similar work in Ethiopia on improved chickpea varieties suggests that the economic returns to adoption plays a major role in the decision to adopt \citep{Michler2018-wk}.
+In our paper, we build on that literature by investigating hybrid maize seed adoption in Ethiopia, focusing on how heterogeneous returns are also driven by climate and how this can affect dis-adoption decisions.
+
+We will explore how the role of heterogeneity in a household’s comparative advantage explains differences in the returns to adoption, and what drives those differences.
+Our objectives center around four main questions: (1) does adoption of improved seed take place where the net benefits of adoption is highest?
+(2) How do adaptation strategies such as water storage, and irrigation augment a household’s comparative advantage in adoption?, (3) what are the drivers of dis-adoption of improved maize seed?
+and (4) is it possible to extend the methodology in \citep{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} by including time-varying characteristics to the Group Random Coefficients (GRC) model?
+
 \section{Context}
+
+Evidence has shown that the adoption of improved seed leads to increases in agricultural yields \citep{Carter2014-fm}, food security \citep{Shiferaw2014-op} and poverty reduction \citep{Minten2008-tj}.
+Despite such positive impacts, the low adoption levels of improved seeds varieties in Sub-Saharan Africa remains a puzzle.
+Multiple explanations have been proposed for observed adoption rates, including the type of technology promotion mechanisms used by governments, NGO’s, and businesses, the characteristics of the improved seeds and their adaptation to specific agro-climatic environments \citep{Bird2020-nt}, as well as growing conditions, such as the type of soil, input use, and other farmer characteristics \citep{Munshi2004-og}.
+Specifically, extension services and on-farm field trials, seed variety characteristics and rainfall have been found to be crucial to the adoption of improved maize in Tanzania \citep{Kaliba2000-jh}.
+In Ethiopia, labor, fertilizer use, farmers’ experience with extension packages, rainfall suitability and prices have been shown to increase the opportunity cost of not adopting improved wheat varieties \citep{Wale2006-bv}.
+Market access and the accessibility of extension services are also critical in the adoption of improved chickpea varieties in Ethiopia \citep{Verkaart2019-ol}.
+Other important determinants of heterogeneous effects include risk aversion \citep{Holden2016-vy}, farm size \citep{Ghimire2015-bd}, and credit constraints \citep{Simtowe2008-jn,Balana2020-hx}.
+The consensus in this literature is that the success of improved varieties depends on a range of factors that need to be better understood.
+Success of improved seeds should be accompanied by complementary conditions and inputs that would allow the full potential of these varieties to be realized.
+A different branch of the literature provides an alternative explanation for such low adoption rates, focusing on heterogeneous potential returns to adoption \citep{Suri2011-oi}.
+This literature suggests that even if returns to adoption are high, farmers may fail to adopt if they face low comparative advantage to adoption.
+Other methods used in the literature fail to account for this heterogeneity, admitting only the presence of time-constant and/or time-varying unobserved heterogeneities (depending on the data structure) for identification purposes \citep{Kassie2018-xn,Falco2011-rt}.
+These methods ignore the possibility of a farmer’s comparative advantage of using an improved technology over a traditional one as playing a central role in their adoption decision.
+We rely on Suri’s method and its extension in \citep{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} to account for such heterogeneous comparative advantages and evaluate how these are explained by climatic conditions.
+In Ethiopia, rainfall deviations from historical averages are found to be the main driver of heterogeneous effects for adopting agronomic packages \citep{Marenya2020-kb}.
+In Malawi, exposure to past droughts is found to increase the adoption of drought-tolerant maize among farmers that also receive subsidies \citep{Katengeza2019-af} and the dis-adoption of traditional local maize \citep{Holden2016-vy}.
+We will contribute to this literature by considering historical climate averages on the suitability of improved varieties to understand whether adoption is indeed realized where its potential is highest.
+Moreover, we will contribute to the understanding of how water conservation innovations and irrigation mitigate the effects of droughts on the potential of improved seeds.
 
@@
 
+The current study uses the Ethiopia Socioeconomic Survey (ESS), a part of the Agricultural Sample Survey (AgSS), a survey 
+designed to obtain production estimates for the major crops.
+It was conducted in collaboration with the Central Statistics Agency of Ethiopia, the World Bank and CGIAR (Consortium of International Agricultural Research Centers) \citep{kosmowski2020shining}.
+The ESS is meant to be representative of the most populous regions of the country.
+The ESS contains several questionnaires targeted at getting an understanding of the state of agriculture.
+This includes a survey that is conducted post-planting and then post-harvest.
+These surveys are more important to our analysis as it gives a close look to inputs used in agriculture, the types of seeds used (in various levels of detail) and yields collected post-harvest.
+There were significant differences between the analysis carried out in ESS 1-3 and the most recent, ESS4.
+Most notably, the ESS4 is no longer a panel dataset made up of the households of ESS 1-3, but it includes important information on the DNA fingerprinting of seed type and misclassification rates.
+This is important information as the first three waves of ESS only included self-reported hybrid maize status.
+As a result, our main analysis will include the first three waves of the ESS and use the fourth wave later for robustness and validation.
+
+Figure \ref{map:regions} shows the spatial distribution of the locations of the our households of study.
+
+Table \ref{tbl:summary} shows a set of the summary statistics for the sample of house
+
 \section{Estimation Strategy}
+
+Methods
+
+Our base analysis will be inspired by the work of \citep{Suri2011-oi}.
+This estimation framework imposes a parametric relationship between farmers' absolute advantage (i.e., the component of productivity that is independent of technology) and their comparative advantage (i.e., their relative productivity of adopting the improved technology over not adopting).
+This method seeks to estimate the following equation:
+
+$$
+yit=+ hit + i+ihit + i+it
+$$
+
+where y is yields or profits for household i at time t, h is an indicator for adoption,  is a household’s absolute advantage,  is a household’s comparative advantage and  describes the gains to adopting, relative to a household’s comparative advantage.
+This framework provides an advantage over other methods \citep{Wooldridge1997-xj,Heckman1998-pt}.
+Firstly, identification does not rely on the existence of a valid instrumental variable.
+Instead, identification comes from a linear projection of an individual's returns to adoption onto the observed history of their adoption, an approach similar to the correlated random effects (CRE) method in \citep{Chamberlain1984-uk}.
+Secondly, it disentangles a household’s absolute advantage when taking part in agriculture and its comparative advantage in adopting.
+This is crucial in understanding the adoption decision as it is not dependent on a household’s agricultural ability, but on the gains from taking it up.
+We and co-authors developed this further in \citep{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} by proposing a more flexible and tractable approach using a group random coefficient strategy that draws on recent developments in the nonparametric panel identification literature.
+Our strategy has a clear economic interpretation and can elucidate potential identification concerns of the CRC model to the practitioner.
+It is also more easily extended to multiple time periods and new levels of heterogeneity than the CRC approach.
+
+Estimating equation 1 will allow us  to analyze the distribution of the gains to adoption and to create a distribution of heterogeneous returns.
+We will investigate whether comparative advantage is driven by two key policy relevant factors: water conservation strategies and deviation from historical climate averages.
+The first will elucidate if adaptation strategies in water management are associated with higher levels of gains to adoption.
+The second will seek to understand whether households are adapting to climate change through their adoption decisions.
+This will require the extension of the method developed in \citep{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc}, by allowing for time-varying characteristics in the identification of comparative advantage.
+
+Implications
+
+Our work has far-reaching policy implications.
+The adoption and impact of agricultural technologies is inherently heterogeneous.
+Our methodology incorporates that from the outset, allowing us to investigate whether synergies between seed adoption and water conservation strategies interact; a key component to understanding how investment into such projects translates to economic returns.
+Moreover, our research informs on not only the adoption but also dis-adoption decisions of technologies as well.
+In this sense, we provide valuable information to policymakers on how to better target seed marketing or infrastructure to help agricultural households improve their livelihoods
 
@@
 
+\bibliography{references}
+
 \end{document}

### 2021-11-24 paper.tex

@@
 \bibliographystyle{rusnat}
+
+\usepackage{graphics}
 
@@
 This is important information as the first three waves of ESS only included self-reported hybrid maize status.
-As a result, our main analysis will include the first three waves of the ESS and use the fourth wave later for robustness and validation.
+Since our estimation strategy requires a balanced panel dataset of maize growers, our main analysis will include the first three waves of the ESS.
+We use the fourth wave later for robustness and validation.
 
-Figure \ref{map:regions} shows the spatial distribution of the locations of the our households of study.
+Figure \ref{map:regions} shows the spatial distribution of the locations of the our households of study. The number of households surveyed did
+
+\begin{figure}
+    \centering
+    \includegraphics{results/figures/map_hhids.pdf}
+    \caption{Mean Spatial Distribution of Households across Ethiopia}
+    \label{map:regions}
+\end{figure}
 

### 2021-11-24 paper.tex
Updates from Overleaf

@@
 
-Figure \ref{map:regions} shows the spatial distribution of the locations of the our households of study. The number of households surveyed did
+Figure \ref{map:regions} shows the spatial distribution of the locations of the our households of study.
+As previously mentioned these are households that grow either improved or conventional maize and that are observed for all three rounds of the survey.
+Most households come from the northwest region of the
 
@@
     \includegraphics{results/figures/map_hhids.pdf}
-    \caption{Mean Spatial Distribution of Households across Ethiopia}
+    \caption{Spatial Distribution of Households Surveyed}
     \label{map:regions}

### 2021-11-24 paper.tex

@@
 As previously mentioned these are households that grow either improved or conventional maize and that are observed for all three rounds of the survey.
-Most households come from the northwest region of the
+Most households come from the northwest region of Amhara, Tigray and Oromiya which is consistent with where maize is mostly grown in Ethiopia \citep{Abate2015-rj}.
 
@@
 
-Table \ref{tbl:summary} shows a set of the summary statistics for the sample of house
+Table \ref{tbl:summary} shows a set of summary statistics for our sample.
+
+% Explain summary stats
+% different definitions of yields
+% hybrid and trajectories
 

### 2021-11-24 paper.tex
Updates from Overleaf

@@
 \bibliographystyle{rusnat}
+\usepackage{multicol}
+\usepackage{multirow}
+\usepackage{booktabs}
+\usepackage{caption}
 
@@
     \caption{Spatial Distribution of Households Surveyed}
+    \figcaption{Note: Green markers denote the locations of households}
     \label{map:regions}
@@
 Table \ref{tbl:summary} shows a set of summary statistics for our sample.
+The mean parcel size for a household is about 0.15 Ha, with a maximum size of a little less than 1.5 Ha.
+Harvest labor is disproportionately skewed to family participation with around 20 days of harvest labor being conducted by family members and only about 2 days for hired labor.
+This
+
+\input{results/tables/summary}
 

### 2021-11-24 paper.tex

@@
 Harvest labor is disproportionately skewed to family participation with around 20 days of harvest labor being conducted by family members and only about 2 days for hired labor.
-This
+This is indicative of either malfunctioning labor markets, income constraints or high supervision costs.
+The distances to some key areas are also highlighted in Table \ref{tbl:summary}.
+On average households are around 13 kilometers away from an asphalt road and almost 60 kilometers away from the nearest market.
+This makes both the purchase of improved seed and the marketing of yields subject to transportation costs and a difficult enterprise.
+Fertilizer costs are on average 8,350 Birr (~178 USD) and with a large amount of the sample not using fertilizer at all.
+This is particularly important in the context of improved maize seed use as it requires fertilizer to be most effective.
+Only 5\% of the sample irrigates their crops making households susceptible weather shocks.
 
 \input{results/tables/summary}
+
+The ESS has has both self-reported
 
@@
 \section{Estimation Strategy}
-
-Methods
 

### 2021-11-24 paper.tex
Updates from Overleaf

@@
 \usepackage{caption}
+\usepackage{amsmath}
 
@@
 
-The ESS has has both self-reported
-
-% Explain summary stats
-% different definitions of yields
-% hybrid and trajectories
+The ESS has has both self-reported and crop cut yields available.
+This provides two sources of yield information, although self-reported yields seem to be overly large compared to either fresh or dried crop cut yields.
+Self-reported yields are both more susceptible to behavioral biases, but also to confusion caused by differences in assumed units.
+In general, though, the sample with available yield information cuts down the available sample immensely, which limits the power of the estimation.
+It isn't clear whether unavailable yields numbers is due to some sort of selection or whether crop cuts were only collected randomly from a few households.
 
@@
 
-Our base analysis will be inspired by the work of \citep{Suri2011-oi}.
-This estimation framework imposes a parametric relationship between farmers' absolute advantage (i.e., the component of productivity that is independent of technology) and their comparative advantage (i.e., their relative productivity of adopting the improved technology over not adopting).
+The advantage of our estimation strategy is that we do not rely on identification from an instrument or the structure of a model, but from the fact that the trajectory that a household takes in their adoption decisions through time is an informative source of identifying information.
+See \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} for more information on the assumptions involved for the Group Random Coefficients estimator.
+Table \ref{tbl:trajectories} maps out the
+
 This method seeks to estimate the following equation:
 
-$$
-yit=+ hit + i+ihit + i+it
-$$
+\begin{align}
+y_{it}&=\sum_{\underline{h}\in\mathcal{H}\backslash (1,1)}\mu_{\underline{h}}1\{h_{i}=\underline{h}\}+\sum_{\underline{h}\in\mathcal{H}_{S}}\Delta_{\underline{h}}h_{it}1\{h_{i}=\underline{h}\}+\kappa_{(1,1)}h_{it}1\{\underline{h}=(1,1)\}+\epsilon_{it}.\label{eq:GRC}
+\end{align}
 

### 2021-11-24 results/tables/traj.tex
add trajectory table

\begin{table}
\centering
\caption{Trajectories of Households}
\label{tbl:trajectories}
\begin{tabular}{lr}
\toprule
{} &  trajectories \\
\midrule
[0.0, 0.0, 0.0] &          1971 \\
[1.0, 1.0, 1.0] &           265 \\
[0.0, 0.0, 1.0] &           172 \\
[0.0, 1.0, 1.0] &           145 \\
[0.0, 1.0, 0.0] &           121 \\
[1.0, 1.0, 0.0] &            62 \\
[1.0, 0.0, 0.0] &            57 \\
[1.0, 0.0, 1.0] &            20 \\
\bottomrule
\end{tabular}
\end{table}

### 2021-11-24 paper.tex
Updates from Overleaf

@@
 See \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} for more information on the assumptions involved for the Group Random Coefficients estimator.
-Table \ref{tbl:trajectories} maps out the
+
+Table \ref{tbl:Trajectories} maps out the frequencies of the different trajectories present in the data, sorted by frequency.
+The largest portion of the data are those that do not adopt improved maize seed at all, followed with sharp decline, to those that use improved maize seed in every period.
+Interestingly, adoption trajectories seem to be more popular for later adoptions than earlier ones.
+
+\input{results/tables/traj}
 

### 2021-11-25 paper.tex
Updates from Overleaf

@@
 Food security is a major challenge in many East African countries-- Ethiopia is no exception, having historically struggled to provide an adequate and reliable food supply \citep{Ramakrishna2002-hv, Jaleta2018-oj}.
-Since the drought of 1984, Ethiopia has become increasingly reliant on maize.
-Despite numerous technological breakthroughs in maize germplasm with increased yields and drought tolerance, adoption of improved seed has remained a challenge.
+Since the drought of 1984, Ethiopia has become increasingly reliant on maize, as of today, it constitutes the most widespread crop in the country.\footnote{FAO stats data shows the total area cultivated with Maize in 2019 is 25\% larger than the one cultivated with the next most widespread crop (sorghum), and 27\% and 139\% larger than the next two following crops, wheat and Barley, respectively.} Despite numerous technological breakthroughs in maize germplasm with increased yields and drought tolerance, adoption of improved seed has remained a challenge.
+As Figure \ref{fig:maize_yields}
+
 There is a broad literature on why sub-Saharan farmers continue to use traditional farming techniques when more modern, higher-return agricultural technologies are available.
@@
 In our paper, we build on that literature by investigating hybrid maize seed adoption in Ethiopia, focusing on how heterogeneous returns are also driven by climate and how this can affect dis-adoption decisions.
+
+\begin{figure}
+    \centering
+    \includegraphics{results/figures/Maize_yields.pdf}
+    \caption{Maize: Total area harvested and Average yields}
+    \figcaption{Source: Authors' calculations using FAO stats data
+    Note: Yield means the harvested production per ha for the area under cultivation}
+    \label{fig:maize_yields}
+\end{figure}
 

### 2021-11-30 paper.tex
Updates from Overleaf

--- 2021-11-24 paper.tex
@@
 
+\begin{itemize}
+    \item Tim Weiss- hybrid subsidies
+    \item is there such a thing as traditional seeds?
+    \item Let's evaluate the effect of adoption in the context of Ethiopia
+    \item find ambiguous results, Why?
+    \item have access to DNA fingerprinting data
+    \item what do they say about adoption?
+    \item turns out that "adoption" isn't really a reliable concept/variable
+    \item most farmers are "adopting" under pretty conservative purity definitions
+    \item Let's see what happens with different varieties of seed, does adoption of that seed lead to positive yield effects?
+    \item can't use these methods but newer data has access to newer varieties where we're relatively more confident
+\end{itemize}
+
 \section{Introduction}
@@
 Food security is a major challenge in many East African countries-- Ethiopia is no exception, having historically struggled to provide an adequate and reliable food supply \citep{Ramakrishna2002-hv, Jaleta2018-oj}.
-Since the drought of 1984, Ethiopia has become increasingly reliant on maize.
-Despite numerous technological breakthroughs in maize germplasm with increased yields and drought tolerance, adoption of improved seed has remained a challenge.
-There is a broad literature on why sub-Saharan farmers continue to use traditional farming techniques when more modern, higher-return agricultural technologies are available.
-Several answers have been proposed to explain this puzzle.
+Since the drought of 1984, Ethiopia has become increasingly reliant on maize, as of today, it constitutes the most widespread crop in the country.\footnote{FAO stats data shows the total area cultivated with Maize in 2019 is 25\% larger than the one cultivated with the next most widespread crop (sorghum), and 27\% and 139\% larger than the next two following crops, wheat and Barley, respectively.} As Figure \ref{fig:maize_yields} shows, there is an upward trend in the total area cultivated by maize, with average yields that have almost double during the period.
+
+%%%%% The puzzle 
+Despite numerous technological breakthroughs in maize germplasm with increased yields and drought tolerance, adoption of improved seed has remained a challenge, with only 13\% of maize farmers consistently using improved seeds varieties, and the lion share of farmers (62\%) not using improved seeds at all during the period where data is available.
+Moreover, there is a large percentage of farmers who adopt new varieties inconsistency limiting their opportunities to learn about their benefits, see Table \ref{tbl:Trajectories}.
+
+The facts sub-Saharan farmers continue to use traditional farming techniques when more modern, higher-return agricultural technologies are available remains an empirical puzzle where several answers have been proposed.
 These include imperfections in credit markets \citep{Croppenstedt2003-pq}, property rights \citep{Place2000-el}, social learning \citep{Conley2010-ue,Foster1995-bz,Munshi2004-og}, lack of commitment devices \cite{Duflo2009-iv}, and high transportation costs that increase the cost of agricultural inputs \citep{Byerlee2013-qk}.
@@
 
+In this paper, we revisit the mechanism proposed by  \citep{Suri2011-oi} that justifies this lack of technology adoption on the existence of farmers heterogeneity.
+In particular, the fact that farmers with high net returns to the technology use new technologies while those with low returns do not.
 Adoption of a technology involves several dimensions such as market integration, access to supplementary inputs like fertilizer, as well agro-ecological considerations such as climate, altitude and soil quality.
@@
 Similar work in Ethiopia on improved chickpea varieties suggests that the economic returns to adoption plays a major role in the decision to adopt \citep{Michler2018-wk}.
+
 In our paper, we build on that literature by investigating hybrid maize seed adoption in Ethiopia, focusing on how heterogeneous returns are also driven by climate and how this can affect dis-adoption decisions.
+We extend the role of heterogeneity in a household’s decision to adopt new technologies in four fronts.
 
-We will explore how the role of heterogeneity in a household’s comparative advantage explains differences in the returns to adoption, and what drives those differences.
-Our objectives center around four main questions: (1) does adoption of improved seed take place where the net benefits of adoption is highest?
-(2) How do adaptation strategies such as water storage, and irrigation augment a household’s comparative advantage in adoption?, (3) what are the drivers of dis-adoption of improved maize seed?
+% We describe what we do here and what we find. We can expand as we have results
+1) First, we test if the adoption of improved seed varieties takes place where the net benefits of adoption is highest?
+
+2) Second, we test how the adoption decision correlates with other adaptation strategies such as water storage, and irrigation augment a household’s comparative advantage in adoption.
+
+3) Third.
+what are the drivers of dis-adoption of improved maize seed?
 and (4) is it possible to extend the methodology in \citep{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} by including time-varying characteristics to the Group Random Coefficients (GRC) model?
+
+% Persistent lack of adoption is a reflection of the distribution of (observable and unobservable) costs and benefits of the technology.
+The approach here models households’ adoption decisions in an environment with household-specific heterogeneity in the costs and benefits, and hence profits, to the technology.
+I estimate how the returns to the technology vary across farmers and then compare these returns to the adoption decisions of farmers.
+I find that farmers with low (or zero) returns to the technology are precisely the farmers who do not adopt it.
+In particular, I use a generalized
+
+\begin{figure}[H]
+    \centering
+    \includegraphics{results/figures/Maize_yields.pdf}
+    \caption{Maize: Total area harvested and Average yields}
+    \figcaption{Source: Authors' calculations using FAO stats data
+    Note: Yield means the harvested production per ha for the area under cultivation}
+    \label{fig:maize_yields}
+\end{figure}
 
@@
 Despite such positive impacts, the low adoption levels of improved seeds varieties in Sub-Saharan Africa remains a puzzle.
+
 Multiple explanations have been proposed for observed adoption rates, including the type of technology promotion mechanisms used by governments, NGO’s, and businesses, the characteristics of the improved seeds and their adaptation to specific agro-climatic environments \citep{Bird2020-nt}, as well as growing conditions, such as the type of soil, input use, and other farmer characteristics \citep{Munshi2004-og}.
+The consensus in this literature is that the success of improved varieties depends on a range of factors that need to be better understood.
+Success of improved seeds should be accompanied by complementary conditions and inputs that would allow the full potential of these varieties to be realized.
 Specifically, extension services and on-farm field trials, seed variety characteristics and rainfall have been found to be crucial to the adoption of improved maize in Tanzania \citep{Kaliba2000-jh}.
-In Ethiopia, labor, fertilizer use, farmers’ experience with extension packages, rainfall suitability and prices have been shown to increase the opportunity cost of not adopting improved wheat varieties \citep{Wale2006-bv}.
+In the specific case of Ethiopia, labor, fertilizer use, farmers’ experience with extension packages, rainfall suitability and prices have been shown to increase the opportunity cost of not adopting improved wheat varieties \citep{Wale2006-bv}.
 Market access and the accessibility of extension services are also critical in the adoption of improved chickpea varieties in Ethiopia \citep{Verkaart2019-ol}.
 Other important determinants of heterogeneous effects include risk aversion \citep{Holden2016-vy}, farm size \citep{Ghimire2015-bd}, and credit constraints \citep{Simtowe2008-jn,Balana2020-hx}.
-The consensus in this literature is that the success of improved varieties depends on a range of factors that need to be better understood.
-Success of improved seeds should be accompanied by complementary conditions and inputs that would allow the full potential of these varieties to be realized.
+
 A different branch of the literature provides an alternative explanation for such low adoption rates, focusing on heterogeneous potential returns to adoption \citep{Suri2011-oi}.
@@
 These methods ignore the possibility of a farmer’s comparative advantage of using an improved technology over a traditional one as playing a central role in their adoption decision.
+
 We rely on Suri’s method and its extension in \citep{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} to account for such heterogeneous comparative advantages and evaluate how these are explained by climatic conditions.
 In Ethiopia, rainfall deviations from historical averages are found to be the main driver of heterogeneous effects for adopting agronomic packages \citep{Marenya2020-kb}.
-In Malawi, exposure to past droughts is found to increase the adoption of drought-tolerant maize among farmers that also receive subsidies \citep{Katengeza2019-af} and the dis-adoption of traditional local maize \citep{Holden2016-vy}.
+Similar experiences from Malawi show exposure to past droughts is found to increase the adoption of drought-tolerant maize among farmers that also receive subsidies \citep{Katengeza2019-af} and the dis-adoption of traditional local maize \citep{Holden2016-vy}.
+
 We will contribute to this literature by considering historical climate averages on the suitability of improved varieties to understand whether adoption is indeed realized where its potential is highest.
@@
 
-The current study uses the Ethiopia Socioeconomic Survey (ESS), a part of the Agricultural Sample Survey (AgSS), a survey 
-designed to obtain production estimates for the major crops.
-It was conducted in collaboration with the Central Statistics Agency of Ethiopia, the World Bank and CGIAR (Consortium of International Agricultural Research Centers) \citep{kosmowski2020shining}.
-The ESS is meant to be representative of the most populous regions of the country.
+The current study uses the Ethiopia Socioeconomic Survey (ESS), a part of the Agricultural Sample Survey (AgSS), a survey designed to obtain production estimates for the major crops.
+The survey is a join effort of the the Central Statistics Agency of Ethiopia, the World Bank and CGIAR (Consortium of International Agricultural Research Centers) \citep{kosmowski2020shining} that aims at collecting information from a representative sample of households in the  the most populous regions of the country.
 The ESS contains several questionnaires targeted at getting an understanding of the state of agriculture.
@@
 These surveys are more important to our analysis as it gives a close look to inputs used in agriculture, the types of seeds used (in various levels of detail) and yields collected post-harvest.
+
 There were significant differences between the analysis carried out in ESS 1-3 and the most recent, ESS4.
@@
 It isn't clear whether unavailable yields numbers is due to some sort of selection or whether crop cuts were only collected randomly from a few households.
+
+\begin{figure}
+    \centering
+    \includegraphics{results/figures/adoption_r4.png}
+    \caption{Adoption rates of improved varieties in 2018 (by self-report, purity level, crop type, source and year of release)}
+    \figcaption{Note: Shares represent the proportion of adopters among maize farmers.
+Acronyms include: SR1 - Self reported 1 [=1 if New, 2nd gen or recycled, =0 if Traditional]; SR2 - Self reported 1 [=1 if New, 2nd gen or recycled, =0 if Traditional]; DNA70 - Improved if purity $>=$70$\%$; DNA90 - Improved if purity $>=$90$\%$; DNA95 - Improved if purity $>=$95$\%$; HYB - Improved DNA, [=1 Hybrid, 0=Open pollinated]; DTM - Drought tolerant Maize [=1 DTM, 0=otherwise], CGIAR source [=1 if germplasm associated to a CGIAR-related thread], Exotic source [=1 if germplasm associated to an exotic source], Released 2000+ [=1 if DNA associated to a variety released on year 2000 or later], Released 2010+ [=1 if DNA associated to a variety released on year 2010 or later].
+QPM is not included as there are too few households planting QPM seeds (n=6).}
+    \label{fig:adoption_r4}
+\end{figure}
 
@@
 
+\input{results/tables/switch1}
+\input{results/tables/switch2}
+\input{results/tables/switch3}
+
 \section{Conclusion}

### 2021-11-30 results/tables/switch1.tex

\begin{table}[htbp]
\centering
\hspace*{-1.2cm}
\begin{threeparttable}
\caption{Effects on yields (log) from adopting improved maize varieties (self-reported)}
\label{tab:switching}
\begin{tabular}{l cccccc}
\hline
\hline
            &Adopters yields&Non-adopters yields&        ATE&          SE&     p-value\\
\hline
\textit{SR1, SR yields}&            &            &            &            &            \\
ATT         &        7.30&        9.07&       -1.76&       0.011&      0.0000\\
%
%
%
ATU         &        9.06&        6.87&        2.19&       0.025&      0.0000\\
%
%
%
\textit{SR1, cropcut yields}&            &            &            &            &            \\
ATT         &        7.57&        7.86&       -0.30&       0.018&      0.0000\\
%
%
%
ATU         &        8.89&        7.11&        1.78&       0.040&      0.0000\\
%
%
%
\textit{SR1, SR yields, full set of controls}&            &            &            &            &            \\
ATT         &        7.28&        7.82&       -0.54&       0.019&      0.0000\\
%
%
%
ATU         &        9.65&        6.85&        2.80&       0.099&      0.0000\\
%
%
%
\textit{SR1, cropcut yields, full set of controls}&            &            &            &            &            \\
ATT         &        7.56&        7.61&      -0.049&       0.020&      0.0133\\
%
%
%
ATU         &        9.24&        7.11&        2.13&       0.084&      0.0000\\
%
%
%
\textit{SR2, SR yields}&            &            &            &            &            \\
ATT         &        7.30&        7.85&       -0.55&       0.020&      0.0000\\
%
%
%
ATU         &        9.63&        6.88&        2.75&       0.104&      0.0000\\
%
%
%
\textit{SR2, cropcut yields}&            &            &            &            &            \\
ATT         &        7.61&        7.62&      -0.013&       0.019&        0.47\\
%
%
%
ATU         &        9.35&        7.14&        2.21&       0.092&    2.7e-127\\
\hline
\hline
\end{tabular}
\begin{tablenotes}
\footnotesize
\item{Note: Full set of controls include: parcel size (in HA), household labor, hired labor, fertilizer costs, other input costs irrigation (dummy), mechanization (dummy), organic fertilizer (dummy). Instruments for the adoption equation include: years of education of hh head, age of hh head, female head, land title, asset index, seed costs. Results for SR2 include full set of controls}
\end{tablenotes}
\end{threeparttable}
\end{table}

### 2021-12-14 results/tables/switch1.tex
Updates from Overleaf

@@
-\begin{table}[htbp]
+\begin{table}[H]
 \centering
@@
 \caption{Effects on yields (log) from adopting improved maize varieties (self-reported)}
-\label{tab:switching}
+\label{tab:switch1}
 \begin{tabular}{l cccccc}
@@
 %
-ATU         &        9.06&        6.87&        2.19&       0.025&      0.0000\\
+ATU         &        9.06&        6.87&        2.19&       0.024&      0.0000\\
 %
@@
 \textit{SR1, cropcut yields}&            &            &            &            &            \\
-ATT         &        7.57&        7.86&       -0.30&       0.018&      0.0000\\
+ATT         &        7.57&        7.86&       -0.30&       0.017&      0.0000\\
 %
@@
 %
-ATU         &        8.89&        7.11&        1.78&       0.040&      0.0000\\
+ATU         &        8.89&        7.11&        1.78&       0.039&      0.0000\\
 %
@@
 %
-ATU         &        9.65&        6.85&        2.80&       0.099&      0.0000\\
+ATU         &        9.65&        6.85&        2.80&       0.100&      0.0000\\
 %
@@
 \textit{SR1, cropcut yields, full set of controls}&            &            &            &            &            \\
-ATT         &        7.56&        7.61&      -0.049&       0.020&      0.0133\\
+ATT         &        7.56&        7.61&      -0.049&       0.019&      0.0087\\
 %
@@
 %
-ATU         &        9.24&        7.11&        2.13&       0.084&      0.0000\\
+ATU         &        9.24&        7.11&        2.13&       0.085&      0.0000\\
 %
@@
 %
-ATU         &        9.63&        6.88&        2.75&       0.104&      0.0000\\
+ATU         &        9.63&        6.88&        2.75&       0.101&      0.0000\\
 %
@@
 %
-ATU         &        9.35&        7.14&        2.21&       0.092&    2.7e-127\\
+ATU         &        9.35&        7.14&        2.21&       0.091&    1.6e-130\\
 \hline
@@
 Instruments for the adoption equation include: years of education of hh head, age of hh head, female head, land title, asset index, seed costs.
-Results for SR2 include full set of controls}
+Results for SR2 include full set of controls.
+SR1=1 if New, 2nd gen or recycled,=0 if Traditional; SR2=1 if New, =0 2nd gen, recycled or traditional}
 \end{tablenotes}

### 2021-12-14 paper.tex
Updates from Overleaf

@@
 \usepackage{amsmath}
+\usepackage{longtable}
+\usepackage{graphicx}
+\usepackage{float}
 
@@
 
+%%%% I added the bullet points where I think they fit what we have as an introduction
+
 \begin{itemize}
@@
     \item is there such a thing as traditional seeds?
+
     \item Let's evaluate the effect of adoption in the context of Ethiopia
@@
 
+%%% Why it is important
 Food security is a major challenge in many East African countries-- Ethiopia is no exception, having historically struggled to provide an adequate and reliable food supply \citep{Ramakrishna2002-hv, Jaleta2018-oj}.
@@
 
+%%%% Let's evaluate the effect of adoption in the context of Ethiopia
+
 In this paper, we revisit the mechanism proposed by  \citep{Suri2011-oi} that justifies this lack of technology adoption on the existence of farmers heterogeneity.
@@
 
-In our paper, we build on that literature by investigating hybrid maize seed adoption in Ethiopia, focusing on how heterogeneous returns are also driven by climate and how this can affect dis-adoption decisions.
-We extend the role of heterogeneity in a household’s decision to adopt new technologies in four fronts.
+%% Expand a bit on the methods
+In our paper, we build on that literature by investigating hybrid maize seed adoption in Ethiopia, focusing on how heterogeneous returns XXXXXXXXXXXXXXXX
+
+%% What we find
+
+Our results show evidence that showcase that the average yields without adoption for
+
+return to adoption
+% Evidence of heterogeneity in response to hybrid adoption for joiners and leavers, however their average yield without adoption ($\mu_{01}$,$\mu_{10}$) are very similar.\\
+
+% are also driven by climate and how this can affect dis-adoption decisions. We extend the role of heterogeneity in a household’s
+% decision to adopt new technologies in four fronts.
 
@@
 There were significant differences between the analysis carried out in ESS 1-3 and the most recent, ESS4.
-Most notably, the ESS4 is no longer a panel dataset made up of the households of ESS 1-3, but it includes important information on the DNA fingerprinting of seed type and misclassification rates.
+Most notably, the ESS4 is no longer a panel dataset made up of the households of ESS 1-3, but it includes important information on the DNA fingerprinting of seed type, with which we can calculate misclassification rates.
 This is important information as the first three waves of ESS only included self-reported hybrid maize status.
 Since our estimation strategy requires a balanced panel dataset of maize growers, our main analysis will include the first three waves of the ESS.
-We use the fourth wave later for robustness and validation.
+We use the fourth wave later for validation, and to explore the impact of specific types of improved seeds.
 
@@
 
+% \resizebox{0.7\textwidth}{!}{
+{\small\tabcolsep=3pt  % hold it local
     \input{results/tables/summary}
+}
 
@@
 
+\subsection{Improved seeds}
+
+Defining what constitutes an improved seed is challenging because of the different characteristics of seeds.
+A crop seed is considered “improved” in Ethiopia if it was tested by breeders and evaluated to be superior to existing varieties (MoA, 2013, SPIA, 2020).
+In Ethiopia, the government controls most of the seed system.
+Local agricultural offices or extension agents assess the demands of different varieties, the MoA decides the seed quantities to be produced, and actors such as regional seed enterprises, seed companies, seed unions, are involved in the production of certified seeds; and seeds are primarily distributed by seed producer cooperatives (SPCs), which account for 37\% of total distribution in the country (SPIA, 2020).
+
+An improved seed can be obtained directly from such distribution centers (which would distribute a first generation seed), or it could be cultivated from second or posterior generations from the originally bred seed.
+Farmers are asked to report whether the variety they used was a first-generation improved seed, second generation,  recycled, or a traditional variety.
+When buying seeds, farmers might be able to identify the variety they are looking for, but it is unlikely that they would be aware of the source.
+
+Fifty four varieties have been released in Ethiopia since 1990, out of which seven were released between 1990 to 1999, 29 between 2000 to 2009, and 18 between 2010 to 2019.
+Thirty four of such varieties contain CIMMYT germplasm, out of which, 10 Drought-tolerant maize varieties (DTMZ) and 8 Quality protein maize (QPM) were released (most of these between 2010 to 2019).
+Out of the CIMMYT-related germplasm, ten are Open-pollinated varieties (OPVs) and 25 are hybrid.
+The fourth wave of the ESS introduced for the first time DNA fingerprinting for barley, maize, and sorghum.
+The collection of crop-cuts for crop yield estimation facilitated the process for DNA fingerprinting.
+For the crop-cuts, a 4 square meter quadrant was randomly selected in each plot, so that total production harvested from such patch was weighted (both when freshly harvested and when dried).
+The DNA material was extracted from samples from such dried crop-cuts, that were then matched with the genetic material of the maize DNA reference library for Ethiopia, obtained from a CIMMYT and EIAR project.
+The collection of DNA samples focused only on eight regions.
+In each EA of such regions, a random sample of plots was selected, of a maximum of 10 fields for each crop.
+As a result, 505 plots were selected for DNA fingerprinting, which corresponded to 447 households.
+
+The DNA fingerprinting process matched the genetic material of each sampled harvest with that of the reference library.
+The resulting matching process returns the name of the crop variety matched, its germplasm origin or their pedigree (whether they are a CIMMYT line, EIAR, IITA or private), the breeder or maintainer, and the year of their release (which could be indicative of the number of generations a seed had been cultivated), as well as the genetic purity of the sample (the degree of no contamination by other genetic varieties).
+It also informs whether the harvested seeds are open-pollinated varieties (OPVs) or hybrids; drought-tolerant maize varieties (DMTZ) or quality protein maize (QPM).
+This rich information about seed varieties provides a unique opportunity to further understand impacts differentiated by type of seed, degree of purity, year of release or source.
+Figure \ref{fig:adoption_r4} shows adoption rates in the fourth round for improved varieties, using the above characteristics as different definitions.
+
+We find that 38\% of households self-report adopting a “new” improved variety (SR2 in Figure \ref{fig:adoption_r4}), and 43\% report adopting a “new”, “second generation” or “recycled” improved variety (SR1).
+Considering a household as adopter if the household planted an improved variety in at least one of their plots, all households where a DNA sample was obtained used a maize improved seed, considering a purity threshold of 70\% (DNA 70).
+When purity thresholds of 90 and 95\% are considered (DNA 90, DNA 95 in Figure \ref{fig:adoption_r4}, respectively), adoption rates change to 91 and 54\% of households, respectively.
+62\% of households adopted a High-yielding variety, compared to 38\% that adopted an open-pollinated variety, and 15\% of households adopted a Drought-tolerant maize variety.
+63\% of households that adopted an improved variety are using a CGIAR-derived germplasm, and 79\% are using an exotic germplasm.
+Finally, 64\% of households adopted a variety that was released on the year 2000 or after, and 24\% adopted a variety that was adopted in the year 2010 or after.
+
 \begin{figure}
@@
     \figcaption{Note: Shares represent the proportion of adopters among maize farmers.
-Acronyms include: SR1 - Self reported 1 [=1 if New, 2nd gen or recycled, =0 if Traditional]; SR2 - Self reported 1 [=1 if New, 2nd gen or recycled, =0 if Traditional]; DNA70 - Improved if purity $>=$70$\%$; DNA90 - Improved if purity $>=$90$\%$; DNA95 - Improved if purity $>=$95$\%$; HYB - Improved DNA, [=1 Hybrid, 0=Open pollinated]; DTM - Drought tolerant Maize [=1 DTM, 0=otherwise], CGIAR source [=1 if germplasm associated to a CGIAR-related thread], Exotic source [=1 if germplasm associated to an exotic source], Released 2000+ [=1 if DNA associated to a variety released on year 2000 or later], Released 2010+ [=1 if DNA associated to a variety released on year 2010 or later].
+Acronyms include: SR1 [=1 if New, 2nd gen or recycled, =0 if Traditional]; SR2  [=1 if New, =0 if 2nd gen, recycled or traditional]; DNA70 - Improved if purity $>=$70$\%$; DNA90 - Improved if purity $>=$90$\%$; DNA95 - Improved if purity $>=$95$\%$; HYB - Improved DNA, [=1 Hybrid, 0=Open pollinated]; DTM - Drought tolerant Maize [=1 DTM, 0=otherwise], CGIAR source [=1 if germplasm associated to a CGIAR-related thread], Exotic source [=1 if germplasm associated to an exotic source], Released 2000+ [=1 if DNA associated to a variety released on year 2000 or later], Released 2010+ [=1 if DNA associated to a variety released on year 2010 or later].
 QPM is not included as there are too few households planting QPM seeds (n=6).}
@@
 
+These ratios are in line with prior findings.
+In the regions of Oromia, Amhara, Tigray, and SNNPR, adoption rates of improved maize varieties of 39\% of households were found, according to self-reported information (Zeng et al., 2015).
+In the regions of Amhara, Benishangul-Gumuz, Oromia, SNNPR, and Tigray, adoption rates of improved maize varieties were found for 27\% of households (Jaleta et al., 2018).
+In Oromia, SNNPR, and Benishangul-Gumuz adoption of improved varieties per farmers’ self-reports was found for 84\% of maize growers in 2011, and 88\% of maize growers in 2013 (Yirga et al., 2017).
+
+Understanding the effects of improved seed adoption is not only challenging because there could be multiple ways of establishing what constitutes an improved seed, but also because such definitions may or may note overlap, or because farmers may also missclassify a variety as improved.
+Figure \ref{fig:missclassification} shows the correlation between the different definitions discussed above.
+For those seeds for which genetic material was obtained, the correlation with the self-reported seeds indicator is low, especially when using lower cutoff thresholds.
+
+\begin{figure}
+    \centering
+    \includegraphics{results/figures/missclassification.png}
+    \caption{Correlation of different definitions of improved seeds, 4th round}
+    \figcaption{Note: Shares represent the proportion of adopters among maize farmers.
+Acronyms include: SR1 [=1 if New, 2nd gen or recycled, =0 if Traditional]; SR2  [=1 if New, =0 if 2nd gen, recycled or traditional]; DNA70 - Improved if purity $>=$70$\%$; DNA90 - Improved if purity $>=$90$\%$; DNA95 - Improved if purity $>=$95$\%$; HYB - Improved DNA, [=1 Hybrid, 0=Open pollinated]; DTM - Drought tolerant Maize [=1 DTM, 0=otherwise], CGIAR source [=1 if germplasm associated to a CGIAR-related thread], Exotic source [=1 if germplasm associated to an exotic source], Released 2000+ [=1 if DNA associated to a variety released on year 2000 or later], Released 2010+ [=1 if DNA associated to a variety released on year 2010 or later].
+QPM is not included as there are too few households planting QPM seeds (n=6).}
+    \label{fig:missclassification}
+\end{figure}
+
 \section{Estimation Strategy}
 
-The advantage of our estimation strategy is that we do not rely on identification from an instrument or the structure of a model, but from the fact that the trajectory that a household takes in their adoption decisions through time is an informative source of identifying information.
+\subsection{Group Random Coefficient Model}
+
+The advantage of our estimation strategy is that we do not rely on identification from a valid instrument, but modelling the heterogeneity directly through a projection of their returns to adoption on their observed "trajectory" of adoption.
+This allows us to directly capture two effects: the heterogeneous effects of the so-called "comparative advantage" of adoption, and, under the assumption of linearity in comparative advantage, a description of the overall economy's sorting of farmers based on their comparative advantage.
+The first is important because it is able to disentangle the effect of a general household's agricultural ability (which can be thought of as a fixed effect) and their adoption-specific potential.
+The sorting parameter can tell us whether the economy sorts such that those that would benefit the most adoption actually adopt, or whether there are barriers that stop that sorting from happening.
 See \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} for more information on the assumptions involved for the Group Random Coefficients estimator.
+In our particular case, we contribute to the li
+
+We estimate comparative advantage using the restricted model outlined in \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc}:
+
+\begin{align}
+y_{it}&=\sum_{\underline{h}\in\mathcal{H}\backslash (1,1)}\mu_{\underline{h}}1\{h_{i}=\underline{h}\}+\sum_{\underline{h}\in\mathcal{H}_{S}}\Delta_{\underline{h}}h_{it}1\{h_{i}=\underline{h}\}+\kappa_{(1,1)}h_{it}1\{\underline{h}=(1,1)\}+\epsilon_{it}.\label{eq:GRC}
+\end{align}
+
+where $y$ is log yields,
 
@@
 This method seeks to estimate the following equation:
-
-\begin{align}
-y_{it}&=\sum_{\underline{h}\in\mathcal{H}\backslash (1,1)}\mu_{\underline{h}}1\{h_{i}=\underline{h}\}+\sum_{\underline{h}\in\mathcal{H}_{S}}\Delta_{\underline{h}}h_{it}1\{h_{i}=\underline{h}\}+\kappa_{(1,1)}h_{it}1\{\underline{h}=(1,1)\}+\epsilon_{it}.\label{eq:GRC}
-\end{align}
 
@@
 
-Implications
+\subsection{Endogenous Switching regression}
+The ESS4 is a unique data-set that includes DNA fingerprinting of maize seed type. This brings new possibilities in terms of exploring.... sayESS4 no longer part of panel but that still useful,....//
+The group random coefficient model used the variation in the adoption trajectories of each observation in the ESS 1-3 panel as a source of information that allows for an unbiased estimation.
+ESS4 is no longer part of this panel data-set, hence, in order to estimate the effect of the adoption of improved seeds on yields, an alternative estimation method is needed.
+We propose an endogenous switching regression model which accounts for unobserved heterogeneity in adoption decisions.
+This method, although it imposes restrictive assumptions regarding the distribution of errors
+
+\section{Results}
+
+\subsection{Group Random Coefficients}
+
+% {\small\tabcolsep=3pt  % hold it local
+\resizebox{1\textwidth}{!}{
+\input{results/tables/grc_YIELD_cropcutdry_tr}
+}
+
+\resizebox{1\textwidth}{!}{
+\centering
+% \input{results/tables/phi_test_merged}
+\input{results/tables/phi_test_self}
+\input{results/tables/phi_test_cropcut}
+}
+% \input{results/tables/grc_YIELD_selfr_tr}
+
+\subsection{Endogenous Switching Regression}
+
+We first examine the results of the Endogenous Switching Regression using a self-reported definition of an improved variety.
+As before, we consider a more loose definition of improved variety, which includes new seeds but also second-generation and recycled (SR1), and a more strict definition, which includes only new varieties (SR2).
+Table \ref{tab:switch1} shows the results of the Switching regression using both definitions, and using self-reported and cropcut yields.
+We see that regardless of the definition of improved seed, yields measurement form and types of controls included, the results consistently show that the Average Treatment effect on the Treated (ATT) is negative and the Average Treatment effect on the Untreated (ATU) is positive, meaning that farmers that adopted a self-reported improved seed would've been better off not adopting, and farmers that did not adopt a self-reported improved seed would've been better off adopting.
+
+\input{results/tables/switch1}
+
+To better understand the above puzzling results, we explore similar yield estimations using the different definitions of improved seeds from the DNA fingerprinting.
+We run similar switching regressions using different thresholds of DNA purity (70, 90 and 95\% purity), for hybrid seeds (compared with open-pollinated varieties), and for Drought-tolerant maize (DTMZ).
+Table \ref{tab:switch2} shows the results.
+
+Given that all the plots sampled for DNA fingerprinting showed a purity percent of 70\% or higher, we need a counter-factual to compare these purity thresholds with non-adopters.
+We therefore consider as non-adopters those that self-reported using a traditional variety.
+We are aware that given the universality of improved varieties with a 70\% threshold or higher and given the high rates of missclassification, the control group for a 70\% threshold is likely to be contaminated with a high share of false-positives.
+But we also know the results are biased downwards and that the control group for higher shares of purity will be less contaminated, which may provide an interesting comparison.
+
+Table \ref{tab:switch2} shows that the estimations for purity thresholds of 70 and 90\% show similar patterns as before: non-adopters would've been better off adopting and adopters would've been better off not adopting.
+But when improved seeds with a purity of 95\% or higher are compared with self-reported traditional varieties or lower purity than 95\%, the ATT becomes positive and the ATU is still positive but significantly decreases (and it is even negative when self-reported yields are used).
+Farmers that adopted a variety with a 95\% purity or greater had significantly greater yields than if they had not adopted, considering that this is even a lower boundary.
+
+We get similar results for DTMZ.
+Farmers that adopted DTMZ (as tracked by the DNA germplasm) did obtain greater yields than if they had not adopted.
+Again this result may be biased downwards because of possible false-positives among the control group.
+The results for the comparison between hybrid and open-pollinated varieties are not significant (when using self-reported yields).
+
+\input{results/tables/switch2}
+
+To further disaggregate the results, we finally explore the source of the improved varieties (CGIAR or exotic) and the year of the seed release.
+We find positive effects for adopters of CGIAR-sourced varieties when using self-reported yields.
+The results seem to be contradictory when using cropcut yields, but we should consider these are different samples, as cropcuts were obtained in a subsample of plots.
+Exotic-sourced varieties also show positive yield effects for adopters; and a negative ATU, meaning non-adopters are also better off not adopting.
+Newly released-seed varieties do consistently show positive ATT, when using both self-reported and cropcut yields.
+
+\input{results/tables/switch3}
+
+\section{Discussion and Implications}
+
+- Why do we see that increasing purity leads to positive ATE
+- may be due to improved seed being more effective, but doesnt tell the whole story
+- barrier to purchasing the seed and associated inputs for poorer households, leading to reuse of seeds and being less effective
 
@@
 
-\section{Results}
-
-\input{results/tables/switch1}
-\input{results/tables/switch2}
-\input{results/tables/switch3}
-
 \section{Conclusion}

### 2021-12-14 paper.tex
Updates from Overleaf

@@
 %%% Why it is important
-Food security is a major challenge in many East African countries-- Ethiopia is no exception, having historically struggled to provide an adequate and reliable food supply \citep{Ramakrishna2002-hv, Jaleta2018-oj}.
-Since the drought of 1984, Ethiopia has become increasingly reliant on maize, as of today, it constitutes the most widespread crop in the country.\footnote{FAO stats data shows the total area cultivated with Maize in 2019 is 25\% larger than the one cultivated with the next most widespread crop (sorghum), and 27\% and 139\% larger than the next two following crops, wheat and Barley, respectively.} As Figure \ref{fig:maize_yields} shows, there is an upward trend in the total area cultivated by maize, with average yields that have almost double during the period.
+Food security is a major challenge in many East African countries, and Ethiopia is no exception, having historically struggled to provide an adequate and reliable food supply \citep{Ramakrishna2002-hv, Jaleta2018-oj}.
+Since the drought of 1984, Ethiopia has become increasingly reliant on maize, as of today, it constitutes the most widespread crop in the country.\footnote{FAO stats data shows the total area cultivated with Maize in 2019 is 25\% larger than the one cultivated with the next most widespread crop (sorghum), and 27\% and 139\% larger than the next two following crops, wheat and Barley, respectively.} As Figure \ref{fig:maize_yields} shows, there is an upward trend in the total area cultivated with maize since 1990, with average yields that have almost doubled during this period.
 
 %%%%% The puzzle 
-Despite numerous technological breakthroughs in maize germplasm with increased yields and drought tolerance, adoption of improved seed has remained a challenge, with only 13\% of maize farmers consistently using improved seeds varieties, and the lion share of farmers (62\%) not using improved seeds at all during the period where data is available.
-Moreover, there is a large percentage of farmers who adopt new varieties inconsistency limiting their opportunities to learn about their benefits, see Table \ref{tbl:Trajectories}.
+Despite numerous technological breakthroughs in maize germplasm with increased yields and drought tolerance, adoption of improved seeds has remained a challenge, with only 13\% of maize farmers consistently using improved seeds varieties, and the lion share of farmers (62\%) not using improved seeds at all during the period where data is available.
+Moreover, there is a large percentage of farmers who adopt new varieties inconsistently limiting their opportunities to learn about their benefits, see Table \ref{tbl:Trajectories}.
 
@@
 See \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} for more information on the assumptions involved for the Group Random Coefficients estimator.
-In our particular case, we contribute to the li
+In our particular case, we contribute to the literature and methodology of \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} by providing reasonable ways to aggregate the results of the GRC model, which may seem difficult to interpret in its raw form.
+We use these aggregations to better understand how adoption affects households across time, as well as whether the frequency of adoption has a strong effect.
+
+\input{results/tables/traj}
+
+Table \ref{tbl:Trajectories} maps out the frequencies of the different trajectories present in the data, sorted by frequency.
+The largest portion of the data are those that do not adopt improved maize seed at all, followed with sharp decline, to those that use improved maize seed in every period.
+Interestingly, adoption trajectories seem to be more popular for later adoptions than earlier ones.
+Most households do not adopt at all (``never-adopters''), but the second most frequent trajectory are those that adopt in every wave (``always-adopters'').
+Let $\mathcal{H}$ be the set of these trajectories in the data.
 
@@
 \begin{align}
-y_{it}&=\sum_{\underline{h}\in\mathcal{H}\backslash (1,1)}\mu_{\underline{h}}1\{h_{i}=\underline{h}\}+\sum_{\underline{h}\in\mathcal{H}_{S}}\Delta_{\underline{h}}h_{it}1\{h_{i}=\underline{h}\}+\kappa_{(1,1)}h_{it}1\{\underline{h}=(1,1)\}+\epsilon_{it}.\label{eq:GRC}
+y_{it}&=\sum_{\underline{h}\in\mathcal{H}\backslash (1,1,1)}\mu_{\underline{h}}+\Delta_{(0,0,1)}h_{it}+\phi(\mu_{(0,0,1)}-\mu_{(0,0,1)})h_{it}1\{h_{i}=(0,0,1)\}\nonumber\\
+&~~~+\left(\mu_{(1,1,1)}+\phi\left(\mu_{(1,1,1)}-\mu_{(0,0,1)}\right)\right)h_{it}1\{h_{i}=(1,1,1)\} + X_{it}\beta +\varepsilon_{it},\label{eq:GRC_Suri}
 \end{align}
 
-where $y$ is log yields,
+where $y$ is log yields, $\mu_{\underline{h}}$ are binary variables for being in trajectory $\underline{h}$, $h$ is a binary variable for whether household $i$ adopted in wave $t$, $X$ are a set of controls and $\Delta_{0,0,1}$ are the returns to adoption for trajectory $(0,0,1)$.
+For reasons outlined in \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc}, we use a base trajectory.
+$\phi$ is the parameter which summarizes sorting into the economy.
+If $\phi$ is positive, then higher yields is associated with higher comparative advantage, denoting that those that adopt also tend to have the highest returns to adoption.
+If it is negative, however, this denotes that there might be barriers, such as a lack of fertilizer access, and bad roads that restrict the ability to buy hybrid seed.
 
-Table \ref{tbl:Trajectories} maps out the frequencies of the different trajectories present in the data, sorted by frequency.
-The largest portion of the data are those that do not adopt improved maize seed at all, followed with sharp decline, to those that use improved maize seed in every period.
-Interestingly, adoption trajectories seem to be more popular for later adoptions than earlier ones.
+Equation \ref{eq:GRC_Suri} can be estimated by way of General Method of Moments.
+Parameters for returns to adoption and comparative advantage can be post-estimated by the delta method, where the expression for returns of each trajectory can be derived from:
 
-\input{results/tables/traj}
+$$
+\Delta_{\underline{h}}=\phi\left(\mu_{\underline{h}}-\mu_{(0,0,1)}\right) + \Delta_{(0,0,1)}
+$$
 
-This method seeks to estimate the following equation:
+And comparative advantage can be derived from:
 
-where y is yields or profits for household i at time t, h is an indicator for adoption,  is a household’s absolute advantage,  is a household’s comparative advantage and  describes the gains to adopting, relative to a household’s comparative advantage.
-This framework provides an advantage over other methods \citep{Wooldridge1997-xj,Heckman1998-pt}.
-Firstly, identification does not rely on the existence of a valid instrumental variable.
-Instead, identification comes from a linear projection of an individual's returns to adoption onto the observed history of their adoption, an approach similar to the correlated random effects (CRE) method in \citep{Chamberlain1984-uk}.
-Secondly, it disentangles a household’s absolute advantage when taking part in agriculture and its comparative advantage in adopting.
-This is crucial in understanding the adoption decision as it is not dependent on a household’s agricultural ability, but on the gains from taking it up.
-We and co-authors developed this further in \citep{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} by proposing a more flexible and tractable approach using a group random coefficient strategy that draws on recent developments in the nonparametric panel identification literature.
-Our strategy has a clear economic interpretation and can elucidate potential identification concerns of the CRC model to the practitioner.
-It is also more easily extended to multiple time periods and new levels of heterogeneity than the CRC approach.
+$$
+\theta_{\underline{h}} = \mu_{\underline{h}} - E(\mu_{\underline{h}})
+$$
 
-Estimating equation 1 will allow us  to analyze the distribution of the gains to adoption and to create a distribution of heterogeneous returns.
-We will investigate whether comparative advantage is driven by two key policy relevant factors: water conservation strategies and deviation from historical climate averages.
-The first will elucidate if adaptation strategies in water management are associated with higher levels of gains to adoption.
-The second will seek to understand whether households are adapting to climate change through their adoption decisions.
-This will require the extension of the method developed in \citep{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc}, by allowing for time-varying characteristics in the identification of comparative advantage.
+where $E(\mu_{\underline{h}}) = \sum_{\mathcal{H}} \pi_h \cdot \mu_h$ for $\pi_h$ being the frequency weight for trajectory $h$.
+The comparative advantage of each trajectory tells us how much a household with a given trajectory benefits from adoption.
+$\theta <0 $ denotes that returns to adoption are negative and that a household loses by adopting.
+
+Once we generate $\theta$ and $\Delta$ for each trajectory, we can examine heterogeneous effects to adoption. In its raw form
 
@@
 We propose an endogenous switching regression model which accounts for unobserved heterogeneity in adoption decisions.
-This method, although it imposes restrictive assumptions regarding the distribution of errors
+This method, although it imposes restrictive assumptions regarding the distribution of the errors
 

### 2021-12-14 paper.tex

@@
 %% Expand a bit on the methods
-In our paper, we build on that literature by investigating hybrid maize seed adoption in Ethiopia, focusing on how heterogeneous returns XXXXXXXXXXXXXXXX
+In our paper, we build on that literature by investigating hybrid maize seed adoption in Ethiopia, focusing on how heterogeneous returns drive the decision to use improved seed varieties.
+We implement two complementary estimation strategies.
+First, we implement at Group Random Coefficients estimation as originally devised by \cite{Suri2011-oi} and later expanded by \citep{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc}.
+The advantage of our estimation strategy is that we do not rely on identification from a valid instrument,
+but modelling the heterogeneity directly through a projection of their returns to adoption on their observed
+”trajectory” of adoption.
+This first estimation strategy uses data from the first three survey rounds that allow us to follow the same households across time (ESS1 to ESS3).
+The results show returns to adoption are consistently negative, a result that runs against the careful lab development of the improved seed varieties.
+To expand on this results, our second estimation strategy implements an endogenous switching regressions that exploits the last round of the survey (ESS4).
+The latest survey round is not a a continuation of the panel of the first three rounds, but has novel data on the genetic finger prints of the seeds used, that we exploit to understand why hybrid seed varieties seem not to increase returns when adopted.
 
@@
 Most households do not adopt at all (``never-adopters''), but the second most frequent trajectory are those that adopt in every wave (``always-adopters'').
-Let $\mathcal{H}$ be the set of these trajectories in the data.
+Let $\mathcal{H}$ be the set of these trajectories in the data and let $\mathcal{H}_S$ be the set of switcher trajectories, i.e.
+$\mathcal{H}_S = \mathcal{H}\backslash\{(0,0,0), (1,1,1)\}$
 
-We estimate comparative advantage using the restricted model outlined in \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc}:
+To generate results for adoption, we use the unrestricted model as outlined in \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc}:
+
+\begin{align}
+y_{it}&=\sum_{\underline{h}\in\mathcal{H}\backslash (1,1,1)}\mu_{\underline{h}}+\sum_{\underline{h}\in\mathcal{H}_{S}}\Delta_{\underline{h}}h_{it}1\{h_{i}=\underline{h}\} + \kappa_{(1,1,1)}h_{it}1\{\underline{h}=(1,1,1)\}+ X_{it}\beta+\varepsilon_{it}.\label{eq:GRC}
+\end{align}
+
+where $y$ is log yields, $\mu_{\underline{h}}$ are binary variables for being in trajectory $\underline{h}$, $h$ is a binary variable for whether household $i$ adopted in wave $t$, $X$ are a set of controls and $\Delta_{0,0,1}$ are the returns to adoption for trajectory $(0,0,1)$.
+$\kappa_{(1,1,1)}$ is the average yield with adoption of the always-adopters, where $\kappa_{(1,1,1)} = \mu_{(1,1,1)} + \Delta_{(1,1,1)}$.
+It is not possible to separately identify $\mu_{(1,1,1)}$ and $\Delta_{(1,1,1)}$ without adding more structure to the model in the form of the linearity in comparative advantage (LCA) restriction, which we explain how to test for below.
+
+We estimate comparative advantage using the restricted model:
 
@@
 
-where $y$ is log yields, $\mu_{\underline{h}}$ are binary variables for being in trajectory $\underline{h}$, $h$ is a binary variable for whether household $i$ adopted in wave $t$, $X$ are a set of controls and $\Delta_{0,0,1}$ are the returns to adoption for trajectory $(0,0,1)$.
-For reasons outlined in \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc}, we use a base trajectory.
-$\phi$ is the parameter which summarizes sorting into the economy.
+where $\phi$ is the parameter which describes sorting in the economy.
 If $\phi$ is positive, then higher yields is associated with higher comparative advantage, denoting that those that adopt also tend to have the highest returns to adoption.
 If it is negative, however, this denotes that there might be barriers, such as a lack of fertilizer access, and bad roads that restrict the ability to buy hybrid seed.
+$\phi$ is estimated in two ways: through the restricted model just mentioned and by a weak-identification brute force grid procedure, which allows for calculating the confidence interval for $\phi$ for multiple sets of trajectories.
+The advantage of the weak-identification robust procedure is that we can test for whether the data satisfies the conditions necessary for an interpretable $\phi$ parameter.
 
@@
 
-And comparative advantage can be derived from:
+for trajectory $\underline{h} \neq (0,0,1)$.\footnote{Since we use $(0,0,1)$ as our base, we already have returns available for that trajectory, namely $\Delta_{(0,0,1)}$} Comparative advantage can be derived from:
 
@@
 
-Once we generate $\theta$ and $\Delta$ for each trajectory, we can examine heterogeneous effects to adoption. In its raw form
+Once we generate $\theta$ and $\Delta$ for each trajectory, we can examine heterogeneous effects to adoption.
+In its raw form, that means that have eight comparative advantages available for comparison, which is rich in its information, but difficult to interpret as a result, we will aggregate $\theta$ in two ways to improve interpretability: (1) which wave the household adopted, and (2) the number of time adoption took place.
+We do this by taking a simple average, grouped by each of these factors.
 
 \subsection{Endogenous Switching regression}
-The ESS4 is a unique data-set that includes DNA fingerprinting of maize seed type. This brings new possibilities in terms of exploring.... sayESS4 no longer part of panel but that still useful,....//
+The ESS4 is a unique data-set that includes DNA fingerprinting of maize seed type.
+This brings new possibilities in terms of exploring....
+sayESS4 no longer part of panel but that still useful,....
+\par
 The group random coefficient model used the variation in the adoption trajectories of each observation in the ESS 1-3 panel as a source of information that allows for an unbiased estimation.
@@
 We propose an endogenous switching regression model which accounts for unobserved heterogeneity in adoption decisions.
-This method, although it imposes restrictive assumptions regarding the distribution of the errors
+This method imposes restrictive assumptions regarding the distribution of the errors of the selection and output equations.
+At the same time, the exclusion restriction needs to be satisfied in order to reduce the bias of the estimation.
+Nevertheless, the endogenous switching regression holds certain advantages over other methodologies used for identifying unbiased estimates with cross-sectional observational data such as propensity score matching or other instrumental variable approaches (\cite{shiferaw2014adoption}) and has been used to study similar questions in the past (\citealt{falco2011does}, \citealt{kabunga2012yield}).
+\par
+The decision to adopt and improved seed is determined by a series of observed and unobserved variables.
+Equation \ref{eq:switch_latent} shows the definition for a latent variable $A^*$ that reflects the benefit from adopting an improved seed relative to planting a traditional one.
+As shown in the equation, household $i$ will decide to adopt  (A_i=1).
+
+\begin{align}
+A_i^*=Z_i\alpha+\varepsilon_i \; \text{with} \, A_i    \begin{cases}
+      1, & \text{if}\ A_i^*>0 \\
+      0, & \text{otherwise}
+    \end{cases} \label{eq:switch_latent}
+\end{align}
 
@@
 \subsection{Group Random Coefficients}
+
+Our results use two definitions of yields: measured cropcuts of dry maize and self-reported yields maize yields.
+Each measure of yields has its benefits and drawbacks.
+Cropcuts, for instance, provides a true estimate of yields for a given parcel of land, but it is at the cost of a reduced sample size.
+Self-reports have larger sample sizes, but may suffer from measurement error.
+For controls we use the years of education of the household head, the age of the household head, the sex of the household head, whether there is a title for the land being cultivated, the parcel size in hectares, as well as the hours of household and hired labor.
+Commensurate with the literature, we also include interactions of these controls with hybrid adoption as added controls.
+The results of the regression are shown in Table \ref{tbl:unres}.
 
@@
 % \input{results/tables/grc_YIELD_selfr_tr}
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=0.75]{results/figures/theta.png}
+    \caption{Comparative Advantage and Returns by Trajectory (Dry Cropcuts)}
+    \label{fig:theta_delta_raw}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=0.75]{results/figures/traj_sum.png}
+    \caption{Aggregation Results: Comparative Advantage and Returns by the Number of Times Adopted (Dry Cropcuts}
+    \label{fig:traj_sum}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=0.75]{results/figures/when_adopt_cropcut.png}
+    \caption{Aggregation Results: Comparative Advantage and Returns by Wave of Adoption (Dry Cropcuts}
+    \label{fig:when_adopt}
+\end{figure}
 

### 2021-12-14 paper.tex

@@
 The latest survey round is not a a continuation of the panel of the first three rounds, but has novel data on the genetic finger prints of the seeds used, that we exploit to understand why hybrid seed varieties seem not to increase returns when adopted.
+
+Varying how strict of a definition for hybrid seeds we use, starting from second-generation and recycled (SR1), and a more strict definition, which includes only new varieties (SR2), we find that  farmers that adopted a seeds with a less pure genetic profile
+
+self-reported improved seed would’ve been better off not adopting, and farmers
+that did not adopt a self-reported improved seed would’ve been better off adopting.
+16
 
@@
 Equation \ref{eq:switch_latent} shows the definition for a latent variable $A^*$ that reflects the benefit from adopting an improved seed relative to planting a traditional one.
-As shown in the equation, household $i$ will decide to adopt  (A_i=1).
+As shown in the equation, household $i$ will decide to adopt ($A_i$=1)if these perceived benefits of adoption are positive.
 
@@
 \end{align}
+Vector $Z$ includes variables that determine if a particular household may benefit from the adoption of improved seeds.
 
@@
 \resizebox{1\textwidth}{!}{
-\input{results/tables/grc_YIELD_cropcutdry_tr}
+\input{results/tables/grc}
 }

### 2021-12-14 paper.tex

--- 2021-12-14 paper.tex
@@
 The latest survey round is not a a continuation of the panel of the first three rounds, but has novel data on the genetic finger prints of the seeds used, that we exploit to understand why hybrid seed varieties seem not to increase returns when adopted.
+Varying how strict of a definition for hybrid seeds we use, starting from second-generation and recycled (SR1), and a more strict definition, which includes only new varieties (SR2), we find that  farmers that adopted a seeds with a less pure genetic profile would have had better returns by simply not adopting.
 
-%% What we find
-
-Our results show evidence that showcase that the average yields without adoption for
-
-return to adoption
-% Evidence of heterogeneity in response to hybrid adoption for joiners and leavers, however their average yield without adoption ($\mu_{01}$,$\mu_{10}$) are very similar.\\
-
-% are also driven by climate and how this can affect dis-adoption decisions. We extend the role of heterogeneity in a household’s
-% decision to adopt new technologies in four fronts.
-
-% We describe what we do here and what we find. We can expand as we have results
-1) First, we test if the adoption of improved seed varieties takes place where the net benefits of adoption is highest?
-
-2) Second, we test how the adoption decision correlates with other adaptation strategies such as water storage, and irrigation augment a household’s comparative advantage in adoption.
-
-3) Third.
-what are the drivers of dis-adoption of improved maize seed?
-and (4) is it possible to extend the methodology in \citep{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} by including time-varying characteristics to the Group Random Coefficients (GRC) model?
-
-% Persistent lack of adoption is a reflection of the distribution of (observable and unobservable) costs and benefits of the technology.
-The approach here models households’ adoption decisions in an environment with household-specific heterogeneity in the costs and benefits, and hence profits, to the technology.
-I estimate how the returns to the technology vary across farmers and then compare these returns to the adoption decisions of farmers.
-I find that farmers with low (or zero) returns to the technology are precisely the farmers who do not adopt it.
-In particular, I use a generalized
+We put together the results from both estimation strategies to suggest that the negative return to adoption we find under the Group Random Coefficients estimation reflect that relying on a loose definition of hybrid seeds
 
@@
 \end{align}
+Vector $Z$ includes variables that determine if a particular household may benefit from the adoption of improved seeds.
+The definition of what an improved seed varies in the estimations presented in the next sections and consider not only what
 

### 2021-12-14 paper.tex
Updates from Overleaf

@@
 Since the drought of 1984, Ethiopia has become increasingly reliant on maize, as of today, it constitutes the most widespread crop in the country.\footnote{FAO stats data shows the total area cultivated with Maize in 2019 is 25\% larger than the one cultivated with the next most widespread crop (sorghum), and 27\% and 139\% larger than the next two following crops, wheat and Barley, respectively.} As Figure \ref{fig:maize_yields} shows, there is an upward trend in the total area cultivated with maize since 1990, with average yields that have almost doubled during this period.
+These increases in yields have had enormous consequences in decreasing undernourishment, as its prevalence has dramatically decreased from from 47\% of the population in 2001 to 14\% in 2018 (The World Bank, 2021).
+Despite such generalized progress, food insecurity remains affecting more than half the population of Ethiopia (The World Bank, 2021), which already speaks about the heterogeneous progress across subpopulations\footnote{The country has also recently seen increases in food insecurity because of COVID-19, the conflict in Ethiopia’s Tigray region and the desert locust outbreak}.
+
+The large disparities in yields and food security between the African continent and the rest of the developing world are often attributed to different adoption rates of green revolution technologies, such as improved seeds.
+Evidence has shown that the adoption of improved seeds leads to increases in agricultural yields (\citep{Carter2014-fm}), food security (\citep{Shiferaw2014-op}), and poverty reduction (\citep{Minten2008-tj}).
+Although critics of green revolution technologies have also claimed that these technological packages have not been as profitable for farmers in Africa as promised, given the high costs of improved seeds compared to the gains in productivity
+
+Evidence has shown that the adoption of improved seed leads to increases in agricultural yields \citep{Carter2014-fm}, food security \citep{Shiferaw2014-op} and poverty reduction \citep{Minten2008-tj}.
+Despite such positive impacts, the low adoption levels of improved seeds varieties in Sub-Saharan Africa remains a puzzle.
 
@@
 Moreover, there is a large percentage of farmers who adopt new varieties inconsistently limiting their opportunities to learn about their benefits, see Table \ref{tbl:Trajectories}.
+Evidence has shown that the adoption of improved seed leads to increases in agricultural yields \citep{Carter2014-fm}, food security \citep{Shiferaw2014-op} and poverty reduction \citep{Minten2008-tj}.
+Despite such positive impacts, the low adoption levels of improved seeds varieties in Sub-Saharan Africa remains a puzzle.
 
@@
 This issue is becoming more important as even current varieties of improved seed are over 20 years old and their effectiveness in terms of yield improvement and disease resistance is waning \citep{Abate2015-rj}.
+
+Multiple explanations have been proposed for observed adoption rates, including the type of technology promotion mechanisms used by governments, NGO’s, and businesses, the characteristics of the improved seeds and their adaptation to specific agro-climatic environments \citep{Bird2020-nt}, as well as growing conditions, such as the type of soil, input use, and other farmer characteristics \citep{Munshi2004-og}.
+The consensus in this literature is that the success of improved varieties depends on a range of factors that need to be better understood.
+Success of improved seeds should be accompanied by complementary conditions and inputs that would allow the full potential of these varieties to be realized.
+Specifically, extension services and on-farm field trials, seed variety characteristics and rainfall have been found to be crucial to the adoption of improved maize in Tanzania \citep{Kaliba2000-jh}.
+In the specific case of Ethiopia, labor, fertilizer use, farmers’ experience with extension packages, rainfall suitability and prices have been shown to increase the opportunity cost of not adopting improved wheat varieties \citep{Wale2006-bv}.
+Market access and the accessibility of extension services are also critical in the adoption of improved chickpea varieties in Ethiopia \citep{Verkaart2019-ol}.
+Other important determinants of heterogeneous effects include risk aversion \citep{Holden2016-vy}, farm size \citep{Ghimire2015-bd}, and credit constraints \citep{Simtowe2008-jn,Balana2020-hx}.
+
+A different branch of the literature provides an alternative explanation for such low adoption rates, focusing on heterogeneous potential returns to adoption \citep{Suri2011-oi}.
+This literature suggests that even if returns to adoption are high, farmers may fail to adopt if they face low comparative advantage to adoption.
+Other methods used in the literature fail to account for this heterogeneity, admitting only the presence of time-constant and/or time-varying unobserved heterogeneities (depending on the data structure) for identification purposes \citep{Kassie2018-xn,Falco2011-rt}.
+These methods ignore the possibility of a farmer’s comparative advantage of using an improved technology over a traditional one as playing a central role in their adoption decision.
+
+We rely on Suri’s method and its extension in \citep{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} to account for such heterogeneous comparative advantages and evaluate how these are explained by climatic conditions.
+In Ethiopia, rainfall deviations from historical averages are found to be the main driver of heterogeneous effects for adopting agronomic packages \citep{Marenya2020-kb}.
+Similar experiences from Malawi show exposure to past droughts is found to increase the adoption of drought-tolerant maize among farmers that also receive subsidies \citep{Katengeza2019-af} and the dis-adoption of traditional local maize \citep{Holden2016-vy}.
 
@@
 
-We put together the results from both estimation strategies to suggest that the negative return to adoption we find under the Group Random Coefficients estimation reflect that relying on a loose definition of hybrid seeds
+We put together the results from both estimation strategies to suggest that the negative return to adoption we find under the Group Random Coefficients estimation reflects the loose definition of hybrid seeds, where only a subset of farmers are using the purest genetic varieties.
+Considering that poorer farmers are the ones paying a premium for genetically diluted hybrids, this result has several policy implication about the need to foster a more inclusive policy when helping farmers to obtain the best performing seeds.
 
@@
     \centering
-    \includegraphics{results/figures/Maize_yields.pdf}
-    \caption{Maize: Total area harvested and Average yields}
+    \caption{Maize: Total area harvested and Average yields}    \includegraphics{results/figures/Maize_yields.pdf}
+
     \figcaption{Source: Authors' calculations using FAO stats data
@@
 \end{figure}
-
-\section{Context}
-
-Evidence has shown that the adoption of improved seed leads to increases in agricultural yields \citep{Carter2014-fm}, food security \citep{Shiferaw2014-op} and poverty reduction \citep{Minten2008-tj}.
-Despite such positive impacts, the low adoption levels of improved seeds varieties in Sub-Saharan Africa remains a puzzle.
-
-Multiple explanations have been proposed for observed adoption rates, including the type of technology promotion mechanisms used by governments, NGO’s, and businesses, the characteristics of the improved seeds and their adaptation to specific agro-climatic environments \citep{Bird2020-nt}, as well as growing conditions, such as the type of soil, input use, and other farmer characteristics \citep{Munshi2004-og}.
-The consensus in this literature is that the success of improved varieties depends on a range of factors that need to be better understood.
-Success of improved seeds should be accompanied by complementary conditions and inputs that would allow the full potential of these varieties to be realized.
-Specifically, extension services and on-farm field trials, seed variety characteristics and rainfall have been found to be crucial to the adoption of improved maize in Tanzania \citep{Kaliba2000-jh}.
-In the specific case of Ethiopia, labor, fertilizer use, farmers’ experience with extension packages, rainfall suitability and prices have been shown to increase the opportunity cost of not adopting improved wheat varieties \citep{Wale2006-bv}.
-Market access and the accessibility of extension services are also critical in the adoption of improved chickpea varieties in Ethiopia \citep{Verkaart2019-ol}.
-Other important determinants of heterogeneous effects include risk aversion \citep{Holden2016-vy}, farm size \citep{Ghimire2015-bd}, and credit constraints \citep{Simtowe2008-jn,Balana2020-hx}.
-
-A different branch of the literature provides an alternative explanation for such low adoption rates, focusing on heterogeneous potential returns to adoption \citep{Suri2011-oi}.
-This literature suggests that even if returns to adoption are high, farmers may fail to adopt if they face low comparative advantage to adoption.
-Other methods used in the literature fail to account for this heterogeneity, admitting only the presence of time-constant and/or time-varying unobserved heterogeneities (depending on the data structure) for identification purposes \citep{Kassie2018-xn,Falco2011-rt}.
-These methods ignore the possibility of a farmer’s comparative advantage of using an improved technology over a traditional one as playing a central role in their adoption decision.
-
-We rely on Suri’s method and its extension in \citep{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} to account for such heterogeneous comparative advantages and evaluate how these are explained by climatic conditions.
-In Ethiopia, rainfall deviations from historical averages are found to be the main driver of heterogeneous effects for adopting agronomic packages \citep{Marenya2020-kb}.
-Similar experiences from Malawi show exposure to past droughts is found to increase the adoption of drought-tolerant maize among farmers that also receive subsidies \citep{Katengeza2019-af} and the dis-adoption of traditional local maize \citep{Holden2016-vy}.
-
-We will contribute to this literature by considering historical climate averages on the suitability of improved varieties to understand whether adoption is indeed realized where its potential is highest.
-Moreover, we will contribute to the understanding of how water conservation innovations and irrigation mitigate the effects of droughts on the potential of improved seeds.
 
@@
     \centering
+    \caption{Spatial Distribution of Households Surveyed}
     \includegraphics{results/figures/map_hhids.pdf}
-    \caption{Spatial Distribution of Households Surveyed}
     \figcaption{Note: Green markers denote the locations of households}
@@
     \centering
+    \caption{Adoption rates of improved varieties in 2018 (by self-report, purity level, crop type, source and year of release)}
     \includegraphics{results/figures/adoption_r4.png}
-    \caption{Adoption rates of improved varieties in 2018 (by self-report, purity level, crop type, source and year of release)}
     \figcaption{Note: Shares represent the proportion of adopters among maize farmers.
@@
     \centering
-    \includegraphics{results/figures/missclassification.png}
     \caption{Correlation of different definitions of improved seeds, 4th round}
+    \includegraphics[width = 0.9\textwidth]{results/figures/missclassification.png}
     \figcaption{Note: Shares represent the proportion of adopters among maize farmers.
@@
 
-where $y$ is log yields, $\mu_{\underline{h}}$ are binary variables for being in trajectory $\underline{h}$, $h$ is a binary variable for whether household $i$ adopted in wave $t$, $X$ are a set of controls and $\Delta_{0,0,1}$ are the returns to adoption for trajectory $(0,0,1)$.
+where $y$ is log yields, $\mu_{\underline{h}}$ are binary variables for being in trajectory $\underline{h}$, $h$ is a binary variable for whether household $i$ adopted in wave $t$, and $X$ are a set of controls.
 $\kappa_{(1,1,1)}$ is the average yield with adoption of the always-adopters, where $\kappa_{(1,1,1)} = \mu_{(1,1,1)} + \Delta_{(1,1,1)}$.
@@
 
-where $\phi$ is the parameter which describes sorting in the economy.
+where $\phi$ is the parameter which describes sorting in the economy and $\Delta_{0,0,1}$ are the returns to adoption for trajectory $(0,0,1)$.
 If $\phi$ is positive, then higher yields is associated with higher comparative advantage, denoting that those that adopt also tend to have the highest returns to adoption.
@@
 The advantage of the weak-identification robust procedure is that we can test for whether the data satisfies the conditions necessary for an interpretable $\phi$ parameter.
+If there are trajectories which yield a $\phi$ with a non-infinite confidence interval, then the LCA is at least partially satisfied.
 
@@
 \par
-The decision to adopt and improved seed is determined by a series of observed and unobserved variables.
+This approach models the effect of improved seed adoption on yields in two stages.
+First, the decision to adopt an improved seed is determined by a series of observed and unobserved variables.
 Equation \ref{eq:switch_latent} shows the definition for a latent variable $A^*$ that reflects the benefit from adopting an improved seed relative to planting a traditional one.
@@
 \end{align}
-Vector $Z$ includes variables that determine if a particular household may benefit from the adoption of improved seeds.
-The definition of what an improved seed varies in the estimations presented in the next sections and consider not only what
+Vector $Z$ includes variables that determine if a particular household may benefit from the adoption of improved seeds, such as farmers' demographic characteristics (education, age, female headed households) and land tenure status of their plots.
+In the estimations presented in the next sections, we consider different definitions of improved seeds, and include not only self-reported seed classification but also the results of DNA fingerprinting.
+\par
+The second stage presents the outcome functions conditional on adoption:
+\begin{align}
+
+\end{align}
 
@@
 Commensurate with the literature, we also include interactions of these controls with hybrid adoption as added controls.
+Interactions with the hybrid variable are meant to account further account for household specific variation that adopter may have.
 The results of the regression are shown in Table \ref{tbl:unres}.
+
+The results for self-reported and cropcut yields are largely similar, pointing to the fact that self-reported yields may not have a large amount of measurement error.
+In the base case, there are some positive effects to adoption, but with all controls included, marginal effects of adoption for each trajectory are negative and significant.
+This translates to average yields of adoption that close to 0 or even negative.
 
@@
     \centering
-    \includegraphics[scale=0.75]{results/figures/theta.png}
     \caption{Comparative Advantage and Returns by Trajectory (Dry Cropcuts)}
-    \label{fig:theta_delta_raw}
+    \includegraphics[scale=0.75]{results/figures/theta.png}\label{fig:theta_delta_raw}
 \end{figure}

### 2022-01-27 results/tables/switch23-sr.tex
Updates from Overleaf

\begin{table}[H]
\tiny
\centering
\hspace*{-1.2cm}
\begin{threeparttable}
\caption{Effects on yields (log) from adopting improved maize varieties (DNA fingerprinting)}
\label{tab:switch2}
\begin{tabular}{l cccccc}
\hline
\hline
            &Adopters yields&Non-adopters yields&         ATE&          SE&     p-value\\
\hline
\textit{DNA 70, SR yields}&            &            &            &            &            \\
ATT         &        7.17&        7.31&       -0.14&       0.026&      0.0000\\
%
%
%
ATU         &        8.76&        6.76&        2.00&       0.036&      0.0000\\
%
%
%
\textit{DNA 90, SR yields}&            &            &            &            &            \\
ATT         &        7.18&        9.07&       -1.90&       0.023&      0.0000\\
%
%
%
ATU         &        7.01&        6.80&        0.21&       0.022&      0.0000\\
%
%
%
\textit{DNA 95, SR yields}&            &            &            &            &            \\
ATT         &        7.30&        7.08&        0.22&       0.036&      0.0000\\
%
%
%
ATU         &        6.87&        6.86&       0.018&       0.029&      0.5496\\
%
%
%
\textit{HYB, SR yields}&            &            &            &            &            \\
ATT         &        7.28&        7.25&       0.030&       0.028&      0.2892\\
%
%
%
ATU         &        7.06&        6.85&        0.21&       0.028&      0.0000\\
%
%
%
\textit{DTMZ, SR yields}&            &            &            &            &            \\
ATT         &        7.42&        6.53&        0.89&       0.074&      0.0000\\
%
%
%
ATU         &        7.11&        7.02&       0.089&       0.022&      0.0001\\
%
%
%
\textit{CGIAR source, SR yields}&            &            &            &            &            \\
ATT         &        7.13&        6.71&        0.43&       0.033&      0.0000\\
%
%
%
ATU         &        9.19&        6.89&        2.31&       0.018&      0.0000\\
%
%
%
\textit{Year 2010+, SR yields}&            &            &            &            &            \\
ATT         &        7.44&        6.92&        0.53&       0.039&      0.0000\\
%
%
%
ATU         &        8.98&        6.91&        2.07&       0.056&      0.0000\\
%
%
%
\hline
\hline
\end{tabular}
\begin{tablenotes}
\tiny
\item{Note: Full set of controls include: parcel size (in HA), household labor, hired labor, fertilizer costs, other input costs irrigation (dummy), mechanization (dummy), organic fertilizer (dummy). Instruments for the adoption equation include: years of education of hh head, age of hh head, female head, land title, asset index, seed costs. All results include full set of controls. DNA 70, 90 and 95 refer to improved varieties defined by DNA finerprinting with purity levels of 70, 90 and 95\%, respectively. HYB equals 1 for a hybrid variety, 0 for an open-pollinated variety. DTMZ equals 1 for a Drought-tolerant maize variety}
\end{tablenotes}
\end{threeparttable}
\end{table}

### 2022-01-27 document/slides/finalpresentation.tex

--- 2021-11-24 document/slides/slides.tex
@@
 %Title page
-\title[Improved Seed Adoption] %optional/
-{Improved Seed Adoption and Climate Change: Hybrid Maize in Ethiopia}
+\title[Genetic Dilution in Improved Maize] %optional/
+{A Case of Genetic Dilution in Improved Maize: Estimating
+Comparative Advantage of Adoption in Ethiopia}
 
-\author[et al]{Barriga-Cabanillas,  Chiarella, Correa, Michuda}
+\author[et al]{Michuda, Chiarella, Barriga-Cabanillas, Correa, }
 
@@
 
-\date[Feb 2021] % (optional)
+\date[January 2022] % (optional)
 {CGIAR-SPIA \\
-February 2021}
+January, 2022}
 
@@
 %---------------------------------------------------------
+%---------------------------------------------------------
+\begin{frame}
+\frametitle{Group Random Coefficients Results}
+
+\end{frame}
+
+%---------------------------------------------------------
+
+%---------------------------------------------------------
+\begin{frame}
+\frametitle{Endogenous Switching Regression Results}
+
+\input{results/tables/switch23-sr}
+
+\end{frame}
+%---------------------------------------------------------
+
+%---------------------------------------------------------
+\begin{frame}
+\frametitle{Discussion}
+
+\begin{itemize}
+    \item Negative returns to adoption from self-reported improved varieties.
+    \item But positive returns to more genetically pure, DTMZ, CG- sourced and newer varieties.
+    \item All DNA fingerprints are 70\% pure or higher. Control group has high share of false positives: no proper control. 
+    \item Negative results from GRC mask a large heterogeneity in returns to adoption.
+\end{itemize}
+
+\end{frame}
+
+%---------------------------------------------------------
 

### 2022-01-27 paper.tex

@@
-\documentclass{article}
-
-\title{Estimating Comparative Advantage in Improved Seed Adoption: Hybrid Maize in Ethiopia}
+\documentclass[11pt]{article}
 
 \usepackage{natbib}
-\bibliographystyle{rusnat}
+\bibliographystyle{abbrvnat}
+\setcitestyle{authoryear,open={(},close={)}} 
 \usepackage{multicol}
@@
 \usepackage{float}
-
+\usepackage{bm}
 \usepackage{graphics}
-
+\usepackage{threeparttable}
+\usepackage{amssymb}
 \usepackage[margin=1in]{geometry}
 \usepackage{setspace}
+\usepackage{afterpage}
 \doublespacing
+\usepackage{subfig}
+\usepackage{epigraph}
+
+\usepackage{enumerate}
+\usepackage{tabularx}
+\usepackage{fancyhdr}
+\usepackage{lscape}
+\usepackage{mdframed}
+\usepackage{mathtools}
+\usepackage{array}
+\usepackage{import}
+\usepackage{fullpage}
 
 \begin{document}
+    \begin{titlepage}
+    \title{A Case of Genetic Dilution in Improved Maize: Estimating Comparative Advantage of Adoption in Ethiopia}
+% This second title is my preferred title at this point (28 Oct 2021). Open to other titles, but it shouldn't be longer than this one!
 
+%     \author[1]{Oscar Barriga-Cabanillas}  
+%     \author[2]{Cristina Chiarella}    
+%     \author[2]{Juan Sebastian Correa}  	
+%     \author[3]{Aleksandr Michuda}
+
+% 	\affil[1]{\small \emph{The World Bank}}
+% 	\affil[2]{\small \emph{FAO}}
+% 	\affil[3]{\small \emph{Assistant Research Professor at the Center for Data Science for Enterprise and Society, Cornell University}}
+\author{%
+ Aleksandr Michuda \footnote{Center for Data Science for Enterprise and Society, Cornell University}%
+ \and Cristina Chiarella\footnote{UC Louvain}%
+ \and Oscar Barriga-Cabanillas\footnote{The World Bank}%
+  \and Juan Sebastian Correa \footnote{FAO}%
+  }
+%   Assistant Research Professor at the 
+	\date{\today}
+
+    \clearpage\maketitle
+	\thispagestyle{empty}
+	\vspace*{-2em}
+    \begin{center}\begin{abstract}
+			\noindent  
+			Despite large investments in breeding and making high-yielding maize varieties available, evidence shows that few farmers consistently use them over time.
+We use an innovative methodology that studies the adoption trajectories of farmers over a three-round panel data set to explain low adoption rates with heterogeneous returns to adoption.
+By relying on self-reported use of improved seeds, we find that farmers do not adopt because they have a low comparative advantage and seemingly small to negative returns to adoption.
+To reconcile this result, we exploit unique information on maize DNA-fingerprinting collected over the same areas in 2018.
+We find that the negative returns to adoption mask heterogeneity driven by the variety of improved seeds used.
+Positive returns to adoption are found for those farmers that use higher-purity germplasm, drought-tolerant maize, and newly released varieties.
+Our findings point to the wide dispersion of older and genetically diluted varieties, for which poorer farmers are paying a premium.
+The implications of our findings speak to the need for policies to better target context and geography, expand accessibility of improved seeds, and make higher yields more inclusive.
+
+			%Despite the large investments on creating better performing maize varieties, very few farmers consistently use them.
+We rely on four rounds of the ESS survey to answer this question.
+We find that relying on self reported use of hybrid seeds suggest there are negative returns to adoption.
+To reconcile this result, we use the information on genetic fingerprinting of seed available on the last ESS survey round to show that positive return for adoption appear only to exits for those farmers using seed with pure germoplast.
+We discuss the implications of our findings, showing that poorer farmers are the ones that consistently rely on second and third generation seeds, whose performance is debatable.
+        \end{abstract}\end{center}
+
+        {\small \noindent\emph{JEL Classification}:
+        	\emph{Keywords}: Improved maize, DNA-fingerprinting, Group Random Coefficient, Endogenous Switching Regression, Heterogeneous returns}
+
+    \end{titlepage}
 \maketitle
+\newpage
 
-%%%% I added the bullet points where I think they fit what we have as an introduction
-
-\begin{itemize}
-    \item Tim Weiss- hybrid subsidies
-    \item is there such a thing as traditional seeds?
-
-    \item Let's evaluate the effect of adoption in the context of Ethiopia
-    \item find ambiguous results, Why?
-    \item have access to DNA fingerprinting data
-    \item what do they say about adoption?
-    \item turns out that "adoption" isn't really a reliable concept/variable
-    \item most farmers are "adopting" under pretty conservative purity definitions
-    \item Let's see what happens with different varieties of seed, does adoption of that seed lead to positive yield effects?
-    \item can't use these methods but newer data has access to newer varieties where we're relatively more confident
-\end{itemize}
+%\begin{itemize}
+%    \item Tim Weiss- hybrid subsidies
+%    \item is there such a thing as traditional seeds?
+%    \item Let's evaluate the effect of adoption in the context of Ethiopia
+%    \item find ambiguous results, Why?
+%    \item have access to DNA fingerprinting data
+%    \item what do they say about adoption?
+%    \item turns out that "adoption" isn't really a reliable concept/variable
+%    \item most farmers are "adopting" under pretty conservative purity definitions
+%    \item Let's see what happens with different varieties of seed, does adoption of that seed lead to positive yield effects?
+%    \item can't use these methods but newer data has access to newer varieties where we're relatively more confident
+%\end{itemize}
 
@@
 Food security is a major challenge in many East African countries, and Ethiopia is no exception, having historically struggled to provide an adequate and reliable food supply \citep{Ramakrishna2002-hv, Jaleta2018-oj}.
-Since the drought of 1984, Ethiopia has become increasingly reliant on maize, as of today, it constitutes the most widespread crop in the country.\footnote{FAO stats data shows the total area cultivated with Maize in 2019 is 25\% larger than the one cultivated with the next most widespread crop (sorghum), and 27\% and 139\% larger than the next two following crops, wheat and Barley, respectively.} As Figure \ref{fig:maize_yields} shows, there is an upward trend in the total area cultivated with maize since 1990, with average yields that have almost doubled during this period.
-These increases in yields have had enormous consequences in decreasing undernourishment, as its prevalence has dramatically decreased from from 47\% of the population in 2001 to 14\% in 2018 (The World Bank, 2021).
-Despite such generalized progress, food insecurity remains affecting more than half the population of Ethiopia (The World Bank, 2021), which already speaks about the heterogeneous progress across subpopulations\footnote{The country has also recently seen increases in food insecurity because of COVID-19, the conflict in Ethiopia’s Tigray region and the desert locust outbreak}.
+Since the drought of 1984, Ethiopia has become increasingly reliant on maize, and as of today, maize constitutes the most widespread crop in the country.\footnote{FAO statistics show the total area cultivated with Maize in 2019 is 25\% larger than the next most widespread crop (sorghum), and 27\% and 139\% larger than the next two following crops, wheat and barley, respectively.} As Figure \ref{fig:maize_yields} shows, there is an upward trend in the total area cultivated with maize since 1990, with average yields almost doubling during this period.
+This increase in maize yields has also coincided with a dramatic decrease in undernourishment which has fallen from 47\% of the population in 2001 to 14\% in 2018 (The World Bank, 2021).
+Despite these positive gains, food insecurity still affects more than half the population of Ethiopia (ibid), reflecting that the progress achieved
+is not necessarily homogeneous across all subpopulations.\footnote{The country also has recently seen increases in food insecurity because of COVID-19, the conflict in Ethiopia’s Tigray region and the desert locust outbreak.}
 
-The large disparities in yields and food security between the African continent and the rest of the developing world are often attributed to different adoption rates of green revolution technologies, such as improved seeds.
-Evidence has shown that the adoption of improved seeds leads to increases in agricultural yields (\citep{Carter2014-fm}), food security (\citep{Shiferaw2014-op}), and poverty reduction (\citep{Minten2008-tj}).
-Although critics of green revolution technologies have also claimed that these technological packages have not been as profitable for farmers in Africa as promised, given the high costs of improved seeds compared to the gains in productivity
-
-Evidence has shown that the adoption of improved seed leads to increases in agricultural yields \citep{Carter2014-fm}, food security \citep{Shiferaw2014-op} and poverty reduction \citep{Minten2008-tj}.
-Despite such positive impacts, the low adoption levels of improved seeds varieties in Sub-Saharan Africa remains a puzzle.
+\begin{figure}[htp]
+    \centering
+    \caption{Maize: Total area harvested and average yields} \label{fig:maize_yields}
+    \includegraphics[width=.6\textwidth]{results/figures/Maize_yields.pdf}
+    \vspace*{-1em}
+    \begin{table}[H]
+        \centering
+        \begin{tabular}{p{0.6\textwidth}} 
+            \begin{tablenotes}[flushleft]
+                  \small
+                  \item Source: Authors' calculations using FAOSTAT data 
+                  \item Note: Yields are measured as the harvested production per ha for the area under cultivation
+            \end{tablenotes}
+        \end{tabular}
+    \end{table}       
+\end{figure}
 
 %%%%% The puzzle 
-Despite numerous technological breakthroughs in maize germplasm with increased yields and drought tolerance, adoption of improved seeds has remained a challenge, with only 13\% of maize farmers consistently using improved seeds varieties, and the lion share of farmers (62\%) not using improved seeds at all during the period where data is available.
-Moreover, there is a large percentage of farmers who adopt new varieties inconsistently limiting their opportunities to learn about their benefits, see Table \ref{tbl:Trajectories}.
-Evidence has shown that the adoption of improved seed leads to increases in agricultural yields \citep{Carter2014-fm}, food security \citep{Shiferaw2014-op} and poverty reduction \citep{Minten2008-tj}.
-Despite such positive impacts, the low adoption levels of improved seeds varieties in Sub-Saharan Africa remains a puzzle.
+The large disparities in yields and food security between the African continent and the rest of the developing world are often attributed to different adoption rates of green revolution technologies, such as improved seeds.
+Evidence has shown that the adoption of improved seed leads to increases in agricultural yields \citep{Carter2014-fm}, food security \citep{Shiferaw2014-op}, and poverty reduction \citep{Minten2008-tj}.
+Nevertheless, critics of green revolution technologies have also claimed that these technological packages have not been as profitable for many farmers in Africa as promised (with the exception of Ethiopia), given the high costs, and their heavy dependence on input subsidies \citep{Wise20}.
+\par
+Even if the net returns to adoption of improved seeds are more conservative than often thought, the evidence regarding increased yields over traditional varieties is consistent across multiple contexts.
+Despite this evidence and numerous technological breakthroughs in maize germplasm with increased yields and drought tolerance, in the period between 2011 and 2015 only 13\% of maize farmers in Ethiopia have adopted improved seeds consistently, a large percentage has adopted new varieties inconsistently, and the lion share of farmers (62\%) have not used improved seeds at all during the same period.
 
-The facts sub-Saharan farmers continue to use traditional farming techniques when more modern, higher-return agricultural technologies are available remains an empirical puzzle where several answers have been proposed.
-These include imperfections in credit markets \citep{Croppenstedt2003-pq}, property rights \citep{Place2000-el}, social learning \citep{Conley2010-ue,Foster1995-bz,Munshi2004-og}, lack of commitment devices \cite{Duflo2009-iv}, and high transportation costs that increase the cost of agricultural inputs \citep{Byerlee2013-qk}.
-This issue is becoming more important as even current varieties of improved seed are over 20 years old and their effectiveness in terms of yield improvement and disease resistance is waning \citep{Abate2015-rj}.
+Several answers have been proposed to explain such low adoption of improved varieties.
+These include mainly the lack of liquidity and other economic constraints \citep{Carter2014-fm}, growing conditions such as the type of soil, input use, and farmer characteristics \citep{Munshi2004-og}, imperfections in credit markets \citep{Croppenstedt2003-pq}, ambiguously defined property rights \citep{Place2000-el}, lack of commitment devices \cite{Duflo2009-iv}, high transportation costs that increase the cost of agricultural inputs \citep{Byerlee2013-qk}, and improved seed supply constraints \citep{Bird2020-nt}.
+In this context, there is still a large percentage of the population not using improved seed varieties, or using varieties that are over 20 years old.
+With such old seed, there is evidence that yield improvement and disease resistance in these seeds are waning\citep{Abate2015-rj}.
+The variety of these constraints and their dispersion across subpopulations and locations speaks to considerable heterogeneous returns to adoption.
 
-Multiple explanations have been proposed for observed adoption rates, including the type of technology promotion mechanisms used by governments, NGO’s, and businesses, the characteristics of the improved seeds and their adaptation to specific agro-climatic environments \citep{Bird2020-nt}, as well as growing conditions, such as the type of soil, input use, and other farmer characteristics \citep{Munshi2004-og}.
-The consensus in this literature is that the success of improved varieties depends on a range of factors that need to be better understood.
-Success of improved seeds should be accompanied by complementary conditions and inputs that would allow the full potential of these varieties to be realized.
+%%%% Let's evaluate the effect of adoption in the context of Ethiopia
+
+In this paper, we revisit the mechanism proposed by \citep{Suri2011-oi}, where a lack of adoption of improved seed varieties is the result of farmers' heterogeneous returns to the new technology.
+Ignoring such heterogeneity can hide the true impact of adoption, as different sub-populations' returns may be impacted by the same technology in different, and even opposite ways, despite it providing high average returns.
+The heterogeneous returns to adoption are modeled as a farmer's unobservable comparative advantage using a Correlated Random Coefficient (CRC) model.
+In the original paper, \citep{Suri2011-oi} finds that farmers’ lack of adoption is driven by insufficient net benefits to the technology stemming from poor infrastructure that increase the cost of access.
+Similar work in Ethiopia on improved chickpea varieties suggests that the economic returns to adoption plays a major role in the decision to adopt \citep{Michler2018-wk}.
+
+%% Expand a bit on the methods
+We build on this literature by investigating hybrid maize seed adoption in Ethiopia, focusing on how heterogeneous returns drive the decision to use improved seed varieties.
+We implement two complementary estimation strategies.
+First, we implement a Group Random Coefficients estimation, which builds on \cite{Suri2011-oi} in \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} using a General Method of Moments (GMM) Estimator that allows us to test the assumptions assumed by \cite{Suri2011-oi}, as well as provide more economically interpretable results.
+The advantage of our estimation strategy is that we do not rely on identification from a valid instrument, but a strategy similar to \cite{Chamberlain1984-uk}.
+For this strategy, we rely on the first three survey rounds of the Ethiopian Socioeconomic Survey (ESS), panel dataset that allows us to follow the same households across time (from 2011 to 2015).
+Using self-reported information on the seed variety (improved or traditional), we compare the adoption trajectories over three rounds.
+The results show that the returns to adoption are in most cases, not statistically significant, and only significant with wave 3 adoption.
+This posits a puzzle that motivates our next estimation strategy.
+
+We expand our results using the fourth round of the ESS (conducted in 2018), which provides a unique opportunity to understand why improved seed varieties seem not to increase returns when adopted.
+The ESS4 collected DNA fingerprints of harvested cropcuts in selected plots and compared its genetic material with a reference library of available varieties in Ethiopia.
+This provides critical information not only about the seed types but also about their purity, sources, and their year of release.
+We exploit this uniquely rich information to compare different definitions of improved seeds.
+Specifically, we compare the returns to the self-reported use of new, second generation or recycled varieties, with those of improved seeds as defined by their genetic material, comparing different purity thresholds (or degrees of no-contamination with other varieties), year of release, types of varieties, and their sources.
+Because the latest survey round is not a continuation of the first three rounds, our second estimation strategy implements an Endogenous Switching Regression (ESR) approach, a standard estimation strategy to study the question of adoption with cross-sectional data (\citealt{Marenya2020-kb}, \citealt{Shiferaw2014-op}), given its capacity to account for unobserved heterogeneity in adoption.
+
+The results of the ESR find negative effects of adopting improved seeds, as per farmer self-reports.
+But the comparison of the different improved seed definitions using the DNA fingerprinting sheds light on possible explanations for this apparent puzzle.
+We find that when the definition of improved seed is based on greater purity of the germplasm (95\% of purity or more), the returns to adoption for those that used such improved varieties becomes positive and significant.
+We obtain similar results when evaluating the specific effects of adopting drought-tolerant maize varieties, CGIAR or Exotic varieties, or more recently released varieties.
+
+Taken together, the negative and not statistically significant returns to adoption we find from both estimation strategies reflect the loose definition of improved seeds when such improved varieties are self-reported by farmers.
+Only a subset of farmers use the purest genetic varieties, drought-tolerant varieties, or newly released varieties, which tend to provide higher returns.
+This implies that poorer farmers are paying a premium for genetically diluted hybrids, which has far-reaching policy implications about the need to foster a more inclusive policy when making the best performing seeds available to all farmers.
+
+Our results contribute to the literature on the drivers of low adoption rates.
+The consensus in this literature is that the success of improved varieties depends on a range of factors, and that the widespread adoption of  improved seeds is the result of complementary conditions and inputs that allow farmers to extract the full potential of these varieties.
 Specifically, extension services and on-farm field trials, seed variety characteristics and rainfall have been found to be crucial to the adoption of improved maize in Tanzania \citep{Kaliba2000-jh}.
@@
 
-A different branch of the literature provides an alternative explanation for such low adoption rates, focusing on heterogeneous potential returns to adoption \citep{Suri2011-oi}.
+% A different branch of the literature provides an alternative explanation for the low adoption rates, focusing on heterogeneous potential returns to adoption \citep{Suri2011-oi}.
 This literature suggests that even if returns to adoption are high, farmers may fail to adopt if they face low comparative advantage to adoption.
 Other methods used in the literature fail to account for this heterogeneity, admitting only the presence of time-constant and/or time-varying unobserved heterogeneities (depending on the data structure) for identification purposes \citep{Kassie2018-xn,Falco2011-rt}.
-These methods ignore the possibility of a farmer’s comparative advantage of using an improved technology over a traditional one as playing a central role in their adoption decision.
 
-We rely on Suri’s method and its extension in \citep{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} to account for such heterogeneous comparative advantages and evaluate how these are explained by climatic conditions.
-In Ethiopia, rainfall deviations from historical averages are found to be the main driver of heterogeneous effects for adopting agronomic packages \citep{Marenya2020-kb}.
-Similar experiences from Malawi show exposure to past droughts is found to increase the adoption of drought-tolerant maize among farmers that also receive subsidies \citep{Katengeza2019-af} and the dis-adoption of traditional local maize \citep{Holden2016-vy}.
+We contribute to this literature by exploring heterogeneous comparative advantages to adoption in Ethiopia between 2011 to 2015, controlling for important drivers of adoption such as rainfall (found to be the main driver of heterogeneous effects for adopting agronomic packages \citealt{Marenya2020-kb}, \citealt{Katengeza2019-af}), labor and input use.
+Our paper also provides a novel way of visualizing GRC results, which become difficult to interpret with longer time variables.
+Our findings also provide an alternative hypothesis to the low-returns to adoption, namely that the wide dispersion of older generations and recycled improved seeds, which may be both resulting in lower yields and altering farmers' input use in response to a mis-classified seed.
+This also implies a methodological discovery when investigating hybrid seed adoption: the absence of a true control group to use for analysis.
+With more time and more genetic dilution, it becomes more and more likely that even those that do report hybrid use are still likely using some level of hybrid.
 
-%%%% Let's evaluate the effect of adoption in the context of Ethiopia
+The article continues as follows.
+In Section \ref{sec:data}, we describe the data used, the sample frame, general farmer and farm characteristics, and improved seed adoption levels.
+Section \ref{sec:strategy} presents the methodological details of our two estimation strategies.
+Section \ref{sec:results} shows the estimations results, and Section \ref{sec:conclusion} presents the conclusion and implications of our analysis.
 
-In this paper, we revisit the mechanism proposed by  \citep{Suri2011-oi} that justifies this lack of technology adoption on the existence of farmers heterogeneity.
-In particular, the fact that farmers with high net returns to the technology use new technologies while those with low returns do not.
-Adoption of a technology involves several dimensions such as market integration, access to supplementary inputs like fertilizer, as well agro-ecological considerations such as climate, altitude and soil quality.
-Ignoring the heterogeneity of these factors can hide the true impact of adoption, as different sub-populations may be impacted by the same technology in different, and even opposite ways.
-One way to tackle this problem, proposed by \citep{Suri2011-oi}, centers around quantifying the comparative advantage of a household to adopting improved seed varieties.
-Even when average returns are high, farmers may face heterogeneous returns based on their own, unobservable, comparative advantage in adopting new technologies.
-Using a correlated random coefficient model, Suri’s seminal paper finds that farmers’ lack of adoption is driven by insufficient net benefits to the technology stemming from poor infrastructure.
-Similar work in Ethiopia on improved chickpea varieties suggests that the economic returns to adoption plays a major role in the decision to adopt \citep{Michler2018-wk}.
+\section{Data}\label{sec:data}
 
-%% Expand a bit on the methods
-In our paper, we build on that literature by investigating hybrid maize seed adoption in Ethiopia, focusing on how heterogeneous returns drive the decision to use improved seed varieties.
-We implement two complementary estimation strategies.
-First, we implement at Group Random Coefficients estimation as originally devised by \cite{Suri2011-oi} and later expanded by \citep{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc}.
-The advantage of our estimation strategy is that we do not rely on identification from a valid instrument,
-but modelling the heterogeneity directly through a projection of their returns to adoption on their observed
-”trajectory” of adoption.
-This first estimation strategy uses data from the first three survey rounds that allow us to follow the same households across time (ESS1 to ESS3).
-The results show returns to adoption are consistently negative, a result that runs against the careful lab development of the improved seed varieties.
-To expand on this results, our second estimation strategy implements an endogenous switching regressions that exploits the last round of the survey (ESS4).
-The latest survey round is not a a continuation of the panel of the first three rounds, but has novel data on the genetic finger prints of the seeds used, that we exploit to understand why hybrid seed varieties seem not to increase returns when adopted.
-Varying how strict of a definition for hybrid seeds we use, starting from second-generation and recycled (SR1), and a more strict definition, which includes only new varieties (SR2), we find that  farmers that adopted a seeds with a less pure genetic profile would have had better returns by simply not adopting.
+This study uses the Ethiopia Socioeconomic Survey (ESS), part of the Agricultural Sample Survey (AgSS), a nationally representative survey designed to collect agricultural information for the major crops in Ethiopia.
+The survey is a joint effort of the the Central Statistics Agency of Ethiopia, the World Bank and CGIAR (Consortium of International Agricultural Research Centers) \citep{kosmowski2020shining}.
+The ESS contains four survey modules designed to obtain specific information about the different stages in the agricultural season.
+A post-planting questionnaire obtains information about the parcel, crop roster, seed roster and input use.
+A post-harvest questionnaire collects information about crop harvests by parcel (which includes cropcut measurements for selected plots), and crop disposition.
+These two instruments are complemented with a household and a community questionnaire, which obtain information on household characteristics, labor, land, expenditure, and community characteristics such as market prices and economic activities, respectively.
 
-We put together the results from both estimation strategies to suggest that the negative return to adoption we find under the Group Random Coefficients estimation reflects the loose definition of hybrid seeds, where only a subset of farmers are using the purest genetic varieties.
-Considering that poorer farmers are the ones paying a premium for genetically diluted hybrids, this result has several policy implication about the need to foster a more inclusive policy when helping farmers to obtain the best performing seeds.
+\subsection{The Three Round Panel}
+
+There were significant differences between the analysis carried out in ESS 1-3 and the most recent, ESS4.
+Most notably, the ESS4 is no longer a panel dataset made up of the households of ESS 1-3, but it includes important information on the DNA fingerprinting of seed type, which can be used to calculate misclassification rates.
+This is important information as the first three waves of ESS only included self-reported hybrid maize status.
+Since our estimation strategy requires a balanced panel dataset of maize growers, our main analysis includes the first three waves of the ESS.
+We use the fourth wave for validation, and to explore the impact of specific types of improved seeds.
+
+Figure \ref{map:regions} shows the spatial distribution of the locations of the households in the study.
+These are households that grow either improved or conventional maize and that are observed for the three rounds of the survey.
+Most households come from the northwest region of Amhara, Tigray and Oromiya which is consistent with where maize is mostly grown in Ethiopia \citep{Abate2015-rj}.
 
@@
     \centering
-    \caption{Maize: Total area harvested and Average yields}    \includegraphics{results/figures/Maize_yields.pdf}
-
-    \figcaption{Source: Authors' calculations using FAO stats data
-    Note: Yield means the harvested production per ha for the area under cultivation}
-    \label{fig:maize_yields}
+    \caption{Spatial Distribution of Households Surveyed}\label{map:regions}
+    \includegraphics[width=.7\textwidth]{results/figures/map_hhids.pdf}
+    \vspace*{-1em}
+    \begin{table}[H]
+        \centering
+        \begin{tabular}{p{0.6\textwidth}} 
+            \begin{tablenotes}
+                  \small
+                  \item Note: Red markers denote the locations of households
+            \end{tablenotes}
+        \end{tabular}
+    \end{table}  
 \end{figure}
 
-\section{Data}
-
-The current study uses the Ethiopia Socioeconomic Survey (ESS), a part of the Agricultural Sample Survey (AgSS), a survey designed to obtain production estimates for the major crops.
-The survey is a join effort of the the Central Statistics Agency of Ethiopia, the World Bank and CGIAR (Consortium of International Agricultural Research Centers) \citep{kosmowski2020shining} that aims at collecting information from a representative sample of households in the  the most populous regions of the country.
-The ESS contains several questionnaires targeted at getting an understanding of the state of agriculture.
-This includes a survey that is conducted post-planting and then post-harvest.
-These surveys are more important to our analysis as it gives a close look to inputs used in agriculture, the types of seeds used (in various levels of detail) and yields collected post-harvest.
-
-There were significant differences between the analysis carried out in ESS 1-3 and the most recent, ESS4.
-Most notably, the ESS4 is no longer a panel dataset made up of the households of ESS 1-3, but it includes important information on the DNA fingerprinting of seed type, with which we can calculate misclassification rates.
-This is important information as the first three waves of ESS only included self-reported hybrid maize status.
-Since our estimation strategy requires a balanced panel dataset of maize growers, our main analysis will include the first three waves of the ESS.
-We use the fourth wave later for validation, and to explore the impact of specific types of improved seeds.
-
-Figure \ref{map:regions} shows the spatial distribution of the locations of the our households of study.
-As previously mentioned these are households that grow either improved or conventional maize and that are observed for all three rounds of the survey.
-Most households come from the northwest region of Amhara, Tigray and Oromiya which is consistent with where maize is mostly grown in Ethiopia \citep{Abate2015-rj}.
-
-\begin{figure}
-    \centering
-    \caption{Spatial Distribution of Households Surveyed}
-    \includegraphics{results/figures/map_hhids.pdf}
-    \figcaption{Note: Green markers denote the locations of households}
-    \label{map:regions}
-\end{figure}
-
-Table \ref{tbl:summary} shows a set of summary statistics for our sample.
+Table \ref{tbl:summary} shows a set of summary statistics for the sample.
 The mean parcel size for a household is about 0.15 Ha, with a maximum size of a little less than 1.5 Ha.
@@
 The distances to some key areas are also highlighted in Table \ref{tbl:summary}.
-On average households are around 13 kilometers away from an asphalt road and almost 60 kilometers away from the nearest market.
+On average, households are around 13 kilometers away from an asphalt road and almost 60 kilometers away from the nearest market.
 This makes both the purchase of improved seed and the marketing of yields subject to transportation costs and a difficult enterprise.
-Fertilizer costs are on average 8,350 Birr (~178 USD) and with a large amount of the sample not using fertilizer at all.
+Fertilizer costs are on average 8,350 Birr ($\thicksim$178 USD), with a large amount of the sample not using fertilizer at all.
 This is particularly important in the context of improved maize seed use as it requires fertilizer to be most effective.
-Only 5\% of the sample irrigates their crops making households susceptible weather shocks.
+Only 5\% of the sample irrigates their crops making households susceptible to weather shocks.
 
-% \resizebox{0.7\textwidth}{!}{
-{\small\tabcolsep=3pt  % hold it local
+\afterpage{{\footnotesize\tabcolsep=1pt  % hold it local
     \input{results/tables/summary}
-}
+    \vspace*{-4em}
+    \begin{table}[H]
+        \centering
+        \begin{tabular}{p{0.7\textwidth}} 
+            \begin{tablenotes}
+                  \small
+                  \item Note: Author's calculations using ESS. Maize farmers only.
+            \end{tablenotes}
+        \end{tabular}
+    \end{table}  
+    \newpage
+}}
 
-The ESS has has both self-reported and crop cut yields available.
+The ESS has both self-reported and crop cut yields available.
 This provides two sources of yield information, although self-reported yields seem to be overly large compared to either fresh or dried crop cut yields.
 Self-reported yields are both more susceptible to behavioral biases, but also to confusion caused by differences in assumed units.
-In general, though, the sample with available yield information cuts down the available sample immensely, which limits the power of the estimation.
-It isn't clear whether unavailable yields numbers is due to some sort of selection or whether crop cuts were only collected randomly from a few households.
+In general, though, the sample with available cropcut yields cuts down the available sample significantly, given that cropcuts were only collected for a subsample of randomly selected households, which limits their power.
 
@@
 
-Defining what constitutes an improved seed is challenging because of the different characteristics of seeds.
-A crop seed is considered “improved” in Ethiopia if it was tested by breeders and evaluated to be superior to existing varieties (MoA, 2013, SPIA, 2020).
-In Ethiopia, the government controls most of the seed system.
-Local agricultural offices or extension agents assess the demands of different varieties, the MoA decides the seed quantities to be produced, and actors such as regional seed enterprises, seed companies, seed unions, are involved in the production of certified seeds; and seeds are primarily distributed by seed producer cooperatives (SPCs), which account for 37\% of total distribution in the country (SPIA, 2020).
-
-An improved seed can be obtained directly from such distribution centers (which would distribute a first generation seed), or it could be cultivated from second or posterior generations from the originally bred seed.
-Farmers are asked to report whether the variety they used was a first-generation improved seed, second generation,  recycled, or a traditional variety.
+A crop seed is considered “improved” in Ethiopia if it was tested by breeders and evaluated to be superior to existing varieties  (\citep{MoA13}; \citep{kosmowski2020shining}).
+In Ethiopia, the government controls most of the seed system, with local agricultural offices or extension agents assessing the demands of different varieties.
+The MoA decides the seed quantities to be produced, and actors such as regional seed enterprises, seed companies, seed unions, are involved in the production of certified seeds, that are primarily distributed by seed producer cooperatives (SPCs), which account for 37\% of the total distribution in the country (\citep{kosmowski2020shining}).
+An improved seed can be obtained directly from such distribution centers (which would distribute a first generation seed), or it could be cultivated from second or later generations from the originally bred seed.
+In the survey, farmers are asked to report whether the variety they used was a first-generation improved seed, second generation, recycled, or a traditional variety.
 When buying seeds, farmers might be able to identify the variety they are looking for, but it is unlikely that they would be aware of the source.
@@
 Fifty four varieties have been released in Ethiopia since 1990, out of which seven were released between 1990 to 1999, 29 between 2000 to 2009, and 18 between 2010 to 2019.
-Thirty four of such varieties contain CIMMYT germplasm, out of which, 10 Drought-tolerant maize varieties (DTMZ) and 8 Quality protein maize (QPM) were released (most of these between 2010 to 2019).
-Out of the CIMMYT-related germplasm, ten are Open-pollinated varieties (OPVs) and 25 are hybrid.
-The fourth wave of the ESS introduced for the first time DNA fingerprinting for barley, maize, and sorghum.
-The collection of crop-cuts for crop yield estimation facilitated the process for DNA fingerprinting.
-For the crop-cuts, a 4 square meter quadrant was randomly selected in each plot, so that total production harvested from such patch was weighted (both when freshly harvested and when dried).
-The DNA material was extracted from samples from such dried crop-cuts, that were then matched with the genetic material of the maize DNA reference library for Ethiopia, obtained from a CIMMYT and EIAR project.
-The collection of DNA samples focused only on eight regions.
-In each EA of such regions, a random sample of plots was selected, of a maximum of 10 fields for each crop.
-As a result, 505 plots were selected for DNA fingerprinting, which corresponded to 447 households.
+Thirty four such varieties contain International Maize and Wheat Improvement Center (CIMMYT) germplasm, out of which, 10 Drought-tolerant maize varieties (DTMZ) and 8 Quality protein maize (QPM) were released (most of these between 2010 to 2019).
+Out of the CIMMYT-related germplasm, ten are open-pollinated varieties (OPVs) and 25 are hybrid (\citep{kosmowski2020shining}).
+
+The fourth wave of the ESS introduced DNA fingerprinting of barley, maize, and sorghum for the first time.
+The collection of cropcuts for crop yield estimation facilitated the process for DNA fingerprinting.
+For the cropcuts, a 4 square meter quadrant was randomly selected in each plot, and total production harvested from the quadrant was weighed (both when freshly harvested and when dried).
+The DNA material was extracted from samples from the dried cropcuts, and were then matched with the genetic material of the maize DNA reference library for Ethiopia, obtained from a CIMMYT and Ethiopian Institute of Agricultural Research (EIAR) project.
+The collection of DNA samples focused on eight regions.
+In each enumeration area (EA) of these regions, a random sample of plots was selected, with a maximum of 10 fields for each crop.
+As a result, 505 plots were selected for DNA fingerprinting, corresponding to 447 households.
 
 The DNA fingerprinting process matched the genetic material of each sampled harvest with that of the reference library.
-The resulting matching process returns the name of the crop variety matched, its germplasm origin or their pedigree (whether they are a CIMMYT line, EIAR, IITA or private), the breeder or maintainer, and the year of their release (which could be indicative of the number of generations a seed had been cultivated), as well as the genetic purity of the sample (the degree of no contamination by other genetic varieties).
+The resulting matching process returns the name of the crop variety matched, its germplasm origin or their pedigree (whether they are a CIMMYT line, EIAR,  International Institute of Tropical Agriculture (IITA) or private), the breeder or maintainer, and the year of their release (which could be indicative of the number of generations a seed had been cultivated), as well as the genetic purity of the sample (the degree of non-contamination with other genetic varieties).
 It also informs whether the harvested seeds are open-pollinated varieties (OPVs) or hybrids; drought-tolerant maize varieties (DMTZ) or quality protein maize (QPM).
@@
 We find that 38\% of households self-report adopting a “new” improved variety (SR2 in Figure \ref{fig:adoption_r4}), and 43\% report adopting a “new”, “second generation” or “recycled” improved variety (SR1).
-Considering a household as adopter if the household planted an improved variety in at least one of their plots, all households where a DNA sample was obtained used a maize improved seed, considering a purity threshold of 70\% (DNA 70).
-When purity thresholds of 90 and 95\% are considered (DNA 90, DNA 95 in Figure \ref{fig:adoption_r4}, respectively), adoption rates change to 91 and 54\% of households, respectively.
-62\% of households adopted a High-yielding variety, compared to 38\% that adopted an open-pollinated variety, and 15\% of households adopted a Drought-tolerant maize variety.
-63\% of households that adopted an improved variety are using a CGIAR-derived germplasm, and 79\% are using an exotic germplasm.
-Finally, 64\% of households adopted a variety that was released on the year 2000 or after, and 24\% adopted a variety that was adopted in the year 2010 or after.
+Considering a household as adopter if the household planted an improved variety in at least one of their plots, leads to classifying all households where a DNA sample was obtained as users of improved seed variety, a result that holds when the purity threshold of 70\% (DNA 70) is used.
+It is only when using purity thresholds of 90 and 95\% (DNA 90, DNA 95 in Figure \ref{fig:adoption_r4}, respectively), that adoption rates of improved varieties change to 91 and 54\% of sampled households, respectively.
+We also find that 62\% of households adopted a hybrid variety, compared to 38\% that adopted an open-pollinated variety, and 15\% of households adopted a drought-tolerant maize variety, while 63\% of households that adopted an improved variety use a CGIAR-derived germplasm, and 79\% use an exotic germplasm.
+Finally, 64\% of households adopted a variety that was released in the year 2000 or after, and 24\% adopted a variety that was adopted in the year 2010 or after.
 
-\begin{figure}
+\begin{figure}[htpb]
     \centering
-    \caption{Adoption rates of improved varieties in 2018 (by self-report, purity level, crop type, source and year of release)}
-    \includegraphics{results/figures/adoption_r4.png}
-    \figcaption{Note: Shares represent the proportion of adopters among maize farmers.
+    \caption{Adoption rates of improved varieties in 2018 (by self-report, purity level, crop type, source and year of release)}\label{fig:adoption_r4}
+    \includegraphics[width=.7\textwidth]{results/figures/adoption_r4.png}
+    \begin{table}[H]
+    \centering
+        \begin{tabular}{p{0.7\textwidth}} 
+            \begin{tablenotes}
+                  \small
+                  \item Note: Shares represent the proportion of adopters among maize farmers.
 Acronyms include: SR1 [=1 if New, 2nd gen or recycled, =0 if Traditional]; SR2  [=1 if New, =0 if 2nd gen, recycled or traditional]; DNA70 - Improved if purity $>=$70$\%$; DNA90 - Improved if purity $>=$90$\%$; DNA95 - Improved if purity $>=$95$\%$; HYB - Improved DNA, [=1 Hybrid, 0=Open pollinated]; DTM - Drought tolerant Maize [=1 DTM, 0=otherwise], CGIAR source [=1 if germplasm associated to a CGIAR-related thread], Exotic source [=1 if germplasm associated to an exotic source], Released 2000+ [=1 if DNA associated to a variety released on year 2000 or later], Released 2010+ [=1 if DNA associated to a variety released on year 2010 or later].
-QPM is not included as there are too few households planting QPM seeds (n=6).}
-    \label{fig:adoption_r4}
+QPM is not included as there are too few households planting QPM seeds (n=6).
+            \end{tablenotes}
+        \end{tabular}
+    \end{table}   
 \end{figure}
@@
 These ratios are in line with prior findings.
-In the regions of Oromia, Amhara, Tigray, and SNNPR, adoption rates of improved maize varieties of 39\% of households were found, according to self-reported information (Zeng et al., 2015).
-In the regions of Amhara, Benishangul-Gumuz, Oromia, SNNPR, and Tigray, adoption rates of improved maize varieties were found for 27\% of households (Jaleta et al., 2018).
-In Oromia, SNNPR, and Benishangul-Gumuz adoption of improved varieties per farmers’ self-reports was found for 84\% of maize growers in 2011, and 88\% of maize growers in 2013 (Yirga et al., 2017).
+In the regions of Oromia, Amhara, Tigray, and SNNPR, adoption rates of improved maize varieties of 39\% of households were found, according to self-reported information (\cite{Zeng15}).
+In the regions of Amhara, Benishangul-Gumuz, Oromia, SNNPR, and Tigray, adoption rates of improved maize varieties were found for 27\% of households (\cite{Jaleta2018-oj}).
+In Oromia, SNNPR, and Benishangul-Gumuz adoption of improved varieties per farmers’ self-reports was found for 84\% of maize growers in 2011, and 88\% of maize growers in 2013 (\cite{Yirga17}).
 
-Understanding the effects of improved seed adoption is not only challenging because there could be multiple ways of establishing what constitutes an improved seed, but also because such definitions may or may note overlap, or because farmers may also missclassify a variety as improved.
+Understanding the effects of improved seed adoption is not only challenging because there could be multiple ways of establishing what constitutes an improved seed, but also because such definitions may or may not overlap, or because farmers may also misclassify a variety as improved.
 Figure \ref{fig:missclassification} shows the correlation between the different definitions discussed above.
-For those seeds for which genetic material was obtained, the correlation with the self-reported seeds indicator is low, especially when using lower cutoff thresholds.
+For those seeds for which genetic material was obtained, the correlation with the self-reported seeds indicator is low, especially when using lower purity cutoff thresholds.
 
-\begin{figure}
+\begin{figure}[htpb]
     \centering
-    \caption{Correlation of different definitions of improved seeds, 4th round}
+    \caption{Correlation of different definitions of improved seeds, 4th round}\label{fig:missclassification}
     \includegraphics[width = 0.9\textwidth]{results/figures/missclassification.png}
-    \figcaption{Note: Shares represent the proportion of adopters among maize farmers.
+    \vspace*{-2em}
+    \begin{table}[H]
+        \centering
+        \begin{tabular}{p{1\textwidth}} 
+            \begin{tablenotes}
+                  \small
+                  \item Note: Shares represent the proportion of adopters among maize farmers.
 Acronyms include: SR1 [=1 if New, 2nd gen or recycled, =0 if Traditional]; SR2  [=1 if New, =0 if 2nd gen, recycled or traditional]; DNA70 - Improved if purity $>=$70$\%$; DNA90 - Improved if purity $>=$90$\%$; DNA95 - Improved if purity $>=$95$\%$; HYB - Improved DNA, [=1 Hybrid, 0=Open pollinated]; DTM - Drought tolerant Maize [=1 DTM, 0=otherwise], CGIAR source [=1 if germplasm associated to a CGIAR-related thread], Exotic source [=1 if germplasm associated to an exotic source], Released 2000+ [=1 if DNA associated to a variety released on year 2000 or later], Released 2010+ [=1 if DNA associated to a variety released on year 2010 or later].
-QPM is not included as there are too few households planting QPM seeds (n=6).}
-    \label{fig:missclassification}
+QPM is not included as there are too few households planting QPM seeds (n=6).
+            \end{tablenotes}
+        \end{tabular}
+    \end{table}
+
 \end{figure}
 
-\section{Estimation Strategy}
+\section{Estimation Strategy}\label{sec:strategy}
 
@@
 
-The advantage of our estimation strategy is that we do not rely on identification from a valid instrument, but modelling the heterogeneity directly through a projection of their returns to adoption on their observed "trajectory" of adoption.
-This allows us to directly capture two effects: the heterogeneous effects of the so-called "comparative advantage" of adoption, and, under the assumption of linearity in comparative advantage, a description of the overall economy's sorting of farmers based on their comparative advantage.
+Since ESS1-ESS3 do no have DNA fingerprinting, we are left with only self-reports of hybrid use.
+We use these variables in a Group Random Coefficiens (GRC) approach to investigate the heterogeneous effects of adoption.
+The advantage of the GRC approach is that we do not rely on identification from a valid instrument, but on modelling the heterogeneity directly through a projection of adoption returns on their observed ``trajectory'' of adoption.
+This allows us to directly capture two effects: heterogeneous effects in the form of so-called ``comparative advantage" of adoption, and, under the assumption of linearity in comparative advantage, a description of the overall economy's sorting of farmers based on their comparative advantage.
 The first is important because it is able to disentangle the effect of a general household's agricultural ability (which can be thought of as a fixed effect) and their adoption-specific potential.
-The sorting parameter can tell us whether the economy sorts such that those that would benefit the most adoption actually adopt, or whether there are barriers that stop that sorting from happening.
-See \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} for more information on the assumptions involved for the Group Random Coefficients estimator.
+The second parameter can tell us whether the economy sort farmers in a way that is commensurate with providing those that would benefit the most from adoption actually adopting, or whether there are barriers interfering with that.
+See \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} for more information on the assumptions involved for the GRC approach.
 In our particular case, we contribute to the literature and methodology of \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} by providing reasonable ways to aggregate the results of the GRC model, which may seem difficult to interpret in its raw form.
-We use these aggregations to better understand how adoption affects households across time, as well as whether the frequency of adoption has a strong effect.
+We use these aggregations to better understand the heterogeneous effects to adoption.
 
@@
 
-Table \ref{tbl:Trajectories} maps out the frequencies of the different trajectories present in the data, sorted by frequency.
-The largest portion of the data are those that do not adopt improved maize seed at all, followed with sharp decline, to those that use improved maize seed in every period.
-Interestingly, adoption trajectories seem to be more popular for later adoptions than earlier ones.
+Table \ref{tbl:Trajectories} maps out the frequencies of the trajectories present in the data, sorted by frequency.
+Adoption trajectories seem to be more popular for later adoptions than earlier ones.
 Most households do not adopt at all (``never-adopters''), but the second most frequent trajectory are those that adopt in every wave (``always-adopters'').
@@
 where $\phi$ is the parameter which describes sorting in the economy and $\Delta_{0,0,1}$ are the returns to adoption for trajectory $(0,0,1)$.
-If $\phi$ is positive, then higher yields is associated with higher comparative advantage, denoting that those that adopt also tend to have the highest returns to adoption.
+If $\phi$ is positive, then higher yields are associated with higher comparative advantage, denoting that those that adopt also tend to have the highest returns to adoption.
 If it is negative, however, this denotes that there might be barriers, such as a lack of fertilizer access, and bad roads that restrict the ability to buy hybrid seed.
 $\phi$ is estimated in two ways: through the restricted model just mentioned and by a weak-identification brute force grid procedure, which allows for calculating the confidence interval for $\phi$ for multiple sets of trajectories.
-The advantage of the weak-identification robust procedure is that we can test for whether the data satisfies the conditions necessary for an interpretable $\phi$ parameter.
+The advantage of the weak-identification robust procedure is that we can test for whether the data satisfies the conditions necessary for an interpretable $\phi$ parameter, before estimating the restricted model.
 If there are trajectories which yield a $\phi$ with a non-infinite confidence interval, then the LCA is at least partially satisfied.
@@
 The comparative advantage of each trajectory tells us how much a household with a given trajectory benefits from adoption.
-$\theta <0 $ denotes that returns to adoption are negative and that a household loses by adopting.
+$\theta <0 $ denotes that adoption-specific potential is negative and that a household loses by adopting.
 
 Once we generate $\theta$ and $\Delta$ for each trajectory, we can examine heterogeneous effects to adoption.
-In its raw form, that means that have eight comparative advantages available for comparison, which is rich in its information, but difficult to interpret as a result, we will aggregate $\theta$ in two ways to improve interpretability: (1) which wave the household adopted, and (2) the number of time adoption took place.
+Conditional on the LCA being satisfied, that means that there will be eight comparative advantages available for comparison, which is rich in its information, but difficult to interpret.
+As a result, we will aggregate $\theta$ in two ways to improve interpretability: (1) by which wave the household adopted, and (2) the number of times adoption took place.
 We do this by taking a simple average, grouped by each of these factors.
+The first aggregation will tell us whether there are any wave-specific differences in comparative advantage.
+The second aggregation will tell us whether adopting more times corresponds to higher comparative advantage and associated returns.
+In both cases, these aggregations tell us something about the economic landscape that drive returns.
 
-\subsection{Endogenous Switching regression}
-The ESS4 is a unique data-set that includes DNA fingerprinting of maize seed type.
-This brings new possibilities in terms of exploring....
-sayESS4 no longer part of panel but that still useful,....
-\par
-The group random coefficient model used the variation in the adoption trajectories of each observation in the ESS 1-3 panel as a source of information that allows for an unbiased estimation.
-ESS4 is no longer part of this panel data-set, hence, in order to estimate the effect of the adoption of improved seeds on yields, an alternative estimation method is needed.
+\subsection{Endogenous Switching Regression}
+
+The GRC approach uses the variation in the adoption trajectories of each observation in the ESS 1-3 panel as a source of information that allows for an unbiased estimation.
+ESS4 is no longer part of this panel data-set, and so in order to estimate the effect of adoption of improved seeds on yields, an alternative estimation method is needed.
 We propose an endogenous switching regression model which accounts for unobserved heterogeneity in adoption decisions.
-This method imposes restrictive assumptions regarding the distribution of the errors of the selection and output equations.
+This method imposes restrictive assumptions regarding the distribution of the errors for the selection and output equations.
 At the same time, the exclusion restriction needs to be satisfied in order to reduce the bias of the estimation.
-Nevertheless, the endogenous switching regression holds certain advantages over other methodologies used for identifying unbiased estimates with cross-sectional observational data such as propensity score matching or other instrumental variable approaches (\cite{shiferaw2014adoption}) and has been used to study similar questions in the past (\citealt{falco2011does}, \citealt{kabunga2012yield}).
+
+The endogenous switching regression holds certain advantages over other methodologies used for identifying unbiased estimates with cross-sectional observational data such as propensity score matching or other instrumental variable approaches (\cite{Shiferaw2014-op}) and has been used to study similar questions in the past (\citealt{falco2011does}, \citealt{kabunga2012yield}).
 \par
@@
 Equation \ref{eq:switch_latent} shows the definition for a latent variable $A^*$ that reflects the benefit from adopting an improved seed relative to planting a traditional one.
-As shown in the equation, household $i$ will decide to adopt ($A_i$=1)if these perceived benefits of adoption are positive.
+As shown in the equation, household $i$ will decide to adopt ($A_i$=1) if these perceived benefits of adoption are positive.
 
 \begin{align}
-A_i^*=Z_i\alpha+\varepsilon_i \; \text{with} \, A_i    \begin{cases}
+A_i^*=\bm{Z}_i\alpha+\varepsilon_i \; \text{with} \, A_i    \begin{cases}
       1, & \text{if}\ A_i^*>0 \\
@@
 \end{align}
-Vector $Z$ includes variables that determine if a particular household may benefit from the adoption of improved seeds, such as farmers' demographic characteristics (education, age, female headed households) and land tenure status of their plots.
+
+Vector $\bm{Z}$ includes variables that determine if a particular household may benefit from the adoption of improved seeds, such as farmers' demographic characteristics (education, age, female headed households) and land tenure status of their plots.
 In the estimations presented in the next sections, we consider different definitions of improved seeds, and include not only self-reported seed classification but also the results of DNA fingerprinting.
 \par
-The second stage presents the outcome functions conditional on adoption:
+The second stage presents the outcome functions conditional on adoption, where farmers choose between adopting an improved seed (regime 1) or using a traditional seed (regime 2):
 \begin{align}
+    \text{Regime 1}: \; y_{1i}=\bm{X}_{1i}\beta_1+ \epsilon_{1i} \; if \; A_i=1 \label{eq:switch_reg1}\\
+     \text{Regime 2}: \; y_{2i}=\bm{X}_{2i}\beta_1+ \epsilon_{2i} \; if \; A_i=0 \label{eq:switch_reg2} 
+\end{align}
+$y_{1i}$ and $y_{2i}$ are yields under regime 1 or regime 2 for household $i$, and $\bm{X}$ is a vector of variables that affect yields including parcel size, household and hired labor, costs of fertilizers and other inputs, whether the household has access to irrigation or farm mechanization, among others.
+\par
+The model assumes that the error terms of Equations \ref{eq:switch_latent}, \ref{eq:switch_reg1}, \ref{eq:switch_reg2} have a trivariate normal distribution with mean equal to 0 and a covariance matrix with diagonal terms equal to $\sigma_{\varepsilon}^2$, $\sigma_1^2$ $\sigma_2^2$, and non-zero covariances between $\varepsilon$ and $\epsilon_1$ ($\sigma_{1\varepsilon}$) and between $\varepsilon$ and $\epsilon_2$ ($\sigma_{2\varepsilon}$), and not defined between $\epsilon_1$ and $\epsilon_2$ since both regimes are not observed simultaneously for any given $i$.
+This error structure is derived from the expected value $\epsilon_1$ ($\epsilon_2$) conditional on adoption  (non-adoption) respectively to be non-zero, reflecting the endogeneity in the adoption decision.
+These conditional expectations are equal to:
+\begin{align*}
+    E(\epsilon_{1i}|A_i=1)=&\sigma_{1\varepsilon}\frac{\phi(\bm{Z}_i\alpha)}{\Phi(\bm{Z}_i\alpha)}=\sigma_{1\varepsilon}\lambda_{1i} \\
+    E(\epsilon_{2i}|A_i=0)=&-\sigma_{2\varepsilon}\frac{\phi(\bm{Z}_i\alpha)}{(1-\Phi(\bm{Z}_i\alpha))}=\sigma_{2\varepsilon}\lambda_{2i}
+\end{align*}
+%
+where $\phi(\cdot)$ and $\Phi(\cdot)$ are the standard normal probability density and cumulative density functions, $\lambda_{1i}=\sigma_{1\varepsilon}\frac{\phi(\bm{Z}_i\alpha)}{\Phi(\bm{Z}_i\alpha)}$ and $\lambda_{2i}=-\sigma_{2\varepsilon}\frac{\phi(\bm{Z}_i\alpha)}{(1-\Phi(\bm{Z}_i\alpha))}$.
+\par
+We estimate this model simultaneously using maximum likelihood estimation following \cite{lokshin2004maximum}, which accounts for the endogeneity in the yield equation.
+Statistically significant estimates for the covariances $\sigma_{1\varepsilon}$ and $\sigma_{2\varepsilon}$ would indicate the presence of endogeneity in the adoption and yield equations.
+Moreover, this approach allows us to compare the expected yields for adopters under the two regimes, the observed (adoption) and the counterfactual (non-adoption).
+This effect is the Average Treatment Effect on the Treated (ATT).
+Analogously, it is also possible to calculate the Average Treatment Effect on the Untreated (ATU) by comparing expected yields under the two regimes for non-adopters: the counterfactual (adoption) and the observed (non-adoption).
+The following equations present the expected yield equations for adopters and non-adopters under both regimes:
 
+\begin{subequations}
+\begin{align}
+    E(y_{1i}|A_i=1)=\bm{X}_1i\beta_1+\sigma_{1\varepsilon} \lambda_{1i} \label{eq:att_1} \\
+    E(y_{2i}|A_i=1)=\bm{X}_1i\beta_2+\sigma_{2\varepsilon} \lambda_{1i} \label{eq:att_0} \\
+    E(y_{2i}|A_i=0)=\bm{X}_2i\beta_2+\sigma_{2\varepsilon} \lambda_{2i} \label{eq:atu_0} \\
+    E(y_{1i}|A_i=0)=\bm{X}_2i\beta_1+\sigma_{1\varepsilon} \lambda_{2i} \label{eq:atu_1}    
 \end{align}
+\end{subequations}
 
-\section{Results}
+The ATT is equal to the difference between Equation \ref{eq:att_1} and Equation \ref{eq:att_0}.
+The ATU is the difference between Equation \ref{eq:atu_0} and Equation \ref{eq:atu_1}.
+Note that the observed expectations are those linked to Equations \ref{eq:att_1} and \ref{eq:atu_0} for adopters and non-adopters, while the counterfactual expectations are those presented in Equations \ref{eq:att_0} and \ref{eq:atu_1} respectively.
+In the next section we report these results with bootstrapped standard errors in order to measure the yield benefit of adoption for adopters and non-adopters.
+
+\section{Results}\label{sec:results}
 
@@
 Each measure of yields has its benefits and drawbacks.
-Cropcuts, for instance, provides a true estimate of yields for a given parcel of land, but it is at the cost of a reduced sample size.
-Self-reports have larger sample sizes, but may suffer from measurement error.
+Cropcuts, for instance, provide a true estimate of yields for a given parcel, but at the cost of a reduced sample size.
+Self-reports have larger sample sizes, but may suffer from measurement error \citep{gollin2021heterogeneity}.
 For controls we use the years of education of the household head, the age of the household head, the sex of the household head, whether there is a title for the land being cultivated, the parcel size in hectares, as well as the hours of household and hired labor.
-Commensurate with the literature, we also include interactions of these controls with hybrid adoption as added controls.
-Interactions with the hybrid variable are meant to account further account for household specific variation that adopter may have.
+Commensurate with the literature, we also include interactions of these controls with improved seed adoption as added controls.
+Interactions with the improved seed variable are meant to further account for household specific variation that adopters may have.
+
+%  self-reported yields are systematically higher than cropcut yields, pointing to the fact that there may be some measurement error at play.
+
 The results of the regression are shown in Table \ref{tbl:unres}.
+The results for self-reported and cropcut yields are largely similar.
+There is some discrepancy in the self-reported yields in the first wave, and specifically with the $(1,0,0)$ trajectory.
+The marginal effect to adoption is negative for cropcuts, but positive for self-reported yields.
+The marginal effects to adoption in many cases, are not statistically significant.
+Only in the case of the trajectory $(0,0,1)$: those that adopted in the second and third waves, is there a large, statistically significant effect.
 
-The results for self-reported and cropcut yields are largely similar, pointing to the fact that self-reported yields may not have a large amount of measurement error.
-In the base case, there are some positive effects to adoption, but with all controls included, marginal effects of adoption for each trajectory are negative and significant.
-This translates to average yields of adoption that close to 0 or even negative.
+In order to calculate comparative advantage, there must be evidence for the LCA being satisfied.
+The lower panel of  Table \ref{tbl:unres} shows the results for the weak identification test for $\phi$.
+We can see here that for the case of cropcuts with controls, there is a confidence interval for the joint test of $\phi$, giving evidence to the fact that the LCA is satisfied most prominently in that case.
+For the aggregations for this paper, then, we will use that specification.
 
 % {\small\tabcolsep=3pt  % hold it local
-\resizebox{1\textwidth}{!}{
+
 \input{results/tables/grc}
-}
 
-\resizebox{1\textwidth}{!}{
-\centering
-% \input{results/tables/phi_test_merged}
-\input{results/tables/phi_test_self}
-\input{results/tables/phi_test_cropcut}
-}
-% \input{results/tables/grc_YIELD_selfr_tr}
+% \resizebox{0.8\textwidth}{!}{
+%     \centering
+%     % \input{results/tables/phi_test_merged}
+%     \input{results/tables/phi_test_self}
+%     \input{results/tables/phi_test_cropcut}
+% }
+% \resizebox{0.8\textwidth}{!}{
+%     \centering
+%     % \input{results/tables/phi_test_merged}
+%     \input{results/tables/phi_test_self}
+%     \input{results/tables/phi_test_cropcut}
+% }
+
+Figure \ref{fig:theta_delta_raw} shows a plot of the comparative advantages and returns for the cropcut model with controls.
+We can see here that returns are very high for the $(0,0,1)$ trajectory, but with an exceedingly negative comparative advantage.
+Conversely, all trajectories that adopted in the first wave have a positive comparative advantage, but negative returns.
+This is indicative of either negative sorting (consistent with the negative $\phi$ parameter that we found with the weak identification results) or with a systematic problem in adoption in the first year.
 
@@
     \caption{Comparative Advantage and Returns by Trajectory (Dry Cropcuts)}
-    \includegraphics[scale=0.75]{results/figures/theta.png}\label{fig:theta_delta_raw}
+    \includegraphics[scale=0.7]{results/figures/theta.png}\label{fig:theta_delta_raw}
+        \vspace*{-2em}
+    \begin{table}[H]
+        \centering
+        \begin{tabular}{p{0.8\textwidth}} 
+            \begin{tablenotes}
+                  \small
+                  \item Note: Author's calculations using ESS. The figures shows the comparative advantage and the returns to adoption for each adoption trajectory. Dry cropcuts only
+            \end{tablenotes}
+        \end{tabular}
+    \end{table}
 \end{figure}
+
+This is illustrated in Figure \ref{fig:when_adopt}, where we calculate an aggregation for comparative advantage and returns whether a household adopted in each wave.
+We can see here that adopting in the first wave is consistent with a positive comparative advantage, but negative returns.
+This is repeated, but less strongly in wave 2.
+This means that for waves 1 and 2, households have gains to adopting, but actually receive negative returns to doing so.
+This is indicative of the fact that there are barriers to those that adopted in waves 1 and 2, or perhaps, that returns from improved seed are not as high as expected.
+This relationship is reversed in the third wave, albeit both comparative advantage and returns are close to 0.
+Figure \ref{fig:when_adopt} provides a curious trend across time to adoption: it would seem that negative returns to adoption reduce across time, but adoption-specific advantages in households also get smaller.
+
+\begin{figure}
+    \caption{Aggregation Results: Comparative Advantage and Returns by Wave of Adoption (Dry Cropcuts)}
+    \label{fig:when_adopt}
+    \centering
+    \includegraphics[scale=0.75]{results/figures/when_adopt_cropcut.png}
+    \vspace*{-2em}
+    \begin{table}[H]
+        \centering
+        \begin{tabular}{p{0.9\textwidth}} 
+            \begin{tablenotes}
+                  \small
+                  \item Note: Author's calculations using ESS.
+The figures shows the aggregate impact from Table \ref{tbl:unres} using on which period a person adopted as the basis for the aggregation.
+            \end{tablenotes}
+        \end{tabular}
+    \end{table}
+\end{figure}
+
+Figure \ref{fig:traj_sum} provides another aggregation of returns and comparative advantage using the times a household adopted hybrid maize.
+Adopting 0 times leads to negative comparative and advantage and negative returns makes sense.
+Those that do not adopt would not receive benefits from adopting.
+However, those that adopt once during waves 1-3, see a positive comparative advantage, but negative returns.
+This same is also starkly true for those that adopted three times.
+For those that adopted twice, it would seem that comparative advantage is negative and there are negative returns to adoption.
+The lack of a pattern in Figure \ref{when_adopt} seems to suggest that the number of times adopting is not as conclusive as \emph{when} households adopted.
+Specifically, later adoption periods lead to smaller negative comparative advantages and less negative returns.
+This suggests that the value of adopting is getting higher, and that this may be due to the introduction of newer and more effective seed.
 
@@
     \centering
+    \caption{Aggregation Results: Comparative Advantage and Returns by the Number of Times Adopted (Dry Cropcuts)}
+    \label{fig:traj_sum}
     \includegraphics[scale=0.75]{results/figures/traj_sum.png}
-    \caption{Aggregation Results: Comparative Advantage and Returns by the Number of Times Adopted (Dry Cropcuts}
-    \label{fig:traj_sum}
-\end{figure}
-
-\begin{figure}
+    \vspace*{-2em}
+    \begin{table}[H]
         \centering
-    \includegraphics[scale=0.75]{results/figures/when_adopt_cropcut.png}
-    \caption{Aggregation Results: Comparative Advantage and Returns by Wave of Adoption (Dry Cropcuts}
-    \label{fig:when_adopt}
+        \begin{tabular}{p{0.75\textwidth}} 
+            \begin{tablenotes}
+                  \small
+                  \item Note: Author's calculations using ESS.
+The figures shows the aggregate impact from Table \ref{tbl:unres} using the number of time a person adopted as the basis for the aggregation.
+            \end{tablenotes}
+        \end{tabular}
+    \end{table}
 \end{figure}
@@
 
-We first examine the results of the Endogenous Switching Regression using a self-reported definition of an improved variety.
-As before, we consider a more loose definition of improved variety, which includes new seeds but also second-generation and recycled (SR1), and a more strict definition, which includes only new varieties (SR2).
+To complement the results of the GRC model, we present the Endogenous Switching Regression results.
+We first examine a self-reported definition of an improved variety to validate the results from the GRC model.
+As before, we consider a looser definition of improved seeds, which includes new seeds but also second-generation and recycled (SR1), and a more strict definition, which includes only new varieties (SR2).
 Table \ref{tab:switch1} shows the results of the Switching regression using both definitions, and using self-reported and cropcut yields.
-We see that regardless of the definition of improved seed, yields measurement form and types of controls included, the results consistently show that the Average Treatment effect on the Treated (ATT) is negative and the Average Treatment effect on the Untreated (ATU) is positive, meaning that farmers that adopted a self-reported improved seed would've been better off not adopting, and farmers that did not adopt a self-reported improved seed would've been better off adopting.
+We see that regardless of the definition of improved seed, yields measurement form and types of controls included, the results consistently show that the ATT is negative and the ATU is positive, meaning that farmers that adopted a self-reported improved seed would've been better off not adopting, and farmers that did not adopt a self-reported improved seed would've been better off adopting.
 
@@
 We therefore consider as non-adopters those that self-reported using a traditional variety.
-We are aware that given the universality of improved varieties with a 70\% threshold or higher and given the high rates of missclassification, the control group for a 70\% threshold is likely to be contaminated with a high share of false-positives.
-But we also know the results are biased downwards and that the control group for higher shares of purity will be less contaminated, which may provide an interesting comparison.
+We are aware that given the universality of improved varieties with a 70\% threshold or higher and given the high rates of miss-classification, the control group for a 70\% threshold is likely to be contaminated with a high share of false-positives.
+But we also know the results are biased downwards and that the control group for higher shares of purity will be less contaminated, which provides an interesting comparison.
 
@@
 Farmers that adopted DTMZ (as tracked by the DNA germplasm) did obtain greater yields than if they had not adopted.
-Again this result may be biased downwards because of possible false-positives among the control group.
+Again, this result may be biased downwards because of possible false-positives among the control group.
 The results for the comparison between hybrid and open-pollinated varieties are not significant (when using self-reported yields).
@@
 
-To further disaggregate the results, we finally explore the source of the improved varieties (CGIAR or exotic) and the year of the seed release.
+To further disaggregate the results, we explore the source of the improved varieties (CGIAR or exotic) and the year of the seed release.
 We find positive effects for adopters of CGIAR-sourced varieties when using self-reported yields.
@@
 
-\section{Discussion and Implications}
+\section{Conclusion}\label{sec:conclusion}
 
-- Why do we see that increasing purity leads to positive ATE
-- may be due to improved seed being more effective, but doesnt tell the whole story
-- barrier to purchasing the seed and associated inputs for poorer households, leading to reuse of seeds and being less effective
+% - Why do we see that increasing purity leads to positive ATE
+% - may be due to improved seed being more effective, but doesnt tell the whole story
+% - barrier to purchasing the seed and associated inputs for poorer households, leading to reuse of seeds and being less effective
 
-Our work has far-reaching policy implications.
-The adoption and impact of agricultural technologies is inherently heterogeneous.
-Our methodology incorporates that from the outset, allowing us to investigate whether synergies between seed adoption and water conservation strategies interact; a key component to understanding how investment into such projects translates to economic returns.
+Our results showcase the challenges of identifying the returns to technology adoption when adoption is not promoted by a randomized control trial and when the characteristics of the technology (genetic purity) are not easily observed.
+We show that in the case of Ethiopia, the return to adoption of improved seed varieties appears to be negative, a result that we attempt to further understand by relying on seeds' DNA fingerprinting, obtained from the last round of the ESS.
+Accounting for the decay of the purity of the genetic material, its source and years upon release, reveals that those farmers that rely on newly bred improved varieties do show positive returns to adoption.
+Even without explicit DNA fingerprinting evidence, the GRC was able to show this trend with the comparative advantage results; returns are negative to those that have a high comparative advantage to adopting, but these returns less negative with time.
+We speculate that this is due to the fact that new and more genetically pure hybrids are introduced giving those that adopt higher yields than before.
+
+But the fact that newly bred hybrids that adapted to local conditions outperform local varieties and older-generation improved seeds is not surprising given the large investments devoted to improve seed performance and the speed at which new high-yielding varieties are being developed.
+The results from the Group Random Coefficient model highlights how the fact that we imperfectly observe the use of newly improved seeds and the degree of mis-classification between self-reported improved seeds and bred hybrids may partially explain why we do not see self-reported improved varieties grant positive returns to adoption.
+Reconciling these findings with the results from the Endogenous Switching Regression, we hypothesize that the negative returns to adoption mask a large heterogeneity in the returns to adoption.
+The switching results show that the positive returns to adoption only appear for those farmers using newly bred hybrid seeds provided directly by distributors.
+
+This result buttresses the fact that a large share of farmers are not able to obtain the promised greater yields from using hybrid varieties.
+This may be partially driven by the low-purchasing rates of newly bred varieties, but also by the under-use of critical inputs needed for the successful performance of hybrid seeds, such as fertilizers and irrigation, especially in drought-prone areas.
+More research is needed to understand the barriers to adoption and input use.
+Such barriers may be crucial for households with less resources, which might be more likely to reuse improved seeds for many generations or buy older-generations from other farmers at a cheaper price.
+The fact that all the sampled plots chosen for DNA fingerprinting showed purity percentages of 70 percent or higher is indicative that this might be the case, given the widespread dilution of improved genetic material.
+Such dispersion of improved genetic material also poses a challenge for the proper identification of the returns to adoption as it confounds the potential benefits of locally adapted hybrids, but also the benefits of traditional local varieties that may be better suited for certain agro-ecological conditions and for contexts of low-input use.
+
+Our findings are in line with prior research.
+\cite{Bird2020-nt} propose a theoretical model where the authors compare retained local traditional varieties with improved non-locally adapted varieties and locally adapted improved varieties.
+The authors postulate that while profits of both types of improved varieties increase with fertilizer application, profits of traditional varieties would decrease with fertilizer application, implying the existence of a threshold of fertilizer application below which it is more profitable to keep using traditional varieties.
+Non-locally adapted varieties are less profitable without complementary fertilizer application.
+Depending on how responsive to fertilizer both improved varieties are and on the costs of seeds, there may be a threshold of fertilizer application below which it is more profitable to use the locally adapted variety than the improved non-adapted one.
+\par
+
+It is possible that similar dynamics explain the patterns showcased in our results.
+The evidence is consistent in that newly bred varieties do offer higher performance than traditional varieties, but such higher-yielding effects diminish over time, as newer generations of the improved seeds continue to be used.
+In contrast, traditional local varieties are well-adapted to the local agro-ecological conditions, and their reproduction over multiple generations shows consistent yields, assuming normal weather conditions.
+Depending on the specific improved varieties and the speed at which high-yielding effects diminish over time, there may be a threshold in time above which it may be more profitable to use a traditional variety than an improved one that has been around for too many generations.
+After such a threshold is reached, because improved varieties are more dependent on external inputs, the costs of the older-generation improved seeds, together with the costs of fertilizers and other inputs may not compensate the yield returns of an older-improved variety.
+\par
+
+Our results have far-reaching policy implications.
+The adoption and impact of genetically improved seeds are inherently heterogeneous.
+Our methodology incorporates this from the outset.
+Our findings highlight the wide distribution of the yield returns to adoption, partially explained by the low purchase of newly bred hybrids and the high dispersion of diluted improved genetic material over time.
+This points to the need for making policies oriented towards increasing the adoption of improved seeds and yields more inclusive, aiming at ensuring that improved varieties are not only available to purchase locally but also affordably.
+It is important to consider that improved varieties should also be locally adapted to specific agro-ecological conditions, such that their success is less dependent on external chemical inputs.
+Our findings also shed light on the need to better understand adaptation strategies when the conditions needed for the successful performance of improved seeds are lacking.
+These contexts include but are not limited to places subject to high weather variability or where low soil fertility is prevalent due to the use of excessive intensification and chemical input dependence.
+Resilience enhancing strategies such as the use of drought-tolerant maize or irrigation use may be critical for such contexts, but their suitability for the heterogeneous pool of farmers needs to be better understood.
+
+%allowing us to investigate whether synergies between seed adoption and water conservation strategies interact; a key component to understanding how investment into such projects translates to economic returns.
 Moreover, our research informs on not only the adoption but also dis-adoption decisions of technologies as well.
 In this sense, we provide valuable information to policymakers on how to better target seed marketing or infrastructure to help agricultural households improve their livelihoods
-
-\section{Conclusion}
 

### 2022-07-21 document/slides/SPIA Final Meeting/finalpresentation.tex

--- 2022-01-27 document/slides/finalpresentation.tex
@@
 \usepackage[utf8]{inputenc}
-
+\usepackage{threeparttable}
 %\usepackage{utopia} %font utopia imported
@@
 
-\author[et al]{Michuda, Chiarella, Barriga-Cabanillas, Correa, }
+\author[et al]{Aleksandr Michuda \\ Cristina Chiarella \\Oscar Barriga-Cabanillas \\Juan Correa }
 
-% {Oscar Barriga-Cabanillas\inst{1} \\
-% (with Travis Lybbert\inst{2})}
-
-\institute[UC Davis] % (optional)
-{
-%  \inst{1}%
-  Agricultural and Resource Economics\\
-  UC Davis
- }
+% \institute[UC Davis] % (optional)
+% {
+% %  \inst{1}%
+%   Agricultural and Resource Economics\\
+%   UC Davis
+%  }
 
@@
 \begin{itemize}
-    \item Struggled to provide an adequate and reliable food supply
+    % \item Struggled to provide an adequate and reliable food supply
     \item Breakthroughs in maize germplasm ($\uparrow$ yields, $\uparrow$ drought tolerance)
@@
 
-\begin{enumerate}
+\begin{enumerate}[ {[}1{]} ]
     \item Market integration
@@
 
-\begin{enumerate}
-    \item Adoption where  net benefits are the highest
-    \item How adaptation strategies augment comparative advantage in adoption
-    \item Drivers of dis-adoption of improved maize seed
+\begin{enumerate}[ {[}1{]} ]
+    \item Explore heterogeneous comparative advantages through GRC: find puzzling no significant results.
+    \item Compare definitions of improved seeds: self-reported, second generation or recycled, purity of genetic material, year of release, types and source.
+    \item Low adoption of newer and purest varieties, which do return in higher yields. Poor farmers paying a premium for genetically diluted improved varieties.
 \end{enumerate}
@@
 \begin{frame}
-\frametitle{Empirical Strategy I}
-
-Correlated Random Coefficient Model and comparative advantage (Suri 2011; Tjernstr\"{o}m et al. 2020; Michler et al. 2019; Barriga et al. 2018)
+\frametitle{Data}
 
 \begin{itemize}
-    \item Farmers face unobservable heterogeneous returns to adoption (comparative advantage)
+    % \item Struggled to provide an adequate and reliable food supply
+    \item We focus on maize producing households 
+    \item Self-reported use of improved seed varieties
+    \item Compare self-reported and cropcut yields
 \end{itemize}
 
+\begin{figure}
+    \centering
+    \includegraphics[scale=.4]{results/figures/map_hhids.pdf}
+\end{figure}
+
+% \textit{Add interesting image}
+\end{frame}
+
+%---------------------------------------------------------
+\begin{frame}
+\frametitle{Empirical Strategy I: Finding the returns from adoption}
+
+\textbf{Understand the returns of adopting improved varieties: }
+\begin{itemize}
+    \item Some farmers have higher returns to adoption (comparative advantage)
+    \item How comparative advantage plays a role on using improved seed varieties.
+\end{itemize}
+
+% (Suri 2011; Tjernstr\"{o}m et al. 2020; Michler et al. 2019; Barriga et al. 2018)
+
+We use a Correlated Random Coefficient Model:
+\vspace{-2em}
 \begin{center}
 $$
-    y_{it}= \delta + \beta h_{it} + \theta_{i} + \phi\theta h_{it} + \tau_{i} + \epsilon_{it}
+    y_{it}= \delta + \Delta h_{it} + \theta_{i} + \phi\theta h_{it} + \tau_{i} + \epsilon_{it}
 $$  
@@
     \item $\tau$:  household’s absolute advantage
-    \item $\theta$: household’s comparative advantage
+    \item \textbf{$\theta$: household’s comparative advantage}
+    \item \textbf{$\Delta$: The marginal effect of adoption}
     \item $\phi$: Gains from adopting, relative to a household’s comparative advantage.
@@
 \begin{frame}
-\frametitle{Empirical Strategy II and Extensions}
+\frametitle{Empirical Strategy I: Finding the returns from adoption}
 
@@
 \begin{enumerate}
-    \item Identification does not rely on the existence of a valid instrumental variable (linear projection of observed adoption history)
-    \item Disentangles a household’s absolute advantage from comparative advantage in adopting. 
+    \item Identification does not rely on a IV 
+    % (linear projection of observed adoption history)
+    \item Disentangles absolute advantage from comparative advantage in adoption. ($\theta_i = \theta_A - \theta_N$)
 \end{enumerate}
 
-Extending the model:
+\input{results/tables/traj}
 
-\begin{itemize}
-    \item Implementing a nonparametric panel identification in a novel GRC (Group Random Coefficients)
-    \item Modeling and incorporating time-varying characteristics 
-    \item Complement ESS data with precipitation (CHIRPS) and temperature (CPC) data
-\end{itemize}  
+% Extending the model:
+
+% \begin{itemize}
+%     \item Implementing a nonparametric panel identification in a novel GRC (Group Random Coefficients)
+%     \item Modeling and incorporating time-varying characteristics 
+%     \item Complement ESS data with precipitation (CHIRPS) and temperature (CPC) data
+% \end{itemize}  
 \end{frame}
@@
 \begin{frame}
-\frametitle{Group Random Coefficients Results}
+\frametitle{Empirical Strategy II: Returns to Different Types of Hybrid}
+% \frametitle{Group Random Coefficients Results}
+\textbf{Why farmers do \textit{(not)} adopt?}
+
+\begin{enumerate}
+    \item There are negative returns from adoption
+    \item Comparative advantage is close to 0 unless you always adopt (perhaps driven by 
+\end{enumerate}
+
+\begin{figure}
+    \centering
+    \label{fig:traj_sum}
+    \includegraphics[scale=.4]{results/figures/traj_sum.png}
+\end{figure}
 
@@
 
+%---------------------------------------------------------
+
+%---------------------------------------------------------
+\begin{frame}
+\frametitle{Endogenous Switching Regression Results}
+
+Why do we see farmers adopting despite the seemingly small negative returns?
+
+\begin{itemize}
+    \item We use the DNA-fingerprinting to shed light on this puzzle
+    \item DNA allows us to assess what varieties (and purity) farmers are using 
+    \item Unfortunately, only available in the last round
+\end{itemize}
+
+We implement a switching regression approach
+
+\end{frame}
 %---------------------------------------------------------
@@
 \frametitle{Discussion}
+Negative and small returns to adoption from self-reported improved varieties.
+\begin{itemize}
+    \item But positive returns to more genetically pure, DTMZ, CG- sourced and newer varieties.
+    \item All DNA fingerprints are 70\% pure or higher. Control group has high share of false positives: no proper control.
+
+\end{itemize}
+Negative GRC results mask a large heterogeneity in returns to adoption. 
+\begin{itemize}
+    \item Low purchase of newly bred hybrids, diluted DNA, low inputs use.
+
+\end{itemize}
+
+Effects of high-yielding varieties diminishes over time, but traditional varieties are well adapted to local agro-ecological conditions.
 
 \begin{itemize}
-    \item Negative returns to adoption from self-reported improved varieties.
-    \item But positive returns to more genetically pure, DTMZ, CG- sourced and newer varieties.
-    \item All DNA fingerprints are 70\% pure or higher. Control group has high share of false positives: no proper control. 
-    \item Negative results from GRC mask a large heterogeneity in returns to adoption.
+    \item There may be a threshold in time above which it may be more profitable to use a traditional variety than an improved one that has been around for too many generations.
+
 \end{itemize}
@@
 %---------------------------------------------------------
+\begin{frame}{}
+    Thank you!
+\end{frame}
+
+\begin{frame}{Raw Trajectories}
+    \begin{figure}
+        \centering
+        \includegraphics[scale=.5]{results/figures/theta.png}
+    \end{figure}
+\end{frame}
+
+\bibliography{references}
 

### 2022-07-21 paper.tex
Updates from Overleaf

@@
 Figure \ref{fig:theta_delta_raw} shows a plot of the comparative advantages and returns for the cropcut model with controls.
-We can see here that returns are very high for the $(0,0,1)$ trajectory, but with an exceedingly negative comparative advantage.
-Conversely, all trajectories that adopted in the first wave have a positive comparative advantage, but negative returns.
-This is indicative of either negative sorting (consistent with the negative $\phi$ parameter that we found with the weak identification results) or with a systematic problem in adoption in the first year.
+For the most part these results highlight two interesting puzzles: that it would seem that for many trajectories, comparative advantage is either very close to 0, implying that there is no difference between farmers that adopt and not adopt, and that for trajectory $(0,1,1)$, it would seem that comparative advantage is exceedingly negative.
+This is peculiar, as it this would imply that farmers that adopted in both waves 2 and 3, would actually be better off not adopting, even though their returns are positive (as shown in the bottom figure).
+Another peculiarity in this figure is that it would seem that if we compare trajectories that only differ by one wave's adoption, comparative advantage and returns seem to fall.
+For example, the transition from $(0,1,0)$ to $(0,1,1)$ leads to a large fall in comparative advantage.
+The same can be seen less starkly for the move from $(1,0,0)$ to $(1,0,1)$ and $(1,1,0)$ to $(1,1,1)$.
+But despite this loss in comparative advantage, returns seem to increase.
+This seems to imply that adopting more often leads to higher comparative advantage, but slightly higher returns.
 
@@
 
-This is illustrated in Figure \ref{fig:when_adopt}, where we calculate an aggregation for comparative advantage and returns whether a household adopted in each wave.
-We can see here that adopting in the first wave is consistent with a positive comparative advantage, but negative returns.
-This is repeated, but less strongly in wave 2.
-This means that for waves 1 and 2, households have gains to adopting, but actually receive negative returns to doing so.
-This is indicative of the fact that there are barriers to those that adopted in waves 1 and 2, or perhaps, that returns from improved seed are not as high as expected.
-This relationship is reversed in the third wave, albeit both comparative advantage and returns are close to 0.
-Figure \ref{fig:when_adopt} provides a curious trend across time to adoption: it would seem that negative returns to adoption reduce across time, but adoption-specific advantages in households also get smaller.
+To explore this further, we run a similar regression to calculate the comparative advantage of adopting a certain number of times.
+This is similar to Eq.
+\ref{eq:GRC_Suri}, but uses the number of times that a household adopted hybrid maize (here, $NAdopt$), not its trajectory.
 
-\begin{figure}
-    \caption{Aggregation Results: Comparative Advantage and Returns by Wave of Adoption (Dry Cropcuts)}
-    \label{fig:when_adopt}
-    \centering
-    \includegraphics[scale=0.75]{results/figures/when_adopt_cropcut.png}
-    \vspace*{-2em}
-    \begin{table}[H]
-        \centering
-        \begin{tabular}{p{0.9\textwidth}} 
-            \begin{tablenotes}
-                  \small
-                  \item Note: Author's calculations using ESS.
-The figures shows the aggregate impact from Table \ref{tbl:unres} using on which period a person adopted as the basis for the aggregation.
-            \end{tablenotes}
-        \end{tabular}
-    \end{table}
-\end{figure}
+\begin{align}
+y_{it}&=\sum_{\underline{na}\in\mathcal{NA}\backslash 3}\nu_{\underline{h}}+\Delta_{(0,0,1)}h_{it}+\phi(\nu_{(0,0,1)}-\nu_{(0,0,1)})_{it}1\{NAdopt_{i}=1\}\nonumber\\
+&~~~+\left(\nu_{(1,1,1)}+\phi\left(\nu_{(1,1,1)}-\nu_{(0,0,1)}\right)\right)h_{it}1\{NAdopt_{i}=3\} + X_{it}\beta +\varepsilon_{it},\label{eq:GRC_NAdopt}
+\end{align}
 
-Figure \ref{fig:traj_sum} provides another aggregation of returns and comparative advantage using the times a household adopted hybrid maize.
-Adopting 0 times leads to negative comparative and advantage and negative returns makes sense.
-Those that do not adopt would not receive benefits from adopting.
-However, those that adopt once during waves 1-3, see a positive comparative advantage, but negative returns.
-This same is also starkly true for those that adopted three times.
-For those that adopted twice, it would seem that comparative advantage is negative and there are negative returns to adoption.
-The lack of a pattern in Figure \ref{when_adopt} seems to suggest that the number of times adopting is not as conclusive as \emph{when} households adopted.
-Specifically, later adoption periods lead to smaller negative comparative advantages and less negative returns.
-This suggests that the value of adopting is getting higher, and that this may be due to the introduction of newer and more effective seed.
+In this, instead of a set of trajectories, we have a set of the number of times adopted $\mathcal{NA}$.
+$\nu$ denotes the average yield from adoption $na$ times.
+We use adopting once as the base.
+Figure \ref{fig:n_adopt} shows the results of the regression and calculation of comparative advantage.
+As expected, adopting 0 times leads to negative comparative advantage, but with positive returns to adoption, implying that there is value to adopting to those that have never adopted hybrid maize seed.
+Adopting once leads to both positive comparative advantage, and positive returns to adoption.
+However, adopting twice actually leads to negative comparative advantage, and less positive returns.
+Adopting all three times leads to positive comparative advantage, but negative returns.
 
@@
     \caption{Aggregation Results: Comparative Advantage and Returns by the Number of Times Adopted (Dry Cropcuts)}
-    \label{fig:traj_sum}
-    \includegraphics[scale=0.75]{results/figures/traj_sum.png}
+    \label{fig:n_adopt}
+    \includegraphics[scale=0.75]{results/figures/num_adoption.png}
     \vspace*{-2em}
@@
 \input{results/tables/switch3}
+
+\subsection{Misclassification-Robust GRC}
+
+The previous results highlight the fact that the definition of hybrid matters.
+Even when households think they are not adopting hybrid maize seed, there is such a large amount of hybrid seed that has been introduced over the years in Ethiopia that farmers may be misclassifying their seed and taking part in sub-optimal growing strategies as a result.
+For instance, if a farmer thinks they are growing tradition seed, but they are actually growing hybrid seed, then they will tend to have low returns; hybrid seeds ofte require more fertilizer and other inputs to work optimally.
+If they are cultivated like traditional seeds, then their performance may end up being worse.
+
+Misclassification of seed is not just a problem in Ethiopia.
+Countries have been subsidizing inputs and seed for decades and many countries might have the same challenges.
+In this section, we will introduce a misclassification-robust approach to the GRC method that will allow us to take misclassification into account when calculating comparative advantage.
+The results from the previous section highlight that there are many definitions to hybrid seed and the fourth wave of the ESS includes DNA fingerprinting.
+This gives us the ability to validate how often a hybrid seed is actually used when a farmers self-reports adopting hybrid seed.
+The methodology in this section builds on work that was done in \cite{michuda2021three}, which used confusion matrices from machine learning algorithms as a way to estimate an causal effect that was robust to misclassification.
+Although, the ESS data does not have results from machine learning algorithms, we can construct a confusion matrix based on the DNA fingerprinting data in the fourth wave and apply it to the previous three waves.
 

### 2022-07-22 results/tables/misclass_95purity.tex
add misclassification matrix

\begin{table}
\caption{Misclassification Matrix for Seed Adoption based on DNA Fingerprinting}
\label{tab:cm}
\begin{tabular}{lllllllll}
\toprule
 & 000 & 001 & 010 & 011 & 100 & 101 & 110 & 111 \\
\midrule
000 & 0.642736 & 0.102033 & 0.102033 & 0.016198 & 0.102033 & 0.016198 & 0.016198 & 0.002571 \\
001 & 0.220303 & 0.524466 & 0.034973 & 0.083258 & 0.034973 & 0.083258 & 0.005552 & 0.013217 \\
010 & 0.220303 & 0.034973 & 0.524466 & 0.083258 & 0.034973 & 0.005552 & 0.083258 & 0.013217 \\
011 & 0.075510 & 0.179765 & 0.179765 & 0.427960 & 0.011987 & 0.028537 & 0.028537 & 0.067938 \\
100 & 0.220303 & 0.034973 & 0.034973 & 0.005552 & 0.524466 & 0.083258 & 0.083258 & 0.013217 \\
101 & 0.075510 & 0.179765 & 0.011987 & 0.028537 & 0.179765 & 0.427960 & 0.028537 & 0.067938 \\
110 & 0.075510 & 0.011987 & 0.179765 & 0.028537 & 0.179765 & 0.028537 & 0.427960 & 0.067938 \\
111 & 0.025882 & 0.061616 & 0.061616 & 0.146687 & 0.061616 & 0.146687 & 0.146687 & 0.349211 \\
\bottomrule
\end{tabular}
\end{table}

### 2022-07-22 paper.tex
Updates from Overleaf

@@
 \doublespacing
+\usepackage{subcaption}
 \usepackage{subfig}
 \usepackage{epigraph}
-
 \usepackage{enumerate}
@@
 
+\begin{table}[]
+    \begin{subtable}{.5\linewidth}
+      \centering
+        \caption{Hybrid DNA Purity $>$ 95\% \label{tab:a}}
+        \begin{tabular}{lll}
+        \toprule
+         & 0 & 1 \\
+        \midrule
+        0 & 0.863 & 0.137 \\
+        1 & 0.296 & 0.704 \\
+        \bottomrule
+        \end{tabular}
+        \vspace{1cm}
+    \end{subtable}%
+    \begin{subtable}{.5\linewidth}
+      \centering
+        \caption{DNA Fingerprinting for Drought-Tolerant Maize \label{tab:b}}
+        \begin{tabular}{lll}
+        \toprule
+         & 0 & 1 \\
+        \midrule
+        0 & 0.971 & 0.029 \\
+        1 & 0.868 & 0.132 \\
+        \bottomrule
+        \end{tabular}
+        \vspace{1cm}
+    \end{subtable} 
+    \begin{subtable}{\linewidth}
+        \centering
+        \caption{DNA Fingerprinting of Seed Released after 2010 \label{tab:c}}
+        \begin{tabular}{lll}
+        \toprule
+         & 0 & 1 \\
+        \midrule
+        0 & 0.859 & 0.141 \\
+        1 & 0.628 & 0.372 \\
+        \bottomrule
+        \end{tabular}
+    \end{subtable}
+    \caption{Misclassification Tables for Farmers' Adoption of Hybrid Seed \label{tab:misclassification}}
+    \begin{tablenotes}
+    \footnotesize
+    \item{Note: Table shows three different definitions for misclassifying adoption of hybrid maize.
+A 0(1)-row denotes that a farmer self-reported using traditional(hybrid) seed.
+A 0(1)-column is whether the seed was in fact traditional(hybrid).
+All definitions calculated from ESS data.
+Sub-table \ref{tab:a} shows the percentage that farmers self-reported using hybrid seed when that seed was at least 95\% pure.
+Sub-table \ref{tab:b} shows the same table but for whether the seed was drought-tolerant maize seed.
+Sub-table \ref{tab:c} shows the same table but for hybrid seed released after 2010.}
+    \end{tablenotes}
+\end{table}
+
+Table \ref{tab:misclassification} shows a table of different hybrid seed definitions and the percentage of times that a farmer self-reported traditional or hybrid seed, and whether they in fact were using traditional or hybrid seed, based on the DNA fingerprinting data.
+You can see from Sub-table \ref{tab:a} that around 86\% of farmers actually use traditional seed when they self-report using traditional seed, and around 70\% of the time, farmers use hybrid seed of at least 95\% purity when they self-report using hybrid.
+There are in fact instances where farmers, however, misclassify their self-reports.
+For instance, in sub-table \ref{tab:b}, a farmer that self-reports using drought-tolerant maize, only in fact uses drought-tolerant maize 13\% of the time.
+There are similarly large misclassifications for seed released after 2010.
+The previous section highlighted that different definition of maize could actually lead to vastly different treatment impacts from seed adoption.
+
+In order to apply these misclassification probabilities to the previous three waves, we made assumptions about the nature of misclassification: that the rate of misclassification is constant across time and exhibits no dynamic behavior.
+Although these are strong assumptions, they come from the limitations of the data; the fourth wave of the ESS was the first to include DNA fingerprinting, and so was the first where a table like the one in Table \ref{tab:misclassification} could be calculated.
+Let $p_{ij}$ be the $i,j$th cell of a misclassification table as in Table \ref{tab:misclassification}.
+Suppose that we construct the probability of being in a true trajectory $h = (i,j,k)$, intersected with being in a reported trajectory $\hat{h} = (\hat{i},\hat{j}, \hat{k})$ as:
+
+\begin{equation}
+\label{eq:misclass}
+P(h = (i,j,k) , \hat{h} = (\hat{i},\hat{j}, \hat{k})) = p_{i, \hat{i}}\cdot p_{j, \hat{j}}\cdot p_{k, \hat{k}}
+\end{equation}
+
+We can use Eq.
+\ref{eq:misclass} to construct the joint distribution, $f(h, \hat{h})$ for each trajectory in the sample.
+Since we had the most confidence in the data from Table \ref{tab:a}, we decided to use it for the main body of the paper.
+
 \section{Conclusion}\label{sec:conclusion}
@@
 Accounting for the decay of the purity of the genetic material, its source and years upon release, reveals that those farmers that rely on newly bred improved varieties do show positive returns to adoption.
-Even without explicit DNA fingerprinting evidence, the GRC was able to show this trend with the comparative advantage results; returns are negative to those that have a high comparative advantage to adopting, but these returns less negative with time.
+Even without explicit DNA fingerprinting evidence, the GRC could show this trend with the comparative advantage results; returns are harmful to those with a high comparative advantage to adopt, but these returns are less negative with time.
 We speculate that this is due to the fact that new and more genetically pure hybrids are introduced giving those that adopt higher yields than before.
@@
 This may be partially driven by the low-purchasing rates of newly bred varieties, but also by the under-use of critical inputs needed for the successful performance of hybrid seeds, such as fertilizers and irrigation, especially in drought-prone areas.
-More research is needed to understand the barriers to adoption and input use.
-Such barriers may be crucial for households with less resources, which might be more likely to reuse improved seeds for many generations or buy older-generations from other farmers at a cheaper price.
+More research is required in order to understand the barriers to adoption and input use.
+Such barriers may be crucial for households with less resources, which might be more likely to reuse improved seeds for many generations or buy older generations from other farmers at a cheaper price.
 The fact that all the sampled plots chosen for DNA fingerprinting showed purity percentages of 70 percent or higher is indicative that this might be the case, given the widespread dilution of improved genetic material.
@@
 These contexts include but are not limited to places subject to high weather variability or where low soil fertility is prevalent due to the use of excessive intensification and chemical input dependence.
-Resilience enhancing strategies such as the use of drought-tolerant maize or irrigation use may be critical for such contexts, but their suitability for the heterogeneous pool of farmers needs to be better understood.
+Resilience-enhancing strategies such as the use of drought-tolerant maize or irrigation use may be critical for such contexts, but their suitability for the heterogeneous pool of farmers needs to be better understood.
 

### 2022-11-01 results/tables/misclass_95purity.tex
change format of misclass table
Updates from Overleaf

@@
 \label{tab:cm}
+\centering
 \begin{tabular}{lllllllll}
@@
 \midrule
-000 & 0.642736 & 0.102033 & 0.102033 & 0.016198 & 0.102033 & 0.016198 & 0.016198 & 0.002571 \\
-001 & 0.220303 & 0.524466 & 0.034973 & 0.083258 & 0.034973 & 0.083258 & 0.005552 & 0.013217 \\
-010 & 0.220303 & 0.034973 & 0.524466 & 0.083258 & 0.034973 & 0.005552 & 0.083258 & 0.013217 \\
-011 & 0.075510 & 0.179765 & 0.179765 & 0.427960 & 0.011987 & 0.028537 & 0.028537 & 0.067938 \\
-100 & 0.220303 & 0.034973 & 0.034973 & 0.005552 & 0.524466 & 0.083258 & 0.083258 & 0.013217 \\
-101 & 0.075510 & 0.179765 & 0.011987 & 0.028537 & 0.179765 & 0.427960 & 0.028537 & 0.067938 \\
-110 & 0.075510 & 0.011987 & 0.179765 & 0.028537 & 0.179765 & 0.028537 & 0.427960 & 0.067938 \\
-111 & 0.025882 & 0.061616 & 0.061616 & 0.146687 & 0.061616 & 0.146687 & 0.146687 & 0.349211 \\
+000 & 0.643 & 0.102 & 0.102 & 0.016 & 0.102 & 0.016 & 0.016 & 0.003 \\
+001 & 0.220 & 0.524 & 0.035 & 0.083 & 0.035 & 0.083 & 0.006 & 0.013 \\
+010 & 0.220 & 0.035 & 0.524 & 0.083 & 0.035 & 0.006 & 0.083 & 0.013 \\
+011 & 0.076 & 0.180 & 0.180 & 0.428 & 0.012 & 0.029 & 0.029 & 0.068 \\
+100 & 0.220 & 0.035 & 0.035 & 0.006 & 0.524 & 0.083 & 0.083 & 0.013 \\
+101 & 0.076 & 0.180 & 0.012 & 0.029 & 0.180 & 0.428 & 0.029 & 0.068 \\
+110 & 0.076 & 0.012 & 0.180 & 0.029 & 0.180 & 0.029 & 0.428 & 0.068 \\
+111 & 0.026 & 0.062 & 0.062 & 0.147 & 0.062 & 0.147 & 0.147 & 0.349 \\
 \bottomrule
 \end{tabular}
+\begin{tablenotes}
+    \item \footnotesize Notes: Table shows results of misclassification probabilities for each trajectory in the sample, using probabilities from Table \ref{tab:a}.
+Misclassification probabilities calculated from DNA fingerprinting data from Wave 4 of the ESS and assumes that true adoption of hybrid seed is when DNA purity of hybrid is at least 95\%.
+\end{tablenotes}
 \end{table}

### 2022-11-01 results/tables/switch1.tex
Updates from Overleaf

@@
 \textit{SR1, SR yields, full set of controls}&            &            &            &            &            \\
-ATT         &        7.28&        7.82&       -0.54&       0.019&      0.0000\\
+ATT         &        7.28&        7.81&       -0.53&       0.019&      0.0000\\
 %
@@
 %
-ATU         &        9.65&        6.85&        2.80&       0.100&      0.0000\\
+ATU         &        9.59&        6.87&        2.72&       0.100&      0.0000\\
 %
@@
 \textit{SR1, cropcut yields, full set of controls}&            &            &            &            &            \\
-ATT         &        7.56&        7.61&      -0.049&       0.019&      0.0087\\
+ATT         &        7.56&        7.62&      -0.051&       0.020&      0.0122\\
 %
@@
 %
-ATU         &        9.24&        7.11&        2.13&       0.085&      0.0000\\
+ATU         &        8.93&        7.13&        1.80&       0.081&      0.0000\\
 %
@@
 \textit{SR2, SR yields}&            &            &            &            &            \\
-ATT         &        7.30&        7.85&       -0.55&       0.020&      0.0000\\
+ATT         &        7.30&        7.91&       -0.61&       0.021&      0.0000\\
 %
@@
 %
-ATU         &        9.63&        6.88&        2.75&       0.101&      0.0000\\
+ATU         &        9.61&        6.90&        2.71&       0.099&      0.0000\\
 %
@@
 \textit{SR2, cropcut yields}&            &            &            &            &            \\
-ATT         &        7.61&        7.62&      -0.013&       0.019&        0.47\\
+ATT         &        7.61&        7.66&      -0.049&       0.019&       0.010\\
 %
@@
 %
-ATU         &        9.35&        7.14&        2.21&       0.091&    1.6e-130\\
+ATU         &        9.32&        7.17&        2.15&       0.091&    1.2e-124\\
 \hline

### 2022-11-01 paper.tex

@@
     \begin{titlepage}
-    \title{A Case of Genetic Dilution in Improved Maize: Estimating Comparative Advantage of Adoption in Ethiopia}
+
+    \title{Genetic Dilution erodes productivity: Exploring farmers' low adoption levels of improved maize in  Ethiopia}
 % This second title is my preferred title at this point (28 Oct 2021). Open to other titles, but it shouldn't be longer than this one!
@@
 			\noindent  
-			Despite large investments in breeding and making high-yielding maize varieties available, evidence shows that few farmers consistently use them over time.
+			Despite large investments in  developing new  maize varieties, evidence shows that few farmers consistently adopt this new technology over time.
 We use an innovative methodology that studies the adoption trajectories of farmers over a three-round panel data set to explain low adoption rates with heterogeneous returns to adoption.
-By relying on self-reported use of improved seeds, we find that farmers do not adopt because they have a low comparative advantage and seemingly small to negative returns to adoption.
-To reconcile this result, we exploit unique information on maize DNA-fingerprinting collected over the same areas in 2018.
-We find that the negative returns to adoption mask heterogeneity driven by the variety of improved seeds used.
+By relying on self-reported use of improved seeds, we find a puzzling answer for why farmers do not use new seed varieties: adoption leads to seemingly small and negative returns.
+Heterogeneity analysis across different levels of comparative advantage show that farmers who adopt don't actually have a comparative advantage in adopting improved seed, nor do they exhibit high returns to adopting.
+To reconcile this result with the expected higher yields for improved seeds, we exploit unique information on maize DNA-fingerprinting collected over the same areas in 2018.
+We find that the negative returns to adoption mask heterogeneity in the actual quality of the genetic material of the seeds farmers categorized as improved varieties in the self-reported data.
 Positive returns to adoption are found for those farmers that use higher-purity germplasm, drought-tolerant maize, and newly released varieties.
-Our findings point to the wide dispersion of older and genetically diluted varieties, for which poorer farmers are paying a premium.
-The implications of our findings speak to the need for policies to better target context and geography, expand accessibility of improved seeds, and make higher yields more inclusive.
+When we account for misclassification in our heterogeneity analysis, we find that the puzzle is resolved.
+Our findings point to the wide dispersion of older and genetically diluted varieties, for which poorer farmers may be paying a premium.
+The implications of our findings speak to the need for policies to better target context and geography, expand accessibility of improved seeds, and make higher yields varieties more inclusive.
 
@@
 Since the drought of 1984, Ethiopia has become increasingly reliant on maize, and as of today, maize constitutes the most widespread crop in the country.\footnote{FAO statistics show the total area cultivated with Maize in 2019 is 25\% larger than the next most widespread crop (sorghum), and 27\% and 139\% larger than the next two following crops, wheat and barley, respectively.} As Figure \ref{fig:maize_yields} shows, there is an upward trend in the total area cultivated with maize since 1990, with average yields almost doubling during this period.
-This increase in maize yields has also coincided with a dramatic decrease in undernourishment which has fallen from 47\% of the population in 2001 to 14\% in 2018 (The World Bank, 2021).
-Despite these positive gains, food insecurity still affects more than half the population of Ethiopia (ibid), reflecting that the progress achieved
-is not necessarily homogeneous across all subpopulations.\footnote{The country also has recently seen increases in food insecurity because of COVID-19, the conflict in Ethiopia’s Tigray region and the desert locust outbreak.}
+This increase in maize yields has coincided with a dramatic decrease in undernourishment of the population which has fallen from 47\% in 1990 to 14\% in 2018 (The World Bank, 2021).
+But despite these positive gains in total area planted and average yields, food insecurity still affects more than half the population of Ethiopia.\footnote{The country has also recently seen increases in food insecurity because of COVID-19, the conflict in Ethiopia’s Tigray region and the desert locust outbreak.} Moreover, a large productivity gap persists between the African continent and the rest of the developing world.
+The persistence of this gap is often attributed to different adoption rates of green revolution technologies, which include improved seeds.
+Evidence has shown that the adoption of improved seed leads to increases in agricultural yields \citep{Carter2014-fm}, food security \citep{Shiferaw2014-op}, and poverty reduction \citep{Minten2008-tj}.\footnote{Nevertheless, critics of green revolution technologies have also claimed that these technological packages have not been as profitable for many farmers in Africa as promised (with the exception of Ethiopia), given the high costs, and their heavy dependence on subsidies \citep{Wise20}.}
 
@@
 %%%%% The puzzle
-The large disparities in yields and food security between the African continent and the rest of the developing world are often attributed to different adoption rates of green revolution technologies, such as improved seeds.
-Evidence has shown that the adoption of improved seed leads to increases in agricultural yields \citep{Carter2014-fm}, food security \citep{Shiferaw2014-op}, and poverty reduction \citep{Minten2008-tj}.
-Nevertheless, critics of green revolution technologies have also claimed that these technological packages have not been as profitable for many farmers in Africa as promised (with the exception of Ethiopia), given the high costs, and their heavy dependence on input subsidies \citep{Wise20}.
-\par
-Even if the net returns to adoption of improved seeds are more conservative than often thought, the evidence regarding increased yields over traditional varieties is consistent across multiple contexts.
-Despite this evidence and numerous technological breakthroughs in maize germplasm with increased yields and drought tolerance, in the period between 2011 and 2015 only 13\% of maize farmers in Ethiopia have adopted improved seeds consistently, a large percentage has adopted new varieties inconsistently, and the lion share of farmers (62\%) have not used improved seeds at all during the same period.
 
+Despite evidence and numerous technological breakthroughs in maize germplasm with increased yields and drought tolerance, only 13\% of maize farmers in Ethiopia have adopted improved seeds consistently between 2011 and 2015.
+A large percentage has adopted new varieties inconsistently, and the largest share of farmers (62\%) have not used improved seeds at all during the same period.
 Several answers have been proposed to explain such low adoption of improved varieties.
-These include mainly the lack of liquidity and other economic constraints \citep{Carter2014-fm}, growing conditions such as the type of soil, input use, and farmer characteristics \citep{Munshi2004-og}, imperfections in credit markets \citep{Croppenstedt2003-pq}, ambiguously defined property rights \citep{Place2000-el}, lack of commitment devices \cite{Duflo2009-iv}, high transportation costs that increase the cost of agricultural inputs \citep{Byerlee2013-qk}, and improved seed supply constraints \citep{Bird2020-nt}.
-In this context, there is still a large percentage of the population not using improved seed varieties, or using varieties that are over 20 years old.
-With such old seed, there is evidence that yield improvement and disease resistance in these seeds are waning\citep{Abate2015-rj}.
+These include lack of liquidity and other economic constraints \citep{Carter2014-fm}, growing conditions such as the type of soil, input use, and farmer characteristics \citep{Munshi2004-og}, imperfections in credit markets \citep{Croppenstedt2003-pq}, ambiguously defined property rights \citep{Place2000-el}, lack of commitment devices \cite{Duflo2009-iv}, high transportation costs that increase the cost of agricultural inputs \citep{Byerlee2013-qk}, and improved seed supply constraints \citep{Bird2020-nt}.
+Additionally, the net returns to adoption of improved seeds are often lower than expected in the long term because a large percentage of farmers unknowingly use diluted improved seed varieties, or ones that are over 20 years old.
+With such old seed, there is evidence that yield improvement and disease resistance wanes over time\citep{Abate2015-rj}.
 The variety of these constraints and their dispersion across subpopulations and locations speaks to considerable heterogeneous returns to adoption.
@@
 In this paper, we revisit the mechanism proposed by \citep{Suri2011-oi}, where a lack of adoption of improved seed varieties is the result of farmers' heterogeneous returns to the new technology.
-Ignoring such heterogeneity can hide the true impact of adoption, as different sub-populations' returns may be impacted by the same technology in different, and even opposite ways, despite it providing high average returns.
+Ignoring such heterogeneity hides the true impact of adoption, as farmers with different characteristics and skill levels observe different returns from adoption despite the technology providing high average returns.
 The heterogeneous returns to adoption are modeled as a farmer's unobservable comparative advantage using a Correlated Random Coefficient (CRC) model.
-In the original paper, \citep{Suri2011-oi} finds that farmers’ lack of adoption is driven by insufficient net benefits to the technology stemming from poor infrastructure that increase the cost of access.
-Similar work in Ethiopia on improved chickpea varieties suggests that the economic returns to adoption plays a major role in the decision to adopt \citep{Michler2018-wk}.
+In the original paper, \citep{Suri2011-oi} finds that farmers’ lack of adoption is driven by insufficient net benefits to the technology stemming from poor infrastructure that increases the cost of access.
+\footnote{Similar work in Ethiopia on improved chickpea varieties suggests that the economic returns to adoption plays a major role in the decision to adopt.
+\citep{Michler2018-wk}.}
 
 %% Expand a bit on the methods
-We build on this literature by investigating hybrid maize seed adoption in Ethiopia, focusing on how heterogeneous returns drive the decision to use improved seed varieties.
-We implement two complementary estimation strategies.
-First, we implement a Group Random Coefficients estimation, which builds on \cite{Suri2011-oi} in \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} using a General Method of Moments (GMM) Estimator that allows us to test the assumptions assumed by \cite{Suri2011-oi}, as well as provide more economically interpretable results.
-The advantage of our estimation strategy is that we do not rely on identification from a valid instrument, but a strategy similar to \cite{Chamberlain1984-uk}.
-For this strategy, we rely on the first three survey rounds of the Ethiopian Socioeconomic Survey (ESS), panel dataset that allows us to follow the same households across time (from 2011 to 2015).
-Using self-reported information on the seed variety (improved or traditional), we compare the adoption trajectories over three rounds.
-The results show that the returns to adoption are in most cases, not statistically significant, and only significant with wave 3 adoption.
+We build on this literature by investigating improved maize seed adoption in Ethiopia.
+In particular, we implement two complementary estimation strategies to address both how heterogeneous returns drive the decision to use improved seed varieties, and the extent that the misclassification of varieties and genetic dilution affect the decision to adopt them by lowering their observed returns.
+First, we implement a Group Random Coefficients estimation, which builds on \cite{Suri2011-oi} in \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} using a General Method of Moments (GMM) Estimator that allows us to test \textbf{ the assumptions made by \cite{Suri2011-oi}}, as well as provide more economically interpretable results.
+The advantage of our estimation strategy is that we do not rely on identification from a valid instrument.
+%\textbf{to assess XXXXXX}.
+%\footnote{Our estimation strategy is similar to \cite{Chamberlain1984-uk} \textbf{XXXXXX}.}
+
+For the first set of results, our empirical strategy relies on the first three survey rounds of the Ethiopian Socioeconomic Survey (ESS).
+The data were collected as a panel, which cover 2011, 2013, and 2015.
+For each round, farmers self-reported on the seed variety they used (improved or traditional).
+By using the adoption trajectories over three rounds, we find the returns to adoption are in most cases, not statistically significant, and only significant with wave 3 adoption.
 This posits a puzzle that motivates our next estimation strategy.
 
-We expand our results using the fourth round of the ESS (conducted in 2018), which provides a unique opportunity to understand why improved seed varieties seem not to increase returns when adopted.
+We expand our results using the fourth round of the ESS (conducted in 2018), which provides
+a unique opportunity to understand why improved seed varieties seem not to increase returns
+when adopted.
 The ESS4 collected DNA fingerprints of harvested cropcuts in selected plots and compared its genetic material with a reference library of available varieties in Ethiopia.
-This provides critical information not only about the seed types but also about their purity, sources, and their year of release.
+This data provides critical information not only about the seed types but also about their purity, sources, and their year of release.
 We exploit this uniquely rich information to compare different definitions of improved seeds.
-Specifically, we compare the returns to the self-reported use of new, second generation or recycled varieties, with those of improved seeds as defined by their genetic material, comparing different purity thresholds (or degrees of no-contamination with other varieties), year of release, types of varieties, and their sources.
+Specifically, we compare returns of verifiably improved seeds against self-reported use of new, second generation or recycled varieties, comparing different purity thresholds (or degrees of no-contamination with other varieties), year of release, types of varieties, and their sources.
 Because the latest survey round is not a continuation of the first three rounds, our second estimation strategy implements an Endogenous Switching Regression (ESR) approach, a standard estimation strategy to study the question of adoption with cross-sectional data (\citealt{Marenya2020-kb}, \citealt{Shiferaw2014-op}), given its capacity to account for unobserved heterogeneity in adoption.
-
-The results of the ESR find negative effects of adopting improved seeds, as per farmer self-reports.
-But the comparison of the different improved seed definitions using the DNA fingerprinting sheds light on possible explanations for this apparent puzzle.
-We find that when the definition of improved seed is based on greater purity of the germplasm (95\% of purity or more), the returns to adoption for those that used such improved varieties becomes positive and significant.
+The results of the ESR show negative effects of adopting improved seeds, as per farmer self-reports.
+But the comparison of different improved seed definitions using the DNA fingerprinting shows that when the definition of improved seed is based on greater purity of the germplasm (95\% of purity or more), the returns to adoption for those that used such improved varieties becomes positive and significant.
 We obtain similar results when evaluating the specific effects of adopting drought-tolerant maize varieties, CGIAR or Exotic varieties, or more recently released varieties.
 
-Taken together, the negative and not statistically significant returns to adoption we find from both estimation strategies reflect the loose definition of improved seeds when such improved varieties are self-reported by farmers.
-Only a subset of farmers use the purest genetic varieties, drought-tolerant varieties, or newly released varieties, which tend to provide higher returns.
-This implies that poorer farmers are paying a premium for genetically diluted hybrids, which has far-reaching policy implications about the need to foster a more inclusive policy when making the best performing seeds available to all farmers.
+Taken together, these results offer a plausible answer to the puzzle of low adoption levels of new varieties.
+The negative and not statistically significant returns to adoption we find from both estimation strategies reflect the loose definition of improved seeds when such improved varieties are self-reported by farmers.
+Only a subset of farmers use the purest genetic varieties, drought-tolerant varieties, or newly released varieties, which provide the highest returns.
+This implies that poorer farmers are paying a premium for genetically diluted hybrids, which has far-reaching policy implications about the need to foster a more inclusive strategy when making the best performing seeds available to all farmers.
 
-Our results contribute to the literature on the drivers of low adoption rates.
+Finally, to integrate our results together, we use the misclassification rates from the 4th round of the data and re-estimate the GRC results to be robust to misclassification.
+This method takes inspiration from \cite{michuda2021three} to account for the fact that the self-reports from the first three rounds of the ESS data may be misclassified and attenuated.
+We find that once we account for misclassification, heterogeneity in adoptions becomes very similar to other contexts; higher comparative advantage is associated with lower returns pointing to significant barriers to adoption for those that would benefit the most from it.
+Incorporating misclassification rates from the fourth round of the ESS into the GRC estimation resolves the puzzle posed by the GRC results.
+
+Our results contribute to the literature on the drivers of low adoption rates of improved seed varieties.
 The consensus in this literature is that the success of improved varieties depends on a range of factors, and that the widespread adoption of  improved seeds is the result of complementary conditions and inputs that allow farmers to extract the full potential of these varieties.
@@
 
-We contribute to this literature by exploring heterogeneous comparative advantages to adoption in Ethiopia between 2011 to 2015, controlling for important drivers of adoption such as rainfall (found to be the main driver of heterogeneous effects for adopting agronomic packages \citealt{Marenya2020-kb}, \citealt{Katengeza2019-af}), labor and input use.
-Our paper also provides a novel way of visualizing GRC results, which become difficult to interpret with longer time variables.
-Our findings also provide an alternative hypothesis to the low-returns to adoption, namely that the wide dispersion of older generations and recycled improved seeds, which may be both resulting in lower yields and altering farmers' input use in response to a mis-classified seed.
+We contribute to this literature by exploring heterogeneous comparative advantages to adoption in Ethiopia between 2011 to 2015.
+%, controlling for important drivers of adoption such as rainfall (found to be the main driver of heterogeneous effects for adopting agronomic packages \citealt{Marenya2020-kb}, \citealt{Katengeza2019-af}), labor and input use.
+Our findings provide an alternative hypothesis to the puzzle of low maize adoption, namely that the wide dispersion of older generations and recycled improved seeds, may be resulting in lower yields and altering farmers' input use.
+The null and negative effects that we see in the GRC results are actually stemming from misclassification in self-reported adoption status.
+This paper is the first to take seriously the bias that this can create when estimating the treatment effect of adoption, both in a homogeneous and heterogeneous setting.
 This also implies a methodological discovery when investigating hybrid seed adoption: the absence of a true control group to use for analysis.
-With more time and more genetic dilution, it becomes more and more likely that even those that do report hybrid use are still likely using some level of hybrid.
+With more time and more genetic dilution, it becomes more likely that even those that do report hybrid use are still likely using some level of hybrid.
 
-The article continues as follows.
-In Section \ref{sec:data}, we describe the data used, the sample frame, general farmer and farm characteristics, and improved seed adoption levels.
+% A different branch of the literature provides an alternative explanation for the low adoption rates, focusing on heterogeneous potential returns to adoption \citep{Suri2011-oi}.
+This literature suggests that even if returns to adoption are high, farmers may fail to adopt if they face low comparative advantage to adoption.
+Other methods used in the literature fail to account for this heterogeneity, admitting only the presence of time-constant and/or time-varying unobserved heterogeneities (depending on the data structure) for identification purposes \citep{Kassie2018-xn,Falco2011-rt}.
+
+The article continues as follows: in Section \ref{sec:data}, we describe the data used, the sample frame, general farmer and farm characteristics, and improved seed adoption levels.
 Section \ref{sec:strategy} presents the methodological details of our two estimation strategies.
@@
 This is important information as the first three waves of ESS only included self-reported hybrid maize status.
-Since our estimation strategy requires a balanced panel dataset of maize growers, our main analysis includes the first three waves of the ESS.
-We use the fourth wave for validation, and to explore the impact of specific types of improved seeds.
+Since our estimation strategy to calculate returns to adoption requires a balanced panel dataset of maize growers, our main analysis includes the first three waves of the ESS.
+We use the fourth wave to provide further light on the low levels of returns from adoption we find in the first three ESS rounds, and to explore the impact of specific types of improved seeds.
 
@@
 Harvest labor is disproportionately skewed to family participation with around 20 days of harvest labor being conducted by family members and only about 2 days for hired labor.
-This is indicative of either malfunctioning labor markets, income constraints or high supervision costs.
+This is indicative of either \textbf{malfunctioning} labor markets, income constraints or high supervision costs.
 The distances to some key areas are also highlighted in Table \ref{tbl:summary}.
@@
 Fertilizer costs are on average 8,350 Birr ($\thicksim$178 USD), with a large amount of the sample not using fertilizer at all.
-This is particularly important in the context of improved maize seed use as it requires fertilizer to be most effective.
+This is particularly important in the context of improved maize seed use as these technologies are intensive in external inputs, such as chemical fertilizers and irrigation, to be effective.
 Only 5\% of the sample irrigates their crops making households susceptible to weather shocks.
@@
 Self-reported yields are both more susceptible to behavioral biases, but also to confusion caused by differences in assumed units.
-In general, though, the sample with available cropcut yields cuts down the available sample significantly, given that cropcuts were only collected for a subsample of randomly selected households, which limits their power.
+In general, though, the sample with available cropcut yields cuts down the available sample significantly, given that cropcuts were only collected for a subsample of randomly selected households.
 
@@
 
-A crop seed is considered “improved” in Ethiopia if it was tested by breeders and evaluated to be superior to existing varieties  (\citep{MoA13}; \citep{kosmowski2020shining}).
+A seed is considered “improved” in Ethiopia if it was tested by breeders and evaluated to be superior to existing varieties  (\citep{MoA13}; \citep{kosmowski2020shining}).
 In Ethiopia, the government controls most of the seed system, with local agricultural offices or extension agents assessing the demands of different varieties.
-The MoA decides the seed quantities to be produced, and actors such as regional seed enterprises, seed companies, seed unions, are involved in the production of certified seeds, that are primarily distributed by seed producer cooperatives (SPCs), which account for 37\% of the total distribution in the country (\citep{kosmowski2020shining}).
-An improved seed can be obtained directly from such distribution centers (which would distribute a first generation seed), or it could be cultivated from second or later generations from the originally bred seed.
+The MoA decides the seed quantities to be produced, and actors such as regional seed enterprises, seed companies, seed unions, are involved in the production of certified seeds, which are later distributed by seed producer cooperatives (SPCs).\footnote{They account for 37\% of the total distribution in the country (\citep{kosmowski2020shining}).} An improved seed can be obtained directly from such distribution centers (which would distribute a first generation seed), however it is often the case that seeds cultivated from second or later generations from the originally bred seed are also exchanged between farmers.
 In the survey, farmers are asked to report whether the variety they used was a first-generation improved seed, second generation, recycled, or a traditional variety.
@@
 
-Fifty four varieties have been released in Ethiopia since 1990, out of which seven were released between 1990 to 1999, 29 between 2000 to 2009, and 18 between 2010 to 2019.
-Thirty four such varieties contain International Maize and Wheat Improvement Center (CIMMYT) germplasm, out of which, 10 Drought-tolerant maize varieties (DTMZ) and 8 Quality protein maize (QPM) were released (most of these between 2010 to 2019).
-Out of the CIMMYT-related germplasm, ten are open-pollinated varieties (OPVs) and 25 are hybrid (\citep{kosmowski2020shining}).
+Fifty four varieties have been released in Ethiopia since 1990.\footnote{Out of which seven were released between 1990 to 1999, 29 between 2000 to 2009, and 18 between 2010 to 2019} Thirty four such varieties contain International Maize and Wheat Improvement Center (CIMMYT) germplasm, out of which, 10 Drought-tolerant maize varieties (DTMZ) and 8 Quality protein maize (QPM) were released (most of these between 2010 to 2019).
+Besides the CIMMYT-related germplasm, ten are open-pollinated varieties (OPVs) and 25 are hybrid (\citep{kosmowski2020shining}).
 
@@
 Considering a household as adopter if the household planted an improved variety in at least one of their plots, leads to classifying all households where a DNA sample was obtained as users of improved seed variety, a result that holds when the purity threshold of 70\% (DNA 70) is used.
-It is only when using purity thresholds of 90 and 95\% (DNA 90, DNA 95 in Figure \ref{fig:adoption_r4}, respectively), that adoption rates of improved varieties change to 91 and 54\% of sampled households, respectively.
-We also find that 62\% of households adopted a hybrid variety, compared to 38\% that adopted an open-pollinated variety, and 15\% of households adopted a drought-tolerant maize variety, while 63\% of households that adopted an improved variety use a CGIAR-derived germplasm, and 79\% use an exotic germplasm.
-Finally, 64\% of households adopted a variety that was released in the year 2000 or after, and 24\% adopted a variety that was adopted in the year 2010 or after.
+It is only when using purity thresholds of 90 and 95\% (DNA 90, DNA 95 in Figure \ref{fig:adoption_r4}, respectively), that adoption rates of improved varieties change to 91 and 54\% of sampled households, respectively.\footnote{We also find that 62\% of households adopted a hybrid variety, compared to 38\% that adopted an open-pollinated variety, and 15\% of households adopted a drought-tolerant maize variety, while 63\% of households that adopted an improved variety use a CGIAR-derived germplasm, and 79\% use an exotic germplasm.} Finally, 64\% of households adopted a variety that was released in the year 2000 or after, and 24\% adopted a variety that was adopted in the year 2010 or after.
+These ratios are in line with prior findings (\cite{Zeng15}).
 
@@
 
-These ratios are in line with prior findings.
-In the regions of Oromia, Amhara, Tigray, and SNNPR, adoption rates of improved maize varieties of 39\% of households were found, according to self-reported information (\cite{Zeng15}).
+We find regional variations in adoption levels.
+In the regions of Oromia, Amhara, Tigray, and SNNPR, around 39\% of households report adopting an improved varierty .
 In the regions of Amhara, Benishangul-Gumuz, Oromia, SNNPR, and Tigray, adoption rates of improved maize varieties were found for 27\% of households (\cite{Jaleta2018-oj}).
@@
 
-Since ESS1-ESS3 do no have DNA fingerprinting, we are left with only self-reports of hybrid use.
+Since ESS1-ESS3 do not have DNA fingerprinting, we are left with only self-reports of hybrid use.
 We use these variables in a Group Random Coefficiens (GRC) approach to investigate the heterogeneous effects of adoption.
@@
 See \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} for more information on the assumptions involved for the GRC approach.
-In our particular case, we contribute to the literature and methodology of \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} by providing reasonable ways to aggregate the results of the GRC model, which may seem difficult to interpret in its raw form.
+
+% we contribute to the literature and methodology of \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} by providing reasonable ways to aggregate the results of the GRC model, which may seem difficult to interpret in its raw form.
 We use these aggregations to better understand the heterogeneous effects to adoption.
@@
 Table \ref{tbl:Trajectories} maps out the frequencies of the trajectories present in the data, sorted by frequency.
-Adoption trajectories seem to be more popular for later adoptions than earlier ones.
 Most households do not adopt at all (``never-adopters''), but the second most frequent trajectory are those that adopt in every wave (``always-adopters'').
@@
 If $\phi$ is positive, then higher yields are associated with higher comparative advantage, denoting that those that adopt also tend to have the highest returns to adoption.
-If it is negative, however, this denotes that there might be barriers, such as a lack of fertilizer access, and bad roads that restrict the ability to buy hybrid seed.
-$\phi$ is estimated in two ways: through the restricted model just mentioned and by a weak-identification brute force grid procedure, which allows for calculating the confidence interval for $\phi$ for multiple sets of trajectories.
+If it is negative, however, it implies farmers who do better on average see a worse yield performance while using improved seeds.
+A results that can reflect  barriers, such as a lack of fertilizer access, and bad roads that restrict the ability to buy hybrid seed.
+We estimate $\phi$ in two ways: through the restricted model just mentioned and by a weak-identification grid procedure, which allows for calculating the confidence interval for $\phi$ for multiple sets of trajectories.
 The advantage of the weak-identification robust procedure is that we can test for whether the data satisfies the conditions necessary for an interpretable $\phi$ parameter, before estimating the restricted model.
@@
 The second aggregation will tell us whether adopting more times corresponds to higher comparative advantage and associated returns.
-In both cases, these aggregations tell us something about the economic landscape that drive returns.
+In both cases, these aggregations reflect the economic landscape that drive returns to adoption.
 
@@
 \par
+
 We estimate this model simultaneously using maximum likelihood estimation following \cite{lokshin2004maximum}, which accounts for the endogeneity in the yield equation.
-Statistically significant estimates for the covariances $\sigma_{1\varepsilon}$ and $\sigma_{2\varepsilon}$ would indicate the presence of endogeneity in the adoption and yield equations.
-Moreover, this approach allows us to compare the expected yields for adopters under the two regimes, the observed (adoption) and the counterfactual (non-adoption).
+Statistically significant estimates for the covariances $\sigma_{1\varepsilon}$ and $\sigma_{2\varepsilon}$ indicate the presence of endogeneity in the adoption and yield equations.
+
+A key advantage of this approach is that it allows us to compare the expected yields for adopters under the two regimes, the observed (adoption) and the counterfactual (non-adoption).
 This effect is the Average Treatment Effect on the Treated (ATT).
@@
 There is some discrepancy in the self-reported yields in the first wave, and specifically with the $(1,0,0)$ trajectory.
-The marginal effect to adoption is negative for cropcuts, but positive for self-reported yields.
-The marginal effects to adoption in many cases, are not statistically significant.
+The return to adoption is negative for cropcuts, but positive for self-reported yields.
+The returns to adoption in many cases, are not statistically significant, or they are negative.
 Only in the case of the trajectory $(0,0,1)$: those that adopted in the second and third waves, is there a large, statistically significant effect.
@@
 In order to calculate comparative advantage, there must be evidence for the LCA being satisfied.
-The lower panel of  Table \ref{tbl:unres} shows the results for the weak identification test for $\phi$.
+The lower panel of  Table \ref{tbl:unres} shows the results for the weak identification test for $\phi$, including each trajectory on its own and a joint test.
 We can see here that for the case of cropcuts with controls, there is a confidence interval for the joint test of $\phi$, giving evidence to the fact that the LCA is satisfied most prominently in that case.
@@
 
-% {\small\tabcolsep=3pt  % hold it local
-
 \input{results/tables/grc}
 
-% \resizebox{0.8\textwidth}{!}{
-%     \centering
-%     % \input{results/tables/phi_test_merged}
-%     \input{results/tables/phi_test_self}
-%     \input{results/tables/phi_test_cropcut}
-% }
-% \resizebox{0.8\textwidth}{!}{
-%     \centering
-%     % \input{results/tables/phi_test_merged}
-%     \input{results/tables/phi_test_self}
-%     \input{results/tables/phi_test_cropcut}
-% }
-
-Figure \ref{fig:theta_delta_raw} shows a plot of the comparative advantages and returns for the cropcut model with controls.
-For the most part these results highlight two interesting puzzles: that it would seem that for many trajectories, comparative advantage is either very close to 0, implying that there is no difference between farmers that adopt and not adopt, and that for trajectory $(0,1,1)$, it would seem that comparative advantage is exceedingly negative.
-This is peculiar, as it this would imply that farmers that adopted in both waves 2 and 3, would actually be better off not adopting, even though their returns are positive (as shown in the bottom figure).
+Figure \ref{fig:theta_delta_raw} shows a plot of the comparative advantages and returns for the cropcuts model with controls.
+These results highlight two interesting puzzles: it would seem that for many trajectories, comparative advantage is either very close to 0, implying that there is no difference between farmers that adopt and not adopt.
+For trajectory $(0,1,1)$, comparative advantage is exceedingly negative.
+This is peculiar, as this would imply that farmers that adopted in both waves 2 and 3, would actually be better off not adopting, even though their returns are positive (as shown in the bottom figure).
 Another peculiarity in this figure is that it would seem that if we compare trajectories that only differ by one wave's adoption, comparative advantage and returns seem to fall.
@@
 Figure \ref{fig:n_adopt} shows the results of the regression and calculation of comparative advantage.
-As expected, adopting 0 times leads to negative comparative advantage, but with positive returns to adoption, implying that there is value to adopting to those that have never adopted hybrid maize seed.
+As expected, adopting zero times leads to negative comparative advantage, but with positive returns to adoption, implying that there is value to adopting to those that have never adopted hybrid maize seed.
 Adopting once leads to both positive comparative advantage, and positive returns to adoption.
@@
 To complement the results of the GRC model, we present the Endogenous Switching Regression results.
+The fourth wave of the ESS data allows us to use DNA-fingerprinting results.
+This is advantageous over waves 1-3, because we do not need to rely on self-reported adoption status.
 We first examine a self-reported definition of an improved variety to validate the results from the GRC model.
@@
 Table \ref{tab:switch1} shows the results of the Switching regression using both definitions, and using self-reported and cropcut yields.
-We see that regardless of the definition of improved seed, yields measurement form and types of controls included, the results consistently show that the ATT is negative and the ATU is positive, meaning that farmers that adopted a self-reported improved seed would've been better off not adopting, and farmers that did not adopt a self-reported improved seed would've been better off adopting.
+We see that regardless of the definition of improved seed, yield measurement and controls included, the results consistently show that the ATT is negative and the ATU is positive, meaning that farmers that adopted a self-reported improved seed would've been better off not adopting.
+Farmers that did not adopt improved seeds would have been better off adopting.
+This is in line with the GRC results as it seems to imply that households are not making profit-maximizing decisions; their adoption decisions seem to be leading to a reduction in welfare, whether they are adopting or not adopting.
 
@@
 
-To better understand the above puzzling results, we explore similar yield estimations using the different definitions of improved seeds from the DNA fingerprinting.
-We run similar switching regressions using different thresholds of DNA purity (70, 90 and 95\% purity), for hybrid seeds (compared with open-pollinated varieties), and for Drought-tolerant maize (DTMZ).
+To better understand the above puzzling results, we explore similar yield estimations using the different definitions of improved seeds from DNA fingerprinting.
+The DNA fingerprinting data allows us to be confident in whether a seed that was used, was in fact improved.
+We run similar switching regressions using different thresholds of DNA purity (70, 90 and 95\% purity), hybrid vs.
+open-pollinated varieties, and drought-tolerant maize (DTMZ).
 Table \ref{tab:switch2} shows the results.
 
-Given that all the plots sampled for DNA fingerprinting showed a purity percent of 70\% or higher, we need a counter-factual to compare these purity thresholds with non-adopters.
-We therefore consider as non-adopters those that self-reported using a traditional variety.
-We are aware that given the universality of improved varieties with a 70\% threshold or higher and given the high rates of miss-classification, the control group for a 70\% threshold is likely to be contaminated with a high share of false-positives.
-But we also know the results are biased downwards and that the control group for higher shares of purity will be less contaminated, which provides an interesting comparison.
+The results show that the initial switching regression results and the GRC results could be driven by seed definition.
+All the plots sampled for DNA fingerprinting showed a purity percent of 70\% or higher, which means that even farmers that report using traditional seed only, are in fact still using a 70\% mix with hybrid seed.
+This implies that we need a counter-factual to compare these purity thresholds with non-adopters.
+We consider non-adopters those who self-reported using a traditional variety.
+Given the universality of improved varieties with a 70\% threshold or higher and given the high rates of misclassification, the control group for a 70\% threshold is likely to be contaminated with a high share of false positives.
+But the results are biased downwards and the control group for higher shares of purity will be less contaminated.
 
 Table \ref{tab:switch2} shows that the estimations for purity thresholds of 70 and 90\% show similar patterns as before: non-adopters would've been better off adopting and adopters would've been better off not adopting.
-But when improved seeds with a purity of 95\% or higher are compared with self-reported traditional varieties or lower purity than 95\%, the ATT becomes positive and the ATU is still positive but significantly decreases (and it is even negative when self-reported yields are used).
-Farmers that adopted a variety with a 95\% purity or greater had significantly greater yields than if they had not adopted, considering that this is even a lower boundary.
+However, when we use a 95\% purity definition, the ATT becomes positive and the ATU is still positive but significantly reduced (and even negative when self-reported yields are used).
+Farmers that adopted a variety with a 95\% purity or greater had significantly greater yields than if they had not adopted.
 
 We get similar results for DTMZ.
-Farmers that adopted DTMZ (as tracked by the DNA germplasm) did obtain greater yields than if they had not adopted.
+Farmers that adopt DTMZ (as tracked by the DNA germplasm) did obtain greater yields than if they had not adopted.
 Again, this result may be biased downwards because of possible false-positives among the control group.
@@
 To further disaggregate the results, we explore the source of the improved varieties (CGIAR or exotic) and the year of the seed release.
+These results are shown in Table \ref{tab:switch3}.
 We find positive effects for adopters of CGIAR-sourced varieties when using self-reported yields.
@@
 
-The previous results highlight the fact that the definition of hybrid matters.
+The previous results highlight the fact that the definition of improved seed matters.
 Even when households think they are not adopting hybrid maize seed, there is such a large amount of hybrid seed that has been introduced over the years in Ethiopia that farmers may be misclassifying their seed and taking part in sub-optimal growing strategies as a result.
-For instance, if a farmer thinks they are growing tradition seed, but they are actually growing hybrid seed, then they will tend to have low returns; hybrid seeds ofte require more fertilizer and other inputs to work optimally.
-If they are cultivated like traditional seeds, then their performance may end up being worse.
+For instance, if a farmer thinks they are growing traditional seed, but they are actually growing hybrid seed, returns to adoption will be low, since hybrid seeds often require more fertilizer and inputs to work optimally.\footnote{ If they are cultivated like traditional seeds, then their performance may end up being worse.}
 
@@
 In this section, we will introduce a misclassification-robust approach to the GRC method that will allow us to take misclassification into account when calculating comparative advantage.
-The results from the previous section highlight that there are many definitions to hybrid seed and the fourth wave of the ESS includes DNA fingerprinting.
+The results from the previous section highlight that there are many definitions to improved seed, which we are able to differentiate in the fourth wave of the ESS given the DNA fingerprinting.
 This gives us the ability to validate how often a hybrid seed is actually used when a farmers self-reports adopting hybrid seed.
 The methodology in this section builds on work that was done in \cite{michuda2021three}, which used confusion matrices from machine learning algorithms as a way to estimate an causal effect that was robust to misclassification.
-Although, the ESS data does not have results from machine learning algorithms, we can construct a confusion matrix based on the DNA fingerprinting data in the fourth wave and apply it to the previous three waves.
+Misclassification introduces non-classical measurement error that may attenuate estimates and bias them towards 0 in the binary case.
+In the multi-regime case, misclassification will tend to make estimates too close to one another, leading to smaller or even wrong-signed estimates.
+We construct a confusion matrix based on the DNA fingerprinting data in the fourth wave and apply it to the previous three waves.
 
@@
 Table \ref{tab:misclassification} shows a table of different hybrid seed definitions and the percentage of times that a farmer self-reported traditional or hybrid seed, and whether they in fact were using traditional or hybrid seed, based on the DNA fingerprinting data.
-You can see from Sub-table \ref{tab:a} that around 86\% of farmers actually use traditional seed when they self-report using traditional seed, and around 70\% of the time, farmers use hybrid seed of at least 95\% purity when they self-report using hybrid.
+Table \ref{tab:a} shows that around 86\% of farmers actually use traditional seed when they self-report using traditional seed, and around 70\% of the time, farmers use hybrid seed of at least 95\% purity when they self-report using hybrid.
 There are in fact instances where farmers, however, misclassify their self-reports.
-For instance, in sub-table \ref{tab:b}, a farmer that self-reports using drought-tolerant maize, only in fact uses drought-tolerant maize 13\% of the time.
+For instance, in Table \ref{tab:b}, a farmer that self-reports using drought-tolerant maize, only in fact uses drought-tolerant maize 13\% of the time.
 There are similarly large misclassifications for seed released after 2010.
@@
 \ref{eq:misclass} to construct the joint distribution, $f(h, \hat{h})$ for each trajectory in the sample.
-Since we had the most confidence in the data from Table \ref{tab:a}, we decided to use it for the main body of the paper.
+Since Table \ref{tab:a} provides a general definition of an improved seed, we use it for the main body of the paper.
+Table \ref{tab:cm} shows the results of the construction of the joint distribution.
+The first thing to notice is that the diagonal elements tend get lower with more adoption of seed.
+This is due to the fact that a farmer only has a 70\% probability of having a true-positive outcome when adopting hybrid seed.
+With three years of adoption, the probability of having at least one year of adoption being misclassified is high.
+Another thing to notice is that if, for example, a farmer adopts hybrid seed for 2 out of 3 years, their misclassification probability does not change if, they reported adopting in wave 1 and wave 2 or wave 2 and wave 3.
+Both these features of the confusion matrix have to do with the fact that these probabilities are independent of any dynamic transition or auto-correlation.
+It would be expected that if there was some learning or signal from buying seed from a particular seller, the misclassification probability would get smaller with more adoption.
+Thus, our case can be thought of as a conservative lower-bound on the effect of misclassification on comparative advantage and returns.
+
+\input{results/tables/misclass_95purity.tex}
+
+Once we construct these probabilities, we estimate a Nonlinear-Maximum Likelihood function that is robust to misclassification.
+Essentially, this involves using the likelihood function obtained from the confusion matrices and apply it on the estimation of Eq.
+\ref{eq:GRC} as a conditional distribution of agricultural yield.
+The likelihood for one observation in this case would then be:
+
+\begin{equation}
+\label{eq:mr-grc}
+l(\mu, \kappa, \Delta, \beta, \sigma|y_{t}, \hat{h}) = \sum_{\underbar{$h$}\in\mathcal{H}} \left[ f_{Y_{t}| \mathcal{H}}(y|h=\underbar{$h$})\cdot  \sum_{\bar{h} \in \mathcal{H}} 1\{\hat{h}=\bar{h}\} p_{\underbar{$h$}, \bar{h}} \right]
+\end{equation}
+
+\noindent where $1$ is the indicator function and $f$ is the probability distribution function conditional on true adoption assignment.
+We assume that this distribution is normal and can be written with a mean of $y_t - \sum_{\underline{h}\in\mathcal{H}\backslash (1,1,1)}\mu_{\underline{h}}-\sum_{\underline{h}\in\mathcal{H}_{S}}\Delta_{\underline{h}}h_{it}1\{h_{i}=\underline{h}\} - \kappa_{(1,1,1)}h_{it}1\{\underline{h}=(1,1,1)\}- X_{it}\beta$ and some nuisance variance parameter $\sigma$, which is estimated.
+Put more succinctly, we use Eq.
+\ref{eq:GRC} in the normal distribution.
+We will call this estimator, Misclassification-Robust GRC.\footnote{For more information on the derivation of the original estimator, as well as a link to a python package of its implementation, see \cite{michuda2021three}.}
+
+Once we estimate the parameters, we once again calculate comparative advantage and returns to adoption.
+Figure \ref{fig:mr-grc} shows the results of the MR-GRC and previous GRC results side-by-side.
+What's striking is that once misclassification is accounted for, comparative advantage to adoption is no longer negative to trajectories that adopt hybrid.
+In fact, the number of adoptions is positively correlated with comparative advantage.
+This suggests that households are sorting based on how well they can use hybrid.
+What is interesting, however, is that once we account for misclassification, returns to adoption become almost universally negative.
+This is indicative of a negative $phi$ term or that, there is negative selection into hybrid.
+Similar to \cite{Suri2011-oi}, a negative $\phi$ term suggests that there may be institutional factors at play that make the returns to adoption negative and this may be the reason why households do not adopt.
+The only trajectory that has positive returns to adoptions is $(0,0,1)$.
+What's peculiar about that trajectory is that comparative advantage is closest to 0 than any other trajectory.
+This may suggest that all other things being equal, adoption still presents positive returns over traditional seed, especially if that adoption occurs in the third wave.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=0.7]{results/figures/mr-theta.png}
+    \caption{Comparative Advantage and Returns to Adoption: GRC and MR-GRC Results}
+    \label{fig:mr-grc}
+            \vspace*{-2em}
+    \begin{table}[H]
+        \centering
+        \begin{tabular}{p{0.8\textwidth}} 
+            \begin{tablenotes}
+                  \small
+                  \item Note: Figure shows GRC and MR-GRC results side-by-side.
+GRC results are identical as the ones from Figure \ref{fig:theta_delta_raw}.
+MR-GRC estimated using likelihood function in Eq.
+\ref{eq:mr-grc}.
+Misclassification defined using DNA purity data from the fourth wave of ESS.
+            \end{tablenotes}
+        \end{tabular}
+    \end{table}
+\end{figure}
+
+\section{Discussion}
+
+To explain the patterns showcased in our results, we propose a simple conceptual model of genetic dilution.
+The evidence is consistent in that newly bred varieties do offer higher performance than traditional varieties, but such higher-yielding effects diminish over time, as newer generations of the improved seeds continue to be used.
+In contrast, traditional local varieties are well-adapted to the local agroecological conditions, and their reproduction over multiple generations shows consistent yields, assuming normal weather conditions.
+Depending on the specific improved varieties and the speed at which high-yielding effects diminish over time, there may be a threshold in time above which it may be more profitable to use a traditional variety than an improved one that has been around for too many generations (Figure \ref{fig:generations}).
+After such a threshold is reached, because improved varieties are more dependent on external inputs, at a constant rate of fertilizer application, the yields of the older-generation improved seeds would sharply decrease, so that local traditional varieties would return in higher productivity.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=0.7]{results/figures/generations.png}
+    \caption{Hypothetical yields for recycled and traditional varieties}
+    \label{fig:generations}
+\end{figure}
+
+This depiction is certainly an over-simplification, as many crucial variables are taken as constant, such as seed costs, fertilizer application, water input use, weather conditions, and the specificities of the improved seeds (that would determine for example how resistant they are to drought conditions).
+Adding some of these features to a similar model, \cite{Bird2020-nt} propose a theoretical model where retained local traditional varieties are compared with two types of improved varieties: one that is non-locally adapted, and one bred for the specific local agro-climatic conditions.
+The authors postulate that while profits of both types of improved varieties increase with fertilizer application, profits of traditional varieties would decrease with fertilizer application, implying the existence of a threshold of fertilizer application below which it is more profitable to keep using traditional varieties.
+Non-locally adapted varieties are less profitable without complementary fertilizer application.
+Depending on how responsive to fertilizer both improved varieties are and on the costs of seeds, there may be a threshold of fertilizer application below which it is more profitable to use the locally adapted variety than the improved non-adapted one.
+
+These findings may also hold for our case, especially taking into consideration the degree of genetic dilution and dispersion of recycled or older-generation improved varieties, which are not as yield responsive to fertilizer application.
+In addition to this, prior evidence also indicates that farmers sub-optimally allocate fertilizer in response to their mislead beliefs about the crop variety (Kosmowski \& Lybbert, 2022; \citealt{wossen2022misperceiving}).
+Such wrong beliefs or misclassification has also been shown to be driven by misperception instead of misreporting (\citealt{wossen2022misperceiving}), which is also the case for the farmers in our sample.
+These findings suggest that the seemingly small to negative returns to adoption that we find through the GRC model are driven by both a lower response to fertilizers of the older-generation improved varieties, but also to farmers’ sub-optimal application of fertilizers (and possibly other inputs) due to misperception.
+
+The suitability of the improved varieties to weather conditions and the weather conditions themselves are two additional factors that may further explain the seemingly negative comparative advantages to adopting the misperceived improved varieties.
+Even moderate droughts have been shown to decrease maize yields of non-drought tolerant varieties by 20\% (\citealt{paul2021heterogeneous}).
+Additionally, even the drought-tolerant improved maize varieties distributed in Ethiopia are only resistant to mild drought but would also fail like any other non-drought tolerant variety in case of severe droughts.
+
+In addition to these factors, we are solely evaluating the comparative advantages of improved varieties over maize yields, but further analysis on net returns would be crucial for this context.
+Obtaining a first-generation high-yielding variety or drought tolerant ones represents a significant investment for farmers that usually use traditional varieties (most of the sample).
+The fertilizer, irrigation and other complementary inputs necessary for the proper response of the improved varieties also represent a substantial investment.
+Because of this, our prior is that the findings from this study would be aggravated if the outcome variable was net returns instead of maize yields, but this remains an open question for future research.
 
@@
 But the fact that newly bred hybrids that adapted to local conditions outperform local varieties and older-generation improved seeds is not surprising given the large investments devoted to improve seed performance and the speed at which new high-yielding varieties are being developed.
-The results from the Group Random Coefficient model highlights how the fact that we imperfectly observe the use of newly improved seeds and the degree of mis-classification between self-reported improved seeds and bred hybrids may partially explain why we do not see self-reported improved varieties grant positive returns to adoption.
+The results from the Group Random Coefficient model highlights how the fact that we imperfectly observe the use of newly improved seeds and the degree of misclassification between self-reported improved seeds and bred hybrids may partially explain why we do not see self-reported improved varieties grant positive returns to adoption.
 Reconciling these findings with the results from the Endogenous Switching Regression, we hypothesize that the negative returns to adoption mask a large heterogeneity in the returns to adoption.
@@
 Such dispersion of improved genetic material also poses a challenge for the proper identification of the returns to adoption as it confounds the potential benefits of locally adapted hybrids, but also the benefits of traditional local varieties that may be better suited for certain agro-ecological conditions and for contexts of low-input use.
-
-Our findings are in line with prior research.
-\cite{Bird2020-nt} propose a theoretical model where the authors compare retained local traditional varieties with improved non-locally adapted varieties and locally adapted improved varieties.
-The authors postulate that while profits of both types of improved varieties increase with fertilizer application, profits of traditional varieties would decrease with fertilizer application, implying the existence of a threshold of fertilizer application below which it is more profitable to keep using traditional varieties.
-Non-locally adapted varieties are less profitable without complementary fertilizer application.
-Depending on how responsive to fertilizer both improved varieties are and on the costs of seeds, there may be a threshold of fertilizer application below which it is more profitable to use the locally adapted variety than the improved non-adapted one.
-\par
-
-It is possible that similar dynamics explain the patterns showcased in our results.
-The evidence is consistent in that newly bred varieties do offer higher performance than traditional varieties, but such higher-yielding effects diminish over time, as newer generations of the improved seeds continue to be used.
-In contrast, traditional local varieties are well-adapted to the local agro-ecological conditions, and their reproduction over multiple generations shows consistent yields, assuming normal weather conditions.
-Depending on the specific improved varieties and the speed at which high-yielding effects diminish over time, there may be a threshold in time above which it may be more profitable to use a traditional variety than an improved one that has been around for too many generations.
-After such a threshold is reached, because improved varieties are more dependent on external inputs, the costs of the older-generation improved seeds, together with the costs of fertilizers and other inputs may not compensate the yield returns of an older-improved variety.
-\par
 
@@
 
+% \textbf{Next steps}
+% \begin{enumerate}
+%     \item We shoudl try to frame our non-experimental results in the light of Suri's: With experimental data, one can test for thepresence of heterogeneous returns, but estimating the distribution of returns requires assump-tions about the underlying selection process, which is randomized away in such an experiment(see Heckman, Smith, and Clements (1997).
+he experimental and IV evidence appeared to indicate thatmaize growers in Kenya were leaving “money on the table” by not adopting hy-brid seed.
+However, the experiments and IV results reveal little else about thereturns.
+The framework of heterogeneous returns in this paper allows the es-timation of not just average returns, but also the distribution of returns acrossthe sample of farmers.
+I find strong evidence of heterogeneity in returns tohybrid maize, with comparative advantage playing an important role in the de-termination of yields and adoption decisions.
+Finally,the heterogeneity in returns to the hybrid technology, on the whole, suggestsquite rational and relatively unconstrained adoption of existing hybrid strains,in contrast to the evidence from the experimental and IV literatures
+%     \item 
+% \end{enumerate}
+
+*/
+
 \bibliography{references}

### 2022-11-04 document/slides/NEUDC 2022/main.tex
Updates from Overleaf

--- 2022-07-21 document/slides/SPIA Final Meeting/finalpresentation.tex
@@
 \usecolortheme{default}
+\usepackage{natbib}
+\bibliographystyle{unsrtnat}
 
@@
 
-\author[et al]{Aleksandr Michuda \\ Cristina Chiarella \\Oscar Barriga-Cabanillas \\Juan Correa }
+\author[et al]{Aleksandr Michuda\inst{1} \and Cristina Chiarella\inst{2} \and Oscar Barriga-Cabanillas\inst{3} \and Juan Correa\inst{4}}
+
+\institute{\inst{1} Cornell University \and \inst{2} UCLouvain \and \inst{3} World Bank Corp. \and \inst{4} FAO}
 
@@
 
-\date[January 2022] % (optional)
-{CGIAR-SPIA \\
-January, 2022}
+\date[November 2022] % (optional)
+{NEUDC   \\
+November 2022}
 
@@
 \begin{frame}
-\frametitle{Research question}
+\frametitle{Introduction}
 
@@
     \item Agro-ecological considerations 
-\end{enumerate}
-
-\textbf{We study how heterogeneity masks the true impact of adoption}
-
-\begin{enumerate}[ {[}1{]} ]
-    \item Explore heterogeneous comparative advantages through GRC: find puzzling no significant results.
-    \item Compare definitions of improved seeds: self-reported, second generation or recycled, purity of genetic material, year of release, types and source.
-    \item Low adoption of newer and purest varieties, which do return in higher yields. Poor farmers paying a premium for genetically diluted improved varieties.
+    \item Lack of learning \citep{Conley2010-ue}
 \end{enumerate}
@@
 
+\begin{frame}{Research Question}
+
+\textbf{How does misclassification in adoption status mask heterogeneous impacts to adoption?}
+\end{frame}
+
+\begin{frame}{Preview of Results}
+
+\begin{enumerate}[ {[}1{]} ]
+    \item Explore heterogeneous comparative advantages through GRC: find puzzling results.
+    \item Endogenous switching regression with different definitions of improved seeds: self-reported, second generation or recycled, purity of genetic material, year of release, types and source.
+    \item Low adoption of newest and purest varieties, which do return in higher yields. 
+    \item When we account for misclassification in GRC, the puzzle is resolved
+\end{enumerate}
+\end{frame}
+
 %---------------------------------------------------------
 \begin{frame}
-\frametitle{Data}
+\frametitle{Ethiopian Maize}
 
@@
     % \item Struggled to provide an adequate and reliable food supply
+    \item Use ESS waves 1-4, agricultural questionnaires
+    \item Waves 1-3 were a panel and included information on yields, inputs and hybrid maize adoption
+    \item Wave 4 is a break from Waves 1-3, but contains similar information $+$ DNA fingerprinting
     \item We focus on maize producing households 
     \item Self-reported use of improved seed varieties
-    \item Compare self-reported and cropcut yields
+    \item Self-reported yields
 \end{itemize}
@@
 
-%---------------------------------------------------------
-\begin{frame}
-\frametitle{Empirical Strategy I: Finding the returns from adoption}
+\begin{frame}{}
 
-\textbf{Understand the returns of adopting improved varieties: }
-\begin{itemize}
-    \item Some farmers have higher returns to adoption (comparative advantage)
-    \item How comparative advantage plays a role on using improved seed varieties.
-\end{itemize}
-
-% (Suri 2011; Tjernstr\"{o}m et al. 2020; Michler et al. 2019; Barriga et al. 2018)
-
-We use a Correlated Random Coefficient Model:
-\vspace{-2em}
-\begin{center}
-$$
-    y_{it}= \delta + \Delta h_{it} + \theta_{i} + \phi\theta h_{it} + \tau_{i} + \epsilon_{it}
-$$  
-\end{center}
-
-\begin{itemize}
-    \item $y_{it}$: yields or profits for household i at time t
-    \item $h$:  indicator for adoption
-    \item $\tau$:  household’s absolute advantage
-    \item \textbf{$\theta$: household’s comparative advantage}
-    \item \textbf{$\Delta$: The marginal effect of adoption}
-    \item $\phi$: Gains from adopting, relative to a household’s comparative advantage.
-\end{itemize}
+\begin{figure}
+    \centering
+    \includegraphics[width=.7\textwidth]{results/figures/adoption_r4.png}
+    \caption{Caption}
+    \label{fig:my_label}
+\end{figure}
 
@@
 
-Advantages:
+\textbf{Understand the returns of adopting improved varieties: }
+\begin{itemize}
+    \item Some farmers are ``more suited'' for adoption (comparative advantage) \citep{Suri2011-oi}
+    \item How comparative advantage plays a role on using improved seed varieties and its returns.
+    \item Use Group Random Coefficient (GrRC) method \citep{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc}
+    \item Can separate returns from those who ``should'' be getting highest returns.
+\end{itemize}
+
+% (Suri 2011; Tjernstr\"{o}m et al. 2020; Michler et al. 2019; Barriga et al. 2018)
+
+We estimate a Correlated Random Coefficient Model by way of GrRC estimation:
+\vspace{-2em}
+\begin{center}
+$$
+    y_{it}= \delta + \Delta h_{it} + \phi\boldsymbol{\theta_i} h_{it} + \tau_{i} + \boldsymbol{\theta_i} + \epsilon_{it}
+$$  
+\end{center}
+
+\end{frame}
+
+%---------------------------------------------------------
+\begin{frame}
+\frametitle{Empirical Strategy}
 \begin{enumerate}
-    \item Identification does not rely on a IV 
-    % (linear projection of observed adoption history)
-    \item Disentangles absolute advantage from comparative advantage in adoption. ($\theta_i = \theta_A - \theta_N$)
+    \item Identification does not rely on an IV or propensity to adopt.
+    \item Identification similar to \cite{Chamberlain1984-uk}: strict exogeneity
+    \item strong assumption, but gives us ability to estimate returns to adoption, as well as \textit{advantage in adopting}
+    \item Requires enough variation in ``switcher trajectories'' 
 \end{enumerate}
@@
 \input{results/tables/traj}
-
-% Extending the model:
-
-% \begin{itemize}
-%     \item Implementing a nonparametric panel identification in a novel GRC (Group Random Coefficients)
-%     \item Modeling and incorporating time-varying characteristics 
-%     \item Complement ESS data with precipitation (CHIRPS) and temperature (CPC) data
-% \end{itemize}  
 \end{frame}
@@
 \begin{frame}
-\frametitle{Empirical Strategy II: Returns to Different Types of Hybrid}
+\frametitle{The Puzzle}
 % \frametitle{Group Random Coefficients Results}
-\textbf{Why farmers do \textit{(not)} adopt?}
-
-\begin{enumerate}
-    \item There are negative returns from adoption
-    \item Comparative advantage is close to 0 unless you always adopt (perhaps driven by 
-\end{enumerate}
 
@@
     \centering
-    \label{fig:traj_sum}
-    \includegraphics[scale=.4]{results/figures/traj_sum.png}
+    \includegraphics[scale=0.5]{results/figures/theta.png}\label{fig:theta_delta_raw}
 \end{figure}
@@
 \begin{frame}
-\frametitle{Endogenous Switching Regression Results}
-
-Why do we see farmers adopting despite the seemingly small negative returns?
+\frametitle{What's going on?}
 
 \begin{itemize}
-    \item We use the DNA-fingerprinting to shed light on this puzzle
-    \item DNA allows us to assess what varieties (and purity) farmers are using 
-    \item Unfortunately, only available in the last round
+    \item Negative advantage to adopting
+    \item Adopting more associated with less advantage
+    \item Negative returns (but that's part of the standard story about access)
 \end{itemize}
 
-We implement a switching regression approach
+\end{frame}
+
+\begin{frame}{How do we resolve this puzzle?}
+\begin{itemize}
+    \item Explore 4th wave
+    \item We use the DNA-fingerprinting to shed light on this puzzle
+    \item DNA allows us to assess what varieties (and purity) farmers are using, 
+    \item We implement an endogenous switching regression approach
+\end{itemize}
 
@@
 
-\input{results/tables/switch23-sr}
+\begin{table}[H]
+\resizebox{.75\textwidth}{!}{
+\centering
+\hspace*{-1.2cm}
+\begin{threeparttable}
+\caption{Effects on yields (log) from adopting improved maize varieties (DNA fingerprinting)}
+\label{tab:switch2}
+\begin{tabular}{l cccccc}
+\hline
+\hline
+            &Adopters yields&Non-adopters yields&         ATE&          SE&     p-value\\
+\hline
+\textit{DNA 70, SR yields}&            &            &            &            &            \\
+ATT         &        7.17&        7.31&      \fcolorbox{red}{lightgray}{-0.14}&       0.026&      0.0000\\
+%
+%
+%
+ATU         &        8.76&        6.76&        2.00&       0.036&      0.0000\\
+%
+%
+%
+\textit{DNA 90, SR yields}&            &            &            &            &            \\
+ATT         &        7.18&        9.07&       \fcolorbox{red}{lightgray}{-1.90}&       0.023&      0.0000\\
+%
+%
+%
+ATU         &        7.01&        6.80&        0.21&       0.022&      0.0000\\
+%
+%
+%
+\textit{DNA 95, SR yields}&            &            &            &            &            \\
+ATT         &        7.30&        7.08&        \fcolorbox{red}{lightgray}{0.22}&       0.036&      0.0000\\
+%
+%
+%
+ATU         &        6.87&        6.86&       0.018&       0.029&      0.5496\\
+%
+%
+%
+\textit{HYB, SR yields}&            &            &            &            &            \\
+ATT         &        7.28&        7.25&       \fcolorbox{red}{lightgray}{0.030}&       0.028&      0.2892\\
+%
+%
+%
+ATU         &        7.06&        6.85&        0.21&       0.028&      0.0000\\
+%
+%
+%
+\textit{DTMZ, SR yields}&            &            &            &            &            \\
+ATT         &        7.42&        6.53&        \fcolorbox{red}{lightgray}{0.89}&       0.074&      0.0000\\
+%
+%
+%
+ATU         &        7.11&        7.02&       0.089&       0.022&      0.0001\\
+%
+%
+%
+\textit{CGIAR source, SR yields}&            &            &            &            &            \\
+ATT         &        7.13&        6.71&        \fcolorbox{red}{lightgray}{0.43}&       0.033&      0.0000\\
+%
+%
+%
+ATU         &        9.19&        6.89&        2.31&       0.018&      0.0000\\
+%
+%
+%
+\textit{Year 2010+, SR yields}&            &            &            &            &            \\
+ATT         &        7.44&        6.92&        \fcolorbox{red}{lightgray}{0.53}&       0.039&      0.0000\\
+%
+%
+%
+ATU         &        8.98&        6.91&        2.07&       0.056&      0.0000\\
+%
+%
+%
+\hline
+\hline
+\end{tabular}
+\end{threeparttable}
+}
+\end{table}
 
@@
     \item All DNA fingerprints are 70\% pure or higher. Control group has high share of false positives: no proper  control.  
-
-\end{itemize}
-Negative GRC results mask a large heterogeneity in returns to adoption. 
-\begin{itemize}
-    \item Low purchase of newly bred hybrids, diluted DNA, low inputs use.
-
 \end{itemize}
 
-Effects of high-yielding varieties diminishes over time, but traditional varieties are well adapted to local agro-ecological conditions.
-
+Could misclassification be masking returns heterogeneity?
 \begin{itemize}
-    \item There may be a threshold in time above which it may be more profitable to use a traditional variety than an improved one that has been around for too many generations.
-
+    \item If farmers using some mix of traditional hybrid maize or using incorrect practices, this would attenuate effects
+    \item Trajectories would be misclassified.
 \end{itemize}
@@
 
-%---------------------------------------------------------
-\begin{frame}{}
-    Thank you!
+\begin{frame}{Misclassification Driving Returns?}
+\begin{itemize}
+    \item Use a misclassification robust variant of GRC
+    \item Maximum likelihood estimation that uses misclassification rates in estimation, inspired by \cite{michuda2021three}
+    \item Use Wave 4 misclassification rates for Waves 1-3
+\end{itemize}
+
+\begin{equation*}
+\label{eq:mr-grc}
+l(\mu, \kappa, \Delta, \beta, \sigma|y_{t}, \hat{h}) = \sum_{\underbar{$h$}\in\mathcal{H}} \left[ f_{Y_{t}| \mathcal{H}}(y|h=\underbar{$h$})\cdot  \sum_{\bar{h} \in \mathcal{H}} \boldsymbol{1\{\hat{h}=\bar{h}\} p_{\underbar{$h$}, \bar{h}}} \right]
+\end{equation*}
+
 \end{frame}
 
-\begin{frame}{Raw Trajectories}
+\begin{frame}{Misclassification Rates from Wave 4}
+
+\begin{itemize}
+    \item Use rate of reported hybrid of at least 95\% purity, against genetic fingerprinting to calculate misclassification
+    \item Farmer who reports using traditional seed actually using hybrid 14\% of the time
+    \item Farmer who reports using hybrid seed is actually using traditional seed almost \textit{30\%} of the time.
+\end{itemize}
+    \begin{table}[]
+        \centering
+        \begin{tabular}{lll}
+        \toprule
+         & 0 & 1 \\
+        \midrule
+        0 & 0.863 & 0.137 \\
+        1 & 0.296 & 0.704 \\
+        \bottomrule
+        \end{tabular}
+    \end{table}
+
+\begin{equation}
+    \label{eq:misclass}
+    P(h = (i,j,k) , \hat{h} = (\hat{i},\hat{j}, \hat{k})) = p_{i, \hat{i}}\cdot p_{j, \hat{j}}\cdot p_{k, \hat{k}}
+\end{equation}
+\end{frame}
+
+\begin{frame}{Misclassified GRC Results}
+
 \begin{figure}
     \centering
-        \includegraphics[scale=.5]{results/figures/theta.png}
+    \includegraphics[scale=0.4]{results/figures/mr-theta.png}
 \end{figure}
+\end{frame}
+
+\begin{frame}{Conclusion}
+
+Accounting for misclassification resolves the issue and makes the adoption problem the classic story about access and a lack of profitability.
+
+\begin{itemize}
+    \item Misclassification may work through a measurement problem, but also through changes in practices
+    \item Mismatch between cultivation practices and seed used
+    \item Effects of high-yielding varieties diminish over time, but traditional varieties are well adapted to local agro-ecological conditions.
+    \begin{itemize}
+    \item There may be a threshold in time above which it may be more profitable to use a traditional variety than an improved one that has been around for too many generations.
+    \item Example: 010 trajectory
+\end{itemize}
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+Thank you!
+
+\begin{itemize}
+    \item I would be happy to discuss any questions and please feel free to email me!
+    \item am2497@cornell.edu
+\end{itemize}
+
 \end{frame}

### 2022-11-04 paper.tex

@@
 Since ESS1-ESS3 do not have DNA fingerprinting, we are left with only self-reports of hybrid use.
-We use these variables in a Group Random Coefficiens (GRC) approach to investigate the heterogeneous effects of adoption.
-The advantage of the GRC approach is that we do not rely on identification from a valid instrument, but on modelling the heterogeneity directly through a projection of adoption returns on their observed ``trajectory'' of adoption.
-This allows us to directly capture two effects: heterogeneous effects in the form of so-called ``comparative advantage" of adoption, and, under the assumption of linearity in comparative advantage, a description of the overall economy's sorting of farmers based on their comparative advantage.
+We use these variables in a Group Random Coefficients (GRC) approach to investigate the heterogeneous effects of adoption.
+The advantage of the GRC approach is that we do not rely on identification from a valid instrument, but on modeling the heterogeneity directly through a projection of adoption returns on their observed ``trajectory'' of adoption.
+This allows us to directly capture two effects: heterogeneous effects in the form of the so-called ``comparative advantage" of adoption, and, under the assumption of linearity in comparative advantage, a description of the overall economy's sorting of farmers based on their comparative advantage.
 The first is important because it is able to disentangle the effect of a general household's agricultural ability (which can be thought of as a fixed effect) and their adoption-specific potential.

### 2023-03-20 document/slides/NEUDC 2022/discussant/main.tex
Updates from Overleaf

--- 2022-01-27 document/slides/finalpresentation.tex
@@
 %Title page
-\title[Genetic Dilution in Improved Maize] %optional/
-{A Case of Genetic Dilution in Improved Maize: Estimating
-Comparative Advantage of Adoption in Ethiopia}
+\title{Discussing: "Seeding the Seeds: Role of Social Structure in
+Agricultural Technology Diffusion"}
 
-\author[et al]{Michuda, Chiarella, Barriga-Cabanillas, Correa, }
+\author[et al]{Alain de Janvry, Manaswini Rao, Elisabeth Sadoulet}
 
-% {Oscar Barriga-Cabanillas\inst{1} \\
-% (with Travis Lybbert\inst{2})}
+% \institute[UC Davis] % (optional)
+% {
+% %  \inst{1}%
+%   Agricultural and Resource Economics\\
+%   UC Davis
+%  }
 
-\institute[UC Davis] % (optional)
-{
-%  \inst{1}%
-  Agricultural and Resource Economics\\
-  UC Davis
- }
-
-\date[January 2022] % (optional)
-{CGIAR-SPIA \\
-January, 2022}
+\date[November 2022] % (optional)
+{Discussant:\\
+Aleksandr Michuda}
 
@@
 %---------------------------------------------------------
-\begin{frame}
-\frametitle{Research question}
-
-Food security is a major challenge: Ethiopia is no exception
-
+\begin{frame}{Short Summary}
     \begin{itemize}
-    \item Struggled to provide an adequate and reliable food supply
-    \item Breakthroughs in maize germplasm ($\uparrow$ yields, $\uparrow$ drought tolerance)
-\end{itemize}
-
-However adoption of improved seed has remained a challenge
-
-\begin{enumerate}
-    \item Market integration
-    \item Access to supplementary inputs
-    \item Agro-ecological considerations
-\end{enumerate}
-
-\textbf{We study how heterogeneity masks the true impact of adoption}
-
-\begin{enumerate}
-    \item Adoption where  net benefits are the highest
-    \item How adaptation strategies augment comparative advantage in adoption
-    \item Drivers of dis-adoption of improved maize seed
-\end{enumerate}
-
-\end{frame}
-
-%---------------------------------------------------------
-\begin{frame}
-\frametitle{Empirical Strategy I}
-
-Correlated Random Coefficient Model and comparative advantage (Suri 2011; Tjernstr\"{o}m et al. 2020; Michler et al. 2019; Barriga et al. 2018)
-
-\begin{itemize}
-    \item Farmers face unobservable heterogeneous returns to adoption (comparative advantage)
-\end{itemize}
-
-\begin{center}
-$$
-    y_{it}= \delta + \beta h_{it} + \theta_{i} + \phi\theta h_{it} + \tau_{i} + \epsilon_{it}
-$$  
-\end{center}
-
-\begin{itemize}
-    \item $y_{it}$: yields or profits for household i at time t
-    \item $h$:  indicator for adoption
-    \item $\tau$:  household’s absolute advantage
-    \item $\theta$: household’s comparative advantage
-    \item $\phi$: Gains from adopting, relative to a household’s comparative advantage. 
-\end{itemize}
-
-\end{frame}
-
-%---------------------------------------------------------
-\begin{frame}
-\frametitle{Empirical Strategy II and Extensions}
-
-Advantages:
-\begin{enumerate}
-    \item Identification does not rely on the existence of a valid instrumental variable (linear projection of observed adoption history)
-    \item Disentangles a household’s absolute advantage from comparative advantage in adopting. 
-\end{enumerate}
-
-Extending the model:
-
-\begin{itemize}
-    \item Implementing a nonparametric panel identification in a novel GRC (Group Random Coefficients)
-    \item Modeling and incorporating time-varying characteristics 
-    \item Complement ESS data with precipitation (CHIRPS) and temperature (CPC) data
+        \item Returning to the site of two-staged randomized experiment in Emerick et al. (2016)
+        \item Introduction of flood resistant paddy variety 
+        \item Evaluating diffusion of seed from treatment to non-treatment households
+        \item Dynamic treatment effects mask heterogeneity found across jati (sub-castes) in regions
+        \item Less village-level fractionalization of jati leads to higher levels of diffusion and adoption by non-treatment households
     \end{itemize}
@@
 
-%---------------------------------------------------------
-%---------------------------------------------------------
-\begin{frame}
-\frametitle{Group Random Coefficients Results}
-
+\begin{frame}{Highlights}
+    \begin{itemize}
+        \item Compelling story; just as important to look at spillovers as for treatment, itself
+        \item Merges two interesting literatures: seed tech. adoption and economics of Indian caste
+        \item has clear policy implications for future policy 
+        interventions 
+        \item In particular, social hierarchy results are very interesting!
+    \end{itemize}    
 \end{frame}
 
-%---------------------------------------------------------
+\begin{frame}{Feedback}
+    \begin{itemize}
+        \item Regression of Flood x Treatment unclear for untreated households
+            \begin{itemize}
+                \item Being in more flood-prone areas will increase rate of adoption?
+                \item perhaps using lagged treatment would do better?
+            \end{itemize}
+        \item Why use an indicator for ELF, and not the continuous variable?
+        \item In many cases, for a particular year, adoption and area cultivated increases with similar jati rank, or having someone from your jati receive the mini-kit, but the seed source outcome stays insignificant or is only significant in 2015
+            \begin{itemize}
+                \item The argument in the paper is that this diffusion happens through one's network, however, the lack of results from seed source makes me want a story about mechanisms.
+How are these seeds being used if not through seed exchange?
+                \item If seed is readily available, is this story more about ``information'' exchange than the physical seed?
+            \end{itemize}
+    \end{itemize}
+\end{frame}
 
-%---------------------------------------------------------
-\begin{frame}
-\frametitle{Endogenous Switching Regression Results}
-
-\input{results/tables/switch23-sr}
-
-\end{frame}
-%---------------------------------------------------------
-
-%---------------------------------------------------------
-\begin{frame}
-\frametitle{Discussion}
+\begin{frame}{Feedback}
 
 \begin{itemize}
-    \item Negative returns to adoption from self-reported improved varieties.
-    \item But positive returns to more genetically pure, DTMZ, CG- sourced and newer varieties.
-    \item All DNA fingerprints are 70\% pure or higher. Control group has high share of false positives: no proper control. 
-    \item Negative results from GRC mask a large heterogeneity in returns to adoption.
+    \item How often does a treated village also get treated with a flood?
+    \item In some cases, I'm inclined to think that the paper might be have power challenges, and whether it might be useful to have a continuous flood measure?
+Villages that might not have full-on floods, but that are close and observe floods might also be more likely to adopt
 \end{itemize}
 
+\textbf{Minor Feedback}
+    \begin{itemize}
+        \item Why not make all tables into figures as you have for other regressions?
+    \end{itemize}
 \end{frame}
 
-%---------------------------------------------------------
+\bibliography{references}
 

### 2023-03-20 document/slides/NEUDC 2022/main.tex
Updates from Overleaf

@@
 %\usepackage{utopia} %font utopia imported
-
 \usepackage{amsthm, booktabs, setspace, natbib, amsfonts,amssymb,epstopdf,textcomp,listings, xcolor}
@@
 
-\author[et al]{Aleksandr Michuda\inst{1} \and Cristina Chiarella\inst{2} \and Oscar Barriga-Cabanillas\inst{3} \and Juan Correa\inst{4}}
+\author[]{Aleksandr Michuda\inst{1} \and Cristina Chiarella\inst{2} \and Oscar Barriga-Cabanillas\inst{3} \and Juan Correa\inst{4}}
 
-\institute{\inst{1} Cornell University \and \inst{2} UCLouvain \and \inst{3} World Bank Corp. \and \inst{4} FAO}
+\institute[]{\inst{1} Cornell University \and \inst{2} UCLouvain \and \inst{3} World Bank Corp. \and \inst{4} FAO}
 
@@
 
-However adoption of improved seed has remained a challenge
+However adoption of improved seed has remained a challenge:
 
 \begin{enumerate}[ {[}1{]} ]
-    \item Market integration
-    \item Access to supplementary inputs
-    \item Agro-ecological considerations 
-    \item Lack of learning \citep{Conley2010-ue}
+    \item Lack of market integration \citep{Carter2014-fm, adejobi2013markets}
+    \item Access to supplementary inputs \citep{Byerlee2013-qk, Munshi2004-og}
+    \item Adaptation to agro-ecological conditions \citep{Bird2020-nt}
+    \item Tech. Transfer Challenges of learning \citep{Conley2010-ue}
 \end{enumerate}
@@
 \begin{enumerate}[ {[}1{]} ]
-    \item Explore heterogeneous comparative advantages through GRC: find puzzling results.
+    \item Explore heterogeneous comparative advantages through GRrC (reformulation of \cite{Suri2011-oi}): find puzzling results.
     \item Endogenous switching regression with different definitions of improved seeds: self-reported, second generation or recycled, purity of genetic material, year of release, types and source.
-    \item Low adoption of newest and purest varieties, which do return in higher yields. 
-    \item When we account for misclassification in GRC, the puzzle is resolved
+    \item Low adoption of newest and purest varieties, which \textit{in fact} return higher yields. 
+    \item Due to genetic dilution, all farmers are using improved seed; biasing results
+    \item When we account for misclassification in GRrC, the puzzle is resolved
 \end{enumerate}
@@
     \item We focus on maize producing households 
-    \item Self-reported use of improved seed varieties
-    \item Self-reported yields
+    \item Self-reported use of improved seed varieties and yields
 \end{itemize}
@@
     \includegraphics[width=.7\textwidth]{results/figures/adoption_r4.png}
-    \caption{Caption}
-    \label{fig:my_label}
+    \caption{Seed varieties: Adoption rate of improved seed varieties in 2018 (by self-report, purity level, crop type, source and year of release)}
 \end{figure}
@@
 \begin{frame}
-\frametitle{Empirical Strategy I: Finding the returns from adoption}
+\frametitle{Empirical Strategy}
 
@@
     \item How comparative advantage plays a role on using improved seed varieties and its returns.
-    \item Use Group Random Coefficient (GrRC) method \citep{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc}
-    \item Can separate returns from those who ``should'' be getting highest returns.
+    \item Use Group Random Coefficient (GrRC) method \citep{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} to  separate returns from those who ``should'' be getting highest returns.
 \end{itemize}
@@
 \begin{enumerate}
-    \item Identification does not rely on an IV or propensity to adopt.
+    \item Identification does not require IV
     \item Identification similar to \cite{Chamberlain1984-uk}: strict exogeneity
-    \item strong assumption, but gives us ability to estimate returns to adoption, as well as \textit{advantage in adopting}
+    \item Strong assumption, but gives us ability to estimate returns to adoption, as well as \textit{advantage in adopting}
     \item Requires enough variation in ``switcher trajectories'' 
 \end{enumerate}
+\end{frame}
 
+\begin{frame}{Trajectories}
 \input{results/tables/traj}
+
 \end{frame}
@@
 \begin{frame}
-\frametitle{The Puzzle}
+\frametitle{The Puzzle: Using high yield varieties does not improved returns}
 % \frametitle{Group Random Coefficients Results}
@@
 \begin{itemize}
-    \item Explore 4th wave
-    \item We use the DNA-fingerprinting to shed light on this puzzle
-    \item DNA allows us to assess what varieties (and purity) farmers are using, 
-    \item We implement an endogenous switching regression approach
+    \item Use the  4th wave of the survey to leverage DNA-fingerprinting information 
+    \item DNA allows us to assess what varieties (and purity) farmers are in reality planting  
+    \item We implement an endogenous switching regression approach to properly measure yield return 
+    \item Assess the genetic purity of improved varieties, year of release, types and sources
 \end{itemize}
@@
 \begin{itemize}
-    \item If farmers using some mix of traditional hybrid maize or using incorrect practices, this would attenuate effects
+    \item If farmers using genetically diluted hybrid, or older generations, this would attenuate effects
     \item Trajectories would be misclassified.
@@
 \begin{itemize}
-    \item Use a misclassification robust variant of GRC
-    \item Maximum likelihood estimation that uses misclassification rates in estimation, inspired by \cite{michuda2021three}
-    \item Use Wave 4 misclassification rates for Waves 1-3
+    \item We estimate a misclassification robust GRrC 
+    \item A Maximum likelihood estimation approach uses misclassification rates to properly estimate coefficients on the returns from adoption of high yield varieties $-$inspired by \cite{michuda2021three}
+    \item We apply wave 4 misclassification rates for waves 1-3
 \end{itemize}
@@
 \begin{equation*}
-\label{eq:mr-grc}
+\label{eq:mr-GRrC}
 l(\mu, \kappa, \Delta, \beta, \sigma|y_{t}, \hat{h}) = \sum_{\underbar{$h$}\in\mathcal{H}} \left[ f_{Y_{t}| \mathcal{H}}(y|h=\underbar{$h$})\cdot  \sum_{\bar{h} \in \mathcal{H}} \boldsymbol{1\{\hat{h}=\bar{h}\} p_{\underbar{$h$}, \bar{h}}} \right]
@@
 \begin{itemize}
-    \item Use rate of reported hybrid of at least 95\% purity, against genetic fingerprinting to calculate misclassification
-    \item Farmer who reports using traditional seed actually using hybrid 14\% of the time
-    \item Farmer who reports using hybrid seed is actually using traditional seed almost \textit{30\%} of the time.
+    \item Use rate of hybrid of at least 95\% purity, against self-reported hybrid use to calculate misclassification
+    \item A farmer who reports using traditional seed is actually using hybrid 14\% of the time
+    \item A farmer who reports using hybrid seed is actually using traditional seed almost \textit{30\%} of the time.
 \end{itemize}
@@
         \end{tabular}
+        \caption{Misclassification Matrix of 95\% genetic purity}
     \end{table}
@@
 
-\begin{frame}{Misclassified GRC Results}
+\begin{frame}{Misclassified GRrC Results}
 
@@
 \begin{itemize}
-    \item Misclassification may work through a measurement problem, but also through changes in practices
-    \item Mismatch between cultivation practices and seed used
+    \item Farmers think they're adopting improved seed varieties but evidence shows considerable dilution and varieties are not yielding as expected
+    \item Misclassification biases estimates through measurement problem, but also through changes in practices
+    \item Mismatch between cultivation practices and seed variety used
     \item Effects of high-yielding varieties diminish over time, but traditional varieties are well adapted to local agro-ecological conditions.
@@
 \begin{frame}
-Thank you!
+\textbf{Thank you!}
 

### 2023-03-20 paper.tex
Updates from Overleaf

@@
  Aleksandr Michuda \footnote{Center for Data Science for Enterprise and Society, Cornell University}%
- \and Cristina Chiarella\footnote{UC Louvain}%
+ \and Cristina Chiarella\footnote{UCLouvain}%
  \and Oscar Barriga-Cabanillas\footnote{The World Bank}%
@@
 			Despite large investments in  developing new  maize varieties, evidence shows that few farmers consistently adopt this new technology over time.
-We use an innovative methodology that studies the adoption trajectories of farmers over a three-round panel data set to explain low adoption rates with heterogeneous returns to adoption.
+We study the adoption trajectories of farmers over a three-round panel data set to explain low adoption rates with heterogeneous returns to adoption.
 By relying on self-reported use of improved seeds, we find a puzzling answer for why farmers do not use new seed varieties: adoption leads to seemingly small and negative returns.
-Heterogeneity analysis across different levels of comparative advantage show that farmers who adopt don't actually have a comparative advantage in adopting improved seed, nor do they exhibit high returns to adopting.
+Heterogeneity analysis across different levels of comparative advantage show that farmers who adopt do not have a comparative advantage in adopting improved seed, nor do they exhibit high returns to adopting.
 To reconcile this result with the expected higher yields for improved seeds, we exploit unique information on maize DNA-fingerprinting collected over the same areas in 2018.
@@
 Food security is a major challenge in many East African countries, and Ethiopia is no exception, having historically struggled to provide an adequate and reliable food supply \citep{Ramakrishna2002-hv, Jaleta2018-oj}.
-Since the drought of 1984, Ethiopia has become increasingly reliant on maize, and as of today, maize constitutes the most widespread crop in the country.\footnote{FAO statistics show the total area cultivated with Maize in 2019 is 25\% larger than the next most widespread crop (sorghum), and 27\% and 139\% larger than the next two following crops, wheat and barley, respectively.} As Figure \ref{fig:maize_yields} shows, there is an upward trend in the total area cultivated with maize since 1990, with average yields almost doubling during this period.
-This increase in maize yields has coincided with a dramatic decrease in undernourishment of the population which has fallen from 47\% in 1990 to 14\% in 2018 (The World Bank, 2021).
-But despite these positive gains in total area planted and average yields, food insecurity still affects more than half the population of Ethiopia.\footnote{The country has also recently seen increases in food insecurity because of COVID-19, the conflict in Ethiopia’s Tigray region and the desert locust outbreak.} Moreover, a large productivity gap persists between the African continent and the rest of the developing world.
+Since the drought of 1984, Ethiopia has become increasingly reliant on maize, and as of today, maize constitutes the most widespread crop in the country.\footnote{FAO statistics show the total area cultivated with Maize in 2019 is 25\% larger than the next most widespread crop (sorghum), and 27\% and 139\% larger than the next two following crops, wheat and barley, respectively.} There was an upward trend in the total area cultivated with maize since 1990, with average yields almost doubling during this period.
+\footnote{Improved security conditions and maize productivity contributed with an important decrease in undernourishment which fell from 47\% in 1990 to 14\% in 2018 (The World Bank, 2021).
+Despite these positive gains in total area planted and average yields, food insecurity still affects more than half the population of Ethiopia.The country has also recently seen increases in food insecurity because of COVID-19, the conflict in Ethiopia’s Tigray region and the desert locust outbreak.} Despite progress a large productivity gap persists between the African continent and the rest of the developing world.
+From 2016 to the present, cereal crop yields in Sub-Saharan Africa have consistently been around 70-80\% lower than Europe or North America \citep{worldbankyield}.
 The persistence of this gap is often attributed to different adoption rates of green revolution technologies, which include improved seeds.
-Evidence has shown that the adoption of improved seed leads to increases in agricultural yields \citep{Carter2014-fm}, food security \citep{Shiferaw2014-op}, and poverty reduction \citep{Minten2008-tj}.\footnote{Nevertheless, critics of green revolution technologies have also claimed that these technological packages have not been as profitable for many farmers in Africa as promised (with the exception of Ethiopia), given the high costs, and their heavy dependence on subsidies \citep{Wise20}.}
+Evidence has shown that the adoption of improved seed leads to increases in agricultural yields \citep{Carter2014-fm}, food security \citep{Shiferaw2014-op}, and poverty reduction \citep{Minten2008-tj}.\footnote{Nevertheless, critics of green revolution technologies have also claimed that these technological packages have not been as profitable for many farmers in Africa as promised.
+Even the most successful of these countries, Ethiopia, is not on track to meet the goals of the Alliance for a Green Revolution for Africa (AGRA).
+This is due to high costs, and heavy dependence on subsidies \citep{Wise20}.}
 
-\begin{figure}[htp]
-    \centering
-    \caption{Maize: Total area harvested and average yields} \label{fig:maize_yields}
-    \includegraphics[width=.6\textwidth]{results/figures/Maize_yields.pdf}
-    \vspace*{-1em}
-    \begin{table}[H]
-        \centering
-        \begin{tabular}{p{0.6\textwidth}} 
-            \begin{tablenotes}[flushleft]
-                  \small
-                  \item Source: Authors' calculations using FAOSTAT data 
-                  \item Note: Yields are measured as the harvested production per ha for the area under cultivation
-            \end{tablenotes}
-        \end{tabular}
-    \end{table}       
-\end{figure}
+% \begin{figure}[htp]
+%     \centering
+%     \caption{Maize: Total area harvested and average yields} \label{fig:maize_yields}
+%     \includegraphics[width=.6\textwidth]{results/figures/Maize_yields.pdf}
+%     \vspace*{-1em}
+%     \begin{table}[H]
+%         \centering
+%         \begin{tabular}{p{0.6\textwidth}} 
+%             \begin{tablenotes}[flushleft]
+%                   \small
+%                   \item Source: Authors' calculations using FAOSTAT data 
+%                   \item Note: Yields are measured as the harvested production per ha for the area under cultivation
+%             \end{tablenotes}
+%         \end{tabular}
+%     \end{table}       
+% \end{figure}
 
@@
 Several answers have been proposed to explain such low adoption of improved varieties.
-These include lack of liquidity and other economic constraints \citep{Carter2014-fm}, growing conditions such as the type of soil, input use, and farmer characteristics \citep{Munshi2004-og}, imperfections in credit markets \citep{Croppenstedt2003-pq}, ambiguously defined property rights \citep{Place2000-el}, lack of commitment devices \cite{Duflo2009-iv}, high transportation costs that increase the cost of agricultural inputs \citep{Byerlee2013-qk}, and improved seed supply constraints \citep{Bird2020-nt}.
-Additionally, the net returns to adoption of improved seeds are often lower than expected in the long term because a large percentage of farmers unknowingly use diluted improved seed varieties, or ones that are over 20 years old.
-With such old seed, there is evidence that yield improvement and disease resistance wanes over time\citep{Abate2015-rj}.
+These include lack of liquidity and other economic constraints \citep{Carter2014-fm}, growing conditions such as the type of soil, input use, and farmer characteristics \citep{Munshi2004-og}, imperfections in credit markets \citep{Croppenstedt2003-pq}, ambiguously defined property rights \citep{Place2000-el}, lack of commitment devices \citep{Duflo2009-iv}, high transportation costs that increase the cost of agricultural inputs \citep{Byerlee2013-qk}, and improved seed supply constraints \citep{Bird2020-nt}.
+Additionally, the net returns to adoption of improved seeds are often lower than expected in the long term because a large percentage of farmers may  use diluted improved seed varieties, or ones that have been recycled over many generations.
+This is a concern as there is evidence that yield improvement and disease resistance wanes over time \citep{Abate2015-rj}.
 The variety of these constraints and their dispersion across subpopulations and locations speaks to considerable heterogeneous returns to adoption.
@@
 
-In this paper, we revisit the mechanism proposed by \citep{Suri2011-oi}, where a lack of adoption of improved seed varieties is the result of farmers' heterogeneous returns to the new technology.
+In the absence of a natural experiment or randomized intervention, it is difficult to properly estimate the effect of adoption as there is potential selection of certain farmers into adoption improved varieties.
+This selection bias is more pervasive when the goal is to explore heterogeneous effects of adoption.
+Alternatively, one can make assumptions about the structure of the underlying error distribution to properly identify returns to adoption.
+In this paper, we revisit the latter mechanism proposed by \cite{Suri2011-oi}, where a lack of adoption of improved seed varieties is the result of farmers' heterogeneous returns to the new technology.
 Ignoring such heterogeneity hides the true impact of adoption, as farmers with different characteristics and skill levels observe different returns from adoption despite the technology providing high average returns.
-The heterogeneous returns to adoption are modeled as a farmer's unobservable comparative advantage using a Correlated Random Coefficient (CRC) model.
-In the original paper, \citep{Suri2011-oi} finds that farmers’ lack of adoption is driven by insufficient net benefits to the technology stemming from poor infrastructure that increases the cost of access.
-\footnote{Similar work in Ethiopia on improved chickpea varieties suggests that the economic returns to adoption plays a major role in the decision to adopt.
-\citep{Michler2018-wk}.}
+The heterogeneous returns to adoption are modeled as a farmer's unobservable comparative advantage.
+In the original paper, \cite{Suri2011-oi} finds that farmers’ lack of adoption is driven by insufficient net benefits to the technology stemming from poor infrastructure that increases the cost of access.
+Similar work in Ethiopia on improved chickpea varieties suggests that the economic returns to adoption plays a major role in the decision to adopt.
+\citep{Michler2018-wk}.
 
 %% Expand a bit on the methods
-We build on this literature by investigating improved maize seed adoption in Ethiopia.
+% We build on this literature by investigating improved maize seed adoption in Ethiopia.
 In particular, we implement two complementary estimation strategies to address both how heterogeneous returns drive the decision to use improved seed varieties, and the extent that the misclassification of varieties and genetic dilution affect the decision to adopt them by lowering their observed returns.
-First, we implement a Group Random Coefficients estimation, which builds on \cite{Suri2011-oi} in \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} using a General Method of Moments (GMM) Estimator that allows us to test \textbf{ the assumptions made by \cite{Suri2011-oi}}, as well as provide more economically interpretable results.
+First, we implement a Group Random Coefficients estimation, which builds on \cite{Suri2011-oi} and \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} using a General Method of Moments (GMM) estimator that allows us to test \textbf{ the assumptions made by \cite{Suri2011-oi}}, as well as provide more economically interpretable results.
 The advantage of our estimation strategy is that we do not rely on identification from a valid instrument.
@@
 
-For the first set of results, our empirical strategy relies on the first three survey rounds of the Ethiopian Socioeconomic Survey (ESS).
-The data were collected as a panel, which cover 2011, 2013, and 2015.
+Our results are divided into two parts.
+We start by using the first three survey rounds of the Ethiopian Socioeconomic Survey (ESS), which is a panel  covering  2011, 2013, and 2015.
 For each round, farmers self-reported on the seed variety they used (improved or traditional).
-By using the adoption trajectories over three rounds, we find the returns to adoption are in most cases, not statistically significant, and only significant with wave 3 adoption.
+By using the adoption trajectories over three rounds, we find the returns to adoption are in most cases, not statistically significant, and only significant for adoption in the third wave of the survey.
 This posits a puzzle that motivates our next estimation strategy.
 
-We expand our results using the fourth round of the ESS (conducted in 2018), which provides
-a unique opportunity to understand why improved seed varieties seem not to increase returns
-when adopted.
-The ESS4 collected DNA fingerprints of harvested cropcuts in selected plots and compared its genetic material with a reference library of available varieties in Ethiopia.
-This data provides critical information not only about the seed types but also about their purity, sources, and their year of release.
-We exploit this uniquely rich information to compare different definitions of improved seeds.
-Specifically, we compare returns of verifiably improved seeds against self-reported use of new, second generation or recycled varieties, comparing different purity thresholds (or degrees of no-contamination with other varieties), year of release, types of varieties, and their sources.
+Our second set of results use the fourth round of the ESS (conducted in 2018), which collected DNA fingerprints of harvested cropcuts in selected plots and compared its genetic material with a reference library of available varieties in Ethiopia, hence, providing a unique opportunity to understand why improved seed varieties seem not to increase returns when adopted.
+These data provides critical information not only about the seed types but also about their purity, sources, and their year of release.
+We exploit this uniquely rich information to compare returns of verifiable improved seed varieties against self-reported use of new, second generation or recycled varieties.
+For this, we leverage different purity thresholds (or degrees of no-contamination with other varieties), year of release, types of varieties, and their sources.
 Because the latest survey round is not a continuation of the first three rounds, our second estimation strategy implements an Endogenous Switching Regression (ESR) approach, a standard estimation strategy to study the question of adoption with cross-sectional data (\citealt{Marenya2020-kb}, \citealt{Shiferaw2014-op}), given its capacity to account for unobserved heterogeneity in adoption.
@@
 This method takes inspiration from \cite{michuda2021three} to account for the fact that the self-reports from the first three rounds of the ESS data may be misclassified and attenuated.
-We find that once we account for misclassification, heterogeneity in adoptions becomes very similar to other contexts; higher comparative advantage is associated with lower returns pointing to significant barriers to adoption for those that would benefit the most from it.
+We find that once we account for misclassification, heterogeneity in adoptions becomes very similar to other contexts: higher comparative advantage is associated with lower returns, pointing to significant barriers to adoption for those that would benefit the most from it.
 Incorporating misclassification rates from the fourth round of the ESS into the GRC estimation resolves the puzzle posed by the GRC results.
@@
 %, controlling for important drivers of adoption such as rainfall (found to be the main driver of heterogeneous effects for adopting agronomic packages \citealt{Marenya2020-kb}, \citealt{Katengeza2019-af}), labor and input use.
-Our findings provide an alternative hypothesis to the puzzle of low maize adoption, namely that the wide dispersion of older generations and recycled improved seeds, may be resulting in lower yields and altering farmers' input use.
+Our findings provide an alternative hypothesis to the puzzle of low maize adoption, namely that the wide dispersion of older generations and recycled improved seeds may be resulting in lower yields and altering farmers' input use.
 The null and negative effects that we see in the GRC results are actually stemming from misclassification in self-reported adoption status.
@@
 
-\subsection{The Three Round Panel}
+\subsection{The Three Round Panel and the Fourth Wave}
 
@@
 Table \ref{tbl:summary} shows a set of summary statistics for the sample.
-The mean parcel size for a household is about 0.15 Ha, with a maximum size of a little less than 1.5 Ha.
+The mean parcel size for a household is about 0.15 Ha, with a maximum size of less than 3.7 Ha.
 Harvest labor is disproportionately skewed to family participation with around 20 days of harvest labor being conducted by family members and only about 2 days for hired labor.
@@
 
-A seed is considered “improved” in Ethiopia if it was tested by breeders and evaluated to be superior to existing varieties  (\citep{MoA13}; \citep{kosmowski2020shining}).
+A seed is considered “improved” in Ethiopia if it was tested by breeders and evaluated to be superior to existing varieties  (\citealp{MoA13}; \citealp{kosmowski2020shining}).
 In Ethiopia, the government controls most of the seed system, with local agricultural offices or extension agents assessing the demands of different varieties.
-The MoA decides the seed quantities to be produced, and actors such as regional seed enterprises, seed companies, seed unions, are involved in the production of certified seeds, which are later distributed by seed producer cooperatives (SPCs).\footnote{They account for 37\% of the total distribution in the country (\citep{kosmowski2020shining}).} An improved seed can be obtained directly from such distribution centers (which would distribute a first generation seed), however it is often the case that seeds cultivated from second or later generations from the originally bred seed are also exchanged between farmers.
+The MoA decides the seed quantities to be produced, and actors such as regional seed enterprises, seed companies, seed unions, are involved in the production of certified seeds, which are later distributed by seed producer cooperatives (SPCs).\footnote{They account for 37\% of the total distribution in the country (\citealp{kosmowski2020shining}).} An improved seed can be obtained directly from such distribution centers (which would distribute a first generation seed), however it is often the case that seeds cultivated from second or later generations from the originally bred seed are also exchanged between farmers.
 In the survey, farmers are asked to report whether the variety they used was a first-generation improved seed, second generation, recycled, or a traditional variety.
@@
 
-Fifty four varieties have been released in Ethiopia since 1990.\footnote{Out of which seven were released between 1990 to 1999, 29 between 2000 to 2009, and 18 between 2010 to 2019} Thirty four such varieties contain International Maize and Wheat Improvement Center (CIMMYT) germplasm, out of which, 10 Drought-tolerant maize varieties (DTMZ) and 8 Quality protein maize (QPM) were released (most of these between 2010 to 2019).
-Besides the CIMMYT-related germplasm, ten are open-pollinated varieties (OPVs) and 25 are hybrid (\citep{kosmowski2020shining}).
+Fifty four varieties have been released in Ethiopia since 1990.\footnote{Out of which seven were released between 1990 to 1999, 29 between 2000 to 2009, and 18 between 2010 to 2019.} Thirty four such varieties contain International Maize and Wheat Improvement Center (CIMMYT) germplasm, out of which, 10 Drought-tolerant maize varieties (DTMZ) and 8 Quality protein maize (QPM) were released (most of these between 2010 to 2019).
+Besides the CIMMYT-related germplasm, ten are open-pollinated varieties (OPVs) and 25 are hybrid (\citealp{kosmowski2020shining}).
 
@@
 The collection of cropcuts for crop yield estimation facilitated the process for DNA fingerprinting.
-For the cropcuts, a 4 square meter quadrant was randomly selected in each plot, and total production harvested from the quadrant was weighed (both when freshly harvested and when dried).
+For the cropcuts, a four square meter quadrant was randomly selected in each plot, and total production harvested from the quadrant was weighed (both when freshly harvested and when dried).
 The DNA material was extracted from samples from the dried cropcuts, and were then matched with the genetic material of the maize DNA reference library for Ethiopia, obtained from a CIMMYT and Ethiopian Institute of Agricultural Research (EIAR) project.
@@
 It is only when using purity thresholds of 90 and 95\% (DNA 90, DNA 95 in Figure \ref{fig:adoption_r4}, respectively), that adoption rates of improved varieties change to 91 and 54\% of sampled households, respectively.\footnote{We also find that 62\% of households adopted a hybrid variety, compared to 38\% that adopted an open-pollinated variety, and 15\% of households adopted a drought-tolerant maize variety, while 63\% of households that adopted an improved variety use a CGIAR-derived germplasm, and 79\% use an exotic germplasm.} Finally, 64\% of households adopted a variety that was released in the year 2000 or after, and 24\% adopted a variety that was adopted in the year 2010 or after.
-These ratios are in line with prior findings (\cite{Zeng15}).
+These ratios are in line with prior findings (\citealp{Zeng15}).
 
@@
 In the regions of Oromia, Amhara, Tigray, and SNNPR, around 39\% of households report adopting an improved varierty .
-In the regions of Amhara, Benishangul-Gumuz, Oromia, SNNPR, and Tigray, adoption rates of improved maize varieties were found for 27\% of households (\cite{Jaleta2018-oj}).
-In Oromia, SNNPR, and Benishangul-Gumuz adoption of improved varieties per farmers’ self-reports was found for 84\% of maize growers in 2011, and 88\% of maize growers in 2013 (\cite{Yirga17}).
+In the regions of Amhara, Benishangul-Gumuz, Oromia, SNNPR, and Tigray, adoption rates of improved maize varieties were found for 27\% of households (\citealp{Jaleta2018-oj}).
+In Oromia, SNNPR, and Benishangul-Gumuz adoption of improved varieties per farmers’ self-reports was found for 84\% of maize growers in 2011, and 88\% of maize growers in 2013 (\citealp{Yirga17}).
 
@@
 
-Since ESS1-ESS3 do not have DNA fingerprinting, we are left with only self-reports of hybrid use.
-We use these variables in a Group Random Coefficients (GRC) approach to investigate the heterogeneous effects of adoption.
+We use self-reported and cropcut yields as outcomes in a Group Random Coefficients (GRC) approach to investigate the heterogeneous effects of adoption.
 The advantage of the GRC approach is that we do not rely on identification from a valid instrument, but on modeling the heterogeneity directly through a projection of adoption returns on their observed ``trajectory'' of adoption.
@@
 
-The endogenous switching regression holds certain advantages over other methodologies used for identifying unbiased estimates with cross-sectional observational data such as propensity score matching or other instrumental variable approaches (\cite{Shiferaw2014-op}) and has been used to study similar questions in the past (\citealt{falco2011does}, \citealt{kabunga2012yield}).
+The endogenous switching regression holds certain advantages over other methodologies used for identifying unbiased estimates with cross-sectional observational data such as propensity score matching or other instrumental variable approaches (\citealp{Shiferaw2014-op}) and has been used to study similar questions in the past (\citealp{falco2011does}; \citealp{kabunga2012yield}).
 \par
@@
 
-The results of the regression are shown in Table \ref{tbl:unres}.
+The results of the GRC regression are shown in Table \ref{tbl:unres}.
 The results for self-reported and cropcut yields are largely similar.
@@
 For trajectory $(0,1,1)$, comparative advantage is exceedingly negative.
-This is peculiar, as this would imply that farmers that adopted in both waves 2 and 3, would actually be better off not adopting, even though their returns are positive (as shown in the bottom figure).
+This is peculiar, as this would imply that farmers that adopted in both waves 2 and 3 would actually be better off not adopting, even though their returns are positive (as shown in the bottom figure).
 Another peculiarity in this figure is that it would seem that if we compare trajectories that only differ by one wave's adoption, comparative advantage and returns seem to fall.
@@
 The results show that the initial switching regression results and the GRC results could be driven by seed definition.
-All the plots sampled for DNA fingerprinting showed a purity percent of 70\% or higher, which means that even farmers that report using traditional seed only, are in fact still using a 70\% mix with hybrid seed.
+All the plots sampled for DNA fingerprinting showed a purity percent of 70\% or higher, which means that even farmers that report using traditional seed only, could be in fact still using a 70\% mix with hybrid seed.
 This implies that we need a counter-factual to compare these purity thresholds with non-adopters.
@@
 This gives us the ability to validate how often a hybrid seed is actually used when a farmers self-reports adopting hybrid seed.
-The methodology in this section builds on work that was done in \cite{michuda2021three}, which used confusion matrices from machine learning algorithms as a way to estimate an causal effect that was robust to misclassification.
+The methodology in this section builds on work that was done in \cite{michuda2021three}, which used confusion matrices from machine learning algorithms as a way to estimate a causal effect that was robust to misclassification.
 Misclassification introduces non-classical measurement error that may attenuate estimates and bias them towards 0 in the binary case.
@@
 With three years of adoption, the probability of having at least one year of adoption being misclassified is high.
-Another thing to notice is that if, for example, a farmer adopts hybrid seed for 2 out of 3 years, their misclassification probability does not change if, they reported adopting in wave 1 and wave 2 or wave 2 and wave 3.
+Another thing to notice is that if, for example, a farmer adopts hybrid seed for 2 out of 3 years, their misclassification probability does not change if they reported adopting in wave 1 and wave 2 or wave 2 and wave 3.
 Both these features of the confusion matrix have to do with the fact that these probabilities are independent of any dynamic transition or auto-correlation.
@@
 Even without explicit DNA fingerprinting evidence, the GRC could show this trend with the comparative advantage results; returns are harmful to those with a high comparative advantage to adopt, but these returns are less negative with time.
-We speculate that this is due to the fact that new and more genetically pure hybrids are introduced giving those that adopt higher yields than before.
+We speculate that this is due to the fact that new and more genetically pure hybrids are introduced, giving those that adopt higher yields than before.
 
-But the fact that newly bred hybrids that adapted to local conditions outperform local varieties and older-generation improved seeds is not surprising given the large investments devoted to improve seed performance and the speed at which new high-yielding varieties are being developed.
-The results from the Group Random Coefficient model highlights how the fact that we imperfectly observe the use of newly improved seeds and the degree of misclassification between self-reported improved seeds and bred hybrids may partially explain why we do not see self-reported improved varieties grant positive returns to adoption.
+But the fact that newly bred hybrids that have adapted well to local conditions outperform local varieties and older-generation improved seeds is not surprising given the large investments devoted to improve seed performance and the speed at which new high-yielding varieties are being developed.
+The results from the GRC model highlights how the fact that we imperfectly observe the use of newly improved seeds and the degree of misclassification between self-reported improved seeds and bred hybrids may partially explain why we do not see self-reported improved varieties grant positive returns to adoption.
 Reconciling these findings with the results from the Endogenous Switching Regression, we hypothesize that the negative returns to adoption mask a large heterogeneity in the returns to adoption.
@@
 
-This result buttresses the fact that a large share of farmers are not able to obtain the promised greater yields from using hybrid varieties.
+This result buttresses the fact that a large share of farmers are not able to obtain the promised greater yields from using improved varieties.
 This may be partially driven by the low-purchasing rates of newly bred varieties, but also by the under-use of critical inputs needed for the successful performance of hybrid seeds, such as fertilizers and irrigation, especially in drought-prone areas.
@@
 Our findings highlight the wide distribution of the yield returns to adoption, partially explained by the low purchase of newly bred hybrids and the high dispersion of diluted improved genetic material over time.
-This points to the need for making policies oriented towards increasing the adoption of improved seeds and yields more inclusive, aiming at ensuring that improved varieties are not only available to purchase locally but also affordably.
+This points to the need for making policies oriented towards increasing the adoption of improved seeds and yields more inclusive, aiming at ensuring that improved varieties are not only available to purchase locally but also affordable.
 It is important to consider that improved varieties should also be locally adapted to specific agro-ecological conditions, such that their success is less dependent on external chemical inputs.
@@
 % \begin{enumerate}
+% add comments from slides
 %     \item We shoudl try to frame our non-experimental results in the light of Suri's: With experimental data, one can test for thepresence of heterogeneous returns, but estimating the distribution of returns requires assump-tions about the underlying selection process, which is randomized away in such an experiment(see Heckman, Smith, and Clements (1997).
@@
 
-*/
-
 \bibliography{references}

### 2023-03-22 paper.tex
Updates from Overleaf

@@
     \begin{center}\begin{abstract}
+    \singlespacing
 			\noindent  
@@
 
-        {\small \noindent\emph{JEL Classification}:
+        {\small \noindent\emph{JEL Classification}: O12, O13, O33, Q12, C13 \\
         	\emph{Keywords}: Improved maize, DNA-fingerprinting, Group Random Coefficient, Endogenous Switching Regression, Heterogeneous returns}
@@
 Food security is a major challenge in many East African countries, and Ethiopia is no exception, having historically struggled to provide an adequate and reliable food supply \citep{Ramakrishna2002-hv, Jaleta2018-oj}.
-Since the drought of 1984, Ethiopia has become increasingly reliant on maize, and as of today, maize constitutes the most widespread crop in the country.\footnote{FAO statistics show the total area cultivated with Maize in 2019 is 25\% larger than the next most widespread crop (sorghum), and 27\% and 139\% larger than the next two following crops, wheat and barley, respectively.} There was an upward trend in the total area cultivated with maize since 1990, with average yields almost doubling during this period.
-\footnote{Improved security conditions and maize productivity contributed with an important decrease in undernourishment which fell from 47\% in 1990 to 14\% in 2018 (The World Bank, 2021).
-Despite these positive gains in total area planted and average yields, food insecurity still affects more than half the population of Ethiopia.The country has also recently seen increases in food insecurity because of COVID-19, the conflict in Ethiopia’s Tigray region and the desert locust outbreak.} Despite progress a large productivity gap persists between the African continent and the rest of the developing world.
+Since the drought of 1984, Ethiopia has become increasingly reliant on maize, and as of today, maize constitutes the most widespread crop in the country, with total area cultivated increasing and average yields almost doubling since 1990.\footnote{FAO statistics show the total area cultivated with Maize in 2019 is 25\% larger than the next most widespread crop (sorghum), and 27\% and 139\% larger than the next two following crops, wheat and barley, respectively.}\footnote{Improved security conditions and maize productivity contributed with an important decrease in undernourishment which fell from 47\% in 1990 to 14\% in 2018 (The World Bank, 2021).
+But despite these gains, food insecurity still affects more than half the population of Ethiopia.
+The country has also recently seen increases in food insecurity because of COVID-19, the conflict in Ethiopia’s Tigray region and the desert locust outbreak.} But despite this, progress a large productivity gap persists between the African continent and the rest of the developing world.\footnote{From 2016 to the present, cereal crop yields in Sub-Saharan Africa have consistently been around 70-80\% lower than Europe or North America \citep{worldbankyield}.} The persistence of this gap is often attributed to different adoption rates of green revolution technologies, which include improved seeds.
+Evidence has shown that the adoption of improved seed leads to increases in agricultural yields \citep{Carter2014-fm}, food security \citep{Shiferaw2014-op}, and poverty reduction \citep{Minten2008-tj}.\footnote{Nevertheless, critics of green revolution technologies have also claimed that these technological packages have not been as profitable for many farmers in Africa as promised.
+Even the most successful of these countries, Ethiopia, is not on track to meet the goals of the Alliance for a Green Revolution for Africa (AGRA).
+This is due to high costs, and heavy dependence on subsidies \citep{Wise20}.} But in fact, only 13\% of maize farmers in Ethiopia have adopted improved seeds consistently between 2011 and 2015.
+A large percentage has adopted new varieties inconsistently, and the largest share of farmers (62\%) have not used improved seeds at all during the same period.
+
+%old version
+% \footnote{FAO statistics show the total area cultivated with Maize in 2019 is 25\% larger than the next most widespread crop (sorghum), and 27\% and 139\% larger than the next two following crops, wheat and barley, respectively.} There was an upward trend in the total area cultivated with maize since 1990, with average yields almost doubling during this period.\footnote{Improved security conditions and maize productivity contributed with an important decrease in undernourishment which fell from 47\% in 1990 to 14\% in 2018 (The World Bank, 2021).
+But despite these positive gains in total area planted and average yields, food insecurity still affects more than half the population of Ethiopia.The country has also recently seen increases in food insecurity because of COVID-19, the conflict in Ethiopia’s Tigray region and the desert locust outbreak.} Despite progress a large productivity gap persists between the African continent and the rest of the developing world.
 From 2016 to the present, cereal crop yields in Sub-Saharan Africa have consistently been around 70-80\% lower than Europe or North America \citep{worldbankyield}.
@@
 
-Despite evidence and numerous technological breakthroughs in maize germplasm with increased yields and drought tolerance, only 13\% of maize farmers in Ethiopia have adopted improved seeds consistently between 2011 and 2015.
-A large percentage has adopted new varieties inconsistently, and the largest share of farmers (62\%) have not used improved seeds at all during the same period.
 Several answers have been proposed to explain such low adoption of improved varieties.
 These include lack of liquidity and other economic constraints \citep{Carter2014-fm}, growing conditions such as the type of soil, input use, and farmer characteristics \citep{Munshi2004-og}, imperfections in credit markets \citep{Croppenstedt2003-pq}, ambiguously defined property rights \citep{Place2000-el}, lack of commitment devices \citep{Duflo2009-iv}, high transportation costs that increase the cost of agricultural inputs \citep{Byerlee2013-qk}, and improved seed supply constraints \citep{Bird2020-nt}.
-Additionally, the net returns to adoption of improved seeds are often lower than expected in the long term because a large percentage of farmers may  use diluted improved seed varieties, or ones that have been recycled over many generations.
-This is a concern as there is evidence that yield improvement and disease resistance wanes over time \citep{Abate2015-rj}.
-The variety of these constraints and their dispersion across subpopulations and locations speaks to considerable heterogeneous returns to adoption.
+An alternative mechanism for a lack of improvement has the misclassification of seed and seed dilution.
+Net returns to improved seed is often lower than expected in the long term because a large percentage of farmers use diluted improved seed varieties, or ones that have been recycled over many generations.
+Evidence suggests that yield improvement and disease resistance wanes over time \citep{Abate2015-rj}.
+Conversely, given the ubiquity of improved seed varieties over the course of decades of access in Ethiopia, some farmers may use improved seed without knowing it, leading to inferior yields than conventional seed due to a lack of the application of complementary inputs.
+
+Misclassification of improved seed has been studied in the case of cassava in Nigeria \citep{wossen2019estimating,macours2019comment}.
+The authors use DNA fingerprinting data to show that there is a 22\% bias in the productivity impact of improved cassava seed.
+In the context of seed adoption, however, there is evidence to suggest that the productivity impacts of seed affects households heterogeneously \citep{Suri2011-oi}, and that knowing average impacts to adoption may not be as informative.
+The observed effect of adoption may be low because households that would have a comparative advantage in adoption of improved seed are constrained by a lack of access or infrastructure.
+Additionally, the non-classical measurement error introducted by misclassification can bias heterogeneity in ways that may attenuate or even reverse productivity impacts.
+This paper provides a way to account for misclassification of seed when additional, error-free data exists, such as DNA fingerprinting data.
+We estimate the comparative advantage of households directly using a Group Random Coefficients (GRC) Model and correct its heterogeneity estimates using misclassification rates from DNA fingerprinting data of the same plots.
+\citep{Suri2011-oi, Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc}.
 
 %%%% Let's evaluate the effect of adoption in the context of Ethiopia
-
-In the absence of a natural experiment or randomized intervention, it is difficult to properly estimate the effect of adoption as there is potential selection of certain farmers into adoption improved varieties.
-This selection bias is more pervasive when the goal is to explore heterogeneous effects of adoption.
-Alternatively, one can make assumptions about the structure of the underlying error distribution to properly identify returns to adoption.
-In this paper, we revisit the latter mechanism proposed by \cite{Suri2011-oi}, where a lack of adoption of improved seed varieties is the result of farmers' heterogeneous returns to the new technology.
-Ignoring such heterogeneity hides the true impact of adoption, as farmers with different characteristics and skill levels observe different returns from adoption despite the technology providing high average returns.
-The heterogeneous returns to adoption are modeled as a farmer's unobservable comparative advantage.
-In the original paper, \cite{Suri2011-oi} finds that farmers’ lack of adoption is driven by insufficient net benefits to the technology stemming from poor infrastructure that increases the cost of access.
-Similar work in Ethiopia on improved chickpea varieties suggests that the economic returns to adoption plays a major role in the decision to adopt.
-\citep{Michler2018-wk}.
 
@@
 
-Our results are divided into two parts.
-We start by using the first three survey rounds of the Ethiopian Socioeconomic Survey (ESS), which is a panel  covering  2011, 2013, and 2015.
-For each round, farmers self-reported on the seed variety they used (improved or traditional).
-By using the adoption trajectories over three rounds, we find the returns to adoption are in most cases, not statistically significant, and only significant for adoption in the third wave of the survey.
-This posits a puzzle that motivates our next estimation strategy.
-
-Our second set of results use the fourth round of the ESS (conducted in 2018), which collected DNA fingerprints of harvested cropcuts in selected plots and compared its genetic material with a reference library of available varieties in Ethiopia, hence, providing a unique opportunity to understand why improved seed varieties seem not to increase returns when adopted.
-These data provides critical information not only about the seed types but also about their purity, sources, and their year of release.
-We exploit this uniquely rich information to compare returns of verifiable improved seed varieties against self-reported use of new, second generation or recycled varieties.
-For this, we leverage different purity thresholds (or degrees of no-contamination with other varieties), year of release, types of varieties, and their sources.
-Because the latest survey round is not a continuation of the first three rounds, our second estimation strategy implements an Endogenous Switching Regression (ESR) approach, a standard estimation strategy to study the question of adoption with cross-sectional data (\citealt{Marenya2020-kb}, \citealt{Shiferaw2014-op}), given its capacity to account for unobserved heterogeneity in adoption.
-The results of the ESR show negative effects of adopting improved seeds, as per farmer self-reports.
-But the comparison of different improved seed definitions using the DNA fingerprinting shows that when the definition of improved seed is based on greater purity of the germplasm (95\% of purity or more), the returns to adoption for those that used such improved varieties becomes positive and significant.
-We obtain similar results when evaluating the specific effects of adopting drought-tolerant maize varieties, CGIAR or Exotic varieties, or more recently released varieties.
-
-Taken together, these results offer a plausible answer to the puzzle of low adoption levels of new varieties.
-The negative and not statistically significant returns to adoption we find from both estimation strategies reflect the loose definition of improved seeds when such improved varieties are self-reported by farmers.
-Only a subset of farmers use the purest genetic varieties, drought-tolerant varieties, or newly released varieties, which provide the highest returns.
-This implies that poorer farmers are paying a premium for genetically diluted hybrids, which has far-reaching policy implications about the need to foster a more inclusive strategy when making the best performing seeds available to all farmers.
-
-Finally, to integrate our results together, we use the misclassification rates from the 4th round of the data and re-estimate the GRC results to be robust to misclassification.
-This method takes inspiration from \cite{michuda2021three} to account for the fact that the self-reports from the first three rounds of the ESS data may be misclassified and attenuated.
-We find that once we account for misclassification, heterogeneity in adoptions becomes very similar to other contexts: higher comparative advantage is associated with lower returns, pointing to significant barriers to adoption for those that would benefit the most from it.
-Incorporating misclassification rates from the fourth round of the ESS into the GRC estimation resolves the puzzle posed by the GRC results.
-
 Our results contribute to the literature on the drivers of low adoption rates of improved seed varieties.
@@
 Other important determinants of heterogeneous effects include risk aversion \citep{Holden2016-vy}, farm size \citep{Ghimire2015-bd}, and credit constraints \citep{Simtowe2008-jn,Balana2020-hx}.
+We are the first to consider how heterogeneous impacts to seed adoption are affected by misclassification of seed.
+
+% This implies that poorer farmers are paying a premium for genetically diluted hybrids, which has far-reaching policy implications about the need to foster a more inclusive strategy when making the best performing seeds available to all farmers
+
+Our approach is divided into two parts.
+We start by using the first three survey rounds of the Ethiopian Socioeconomic Survey (ESS), which is a panel  covering  2011, 2013, and 2015.
+For each round, farmers self-reported on the seed variety they used (improved or traditional).
+To illustrate the effects of misclassification, we show the results of the GRC model for each ``adoption trajectory'' which is a sub-group characterized by the sequence of adoption choices made by the household.
+We find counter-intuitive results that seemingly show that households that adopt do not have a comparative advantage in adoption, or that households that adopt more often are less advantaged than those that adopt less.
+Returns to adoption are lowest for those that adopt more often.
+
+Our second set of results use the fourth round of the ESS (conducted in 2018), which collected DNA fingerprints of harvested cropcuts in selected plots and compared its genetic material with a reference library of available varieties in Ethiopia, hence, providing a unique opportunity to understand why improved seed varieties seem not to increase returns when adopted.
+These data provides critical information not only about the seed types but also about their purity, sources, and their year of release.
+We exploit this uniquely rich information to compare returns of verifiable improved seed varieties against self-reported use of new, second generation or recycled varieties.
+For this, we leverage different purity thresholds (or degrees of no-contamination with other varieties), year of release, types of varieties, and their sources.
+Since many categorizations for DNA-fingerprinting exist, we implement an Endogenous Switching Regression (ESR) approach, to understand each categories' impact on adoption.\footnote{ESR is a standard estimation strategy to study the question of adoption with cross-sectional data (\citealt{Marenya2020-kb}, \citealt{Shiferaw2014-op}), given its capacity to account for unobserved heterogeneity in adoption.} The results of the ESR show negative effects of adopting improved seeds.
+But the comparison of different improved seed definitions using the DNA fingerprinting shows that when the definition of improved seed is based on greater purity of the germplasm (95\% of purity or more), returns to adoption becomes positive and significant.
+We obtain similar results when evaluating the specific effects of adopting drought-tolerant maize varieties, CGIAR or Exotic varieties, or more recently released varieties.
+
+Finally, to integrate our results together, we use the misclassification rates from the fourth round of the data and re-estimate the GRC results to be robust to misclassification \cite{michuda2021three}.
+We find that once we account for misclassification, heterogeneity in adoption becomes similar to other contexts: higher comparative advantage is associated with lower returns, pointing to significant barriers to adoption for those that would benefit the most from it.
 
@@
 
-We contribute to this literature by exploring heterogeneous comparative advantages to adoption in Ethiopia between 2011 to 2015.
+% We contribute to this literature by exploring heterogeneous comparative advantages to adoption in Ethiopia between 2011 to 2015.
 %, controlling for important drivers of adoption such as rainfall (found to be the main driver of heterogeneous effects for adopting agronomic packages \citealt{Marenya2020-kb}, \citealt{Katengeza2019-af}), labor and input use.
-Our findings provide an alternative hypothesis to the puzzle of low maize adoption, namely that the wide dispersion of older generations and recycled improved seeds may be resulting in lower yields and altering farmers' input use.
+% Our findings provide an alternative hypothesis to the puzzle of low maize adoption, namely that the wide dispersion of older generations and recycled improved seeds may be resulting in lower yields and altering farmers' input use.
 The null and negative effects that we see in the GRC results are actually stemming from misclassification in self-reported adoption status.
@@
 
-There were significant differences between the analysis carried out in ESS 1-3 and the most recent, ESS4.
-Most notably, the ESS4 is no longer a panel dataset made up of the households of ESS 1-3, but it includes important information on the DNA fingerprinting of seed type, which can be used to calculate misclassification rates.
+There were significant differences between the analysis carried out with the panel survey rounds (ESS 1-3) and the most recent survey round (ESS4).
+Most notably, the ESS4 is no longer a panel dataset made up of the households of ESS 1-3, but it includes important unique information on the DNA fingerprinting of seed type, which can be used to calculate misclassification rates.
 This is important information as the first three waves of ESS only included self-reported hybrid maize status.
@@
 Harvest labor is disproportionately skewed to family participation with around 20 days of harvest labor being conducted by family members and only about 2 days for hired labor.
-This is indicative of either \textbf{malfunctioning} labor markets, income constraints or high supervision costs.
+This is indicative of either malfunctioning labor markets, income constraints or high supervision costs.
 The distances to some key areas are also highlighted in Table \ref{tbl:summary}.
@@
 
-We use self-reported and cropcut yields as outcomes in a Group Random Coefficients (GRC) approach to investigate the heterogeneous effects of adoption.
+We use a Group Random Coefficients (GRC) approach to investigate the heterogeneous effects of adoption.
 The advantage of the GRC approach is that we do not rely on identification from a valid instrument, but on modeling the heterogeneity directly through a projection of adoption returns on their observed ``trajectory'' of adoption.
@@
 The first is important because it is able to disentangle the effect of a general household's agricultural ability (which can be thought of as a fixed effect) and their adoption-specific potential.
-The second parameter can tell us whether the economy sort farmers in a way that is commensurate with providing those that would benefit the most from adoption actually adopting, or whether there are barriers interfering with that.
+The second parameter can tell us whether the economy sorts farmers in a way that is commensurate with providing those that would benefit the most from adoption actually adopting, or whether there are barriers interfering with that.
 See \cite{Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc} for more information on the assumptions involved for the GRC approach.
@@
 
+\afterpage{{\footnotesize\tabcolsep=1pt  % hold it local
+    \centering
     \input{results/tables/traj}
+    \vspace*{-4em}
+    \begin{table}[H]
+        \centering
+        \begin{tabular}{p{0.7\textwidth}} 
+            \begin{tablenotes}
+                  \small
+                  \item Note: Author's calculations using ESS.
+Maize farmers only.
+Table shows household adoption trajectories, their frequency and their share in the sample.
+1 denotes adoption of improved maize seed, 0 otherwise.
+The first digit denotes adopting in wave 1, the second in wave 2, and the third in wave 3.
+            \end{tablenotes}
+        \end{tabular}
+    \end{table}  
+    \newpage
+}}
 
@@
 It is not possible to separately identify $\mu_{(1,1,1)}$ and $\Delta_{(1,1,1)}$ without adding more structure to the model in the form of the linearity in comparative advantage (LCA) restriction, which we explain how to test for below.
+
+Our results use two definitions of yields: measured cropcuts of dry maize and self-reported maize yields.
+Each measure of yields has its benefits and drawbacks.
+Cropcuts, for instance, provide a true estimate of yields for a given parcel, but at the cost of a reduced sample size.
+Self-reports have larger sample sizes, but may suffer from measurement error \citep{gollin2021heterogeneity}.
+For controls we use the years of education of the household head, the age of the household head, the sex of the household head, whether there is a title for the land being cultivated, the parcel size in hectares, as well as the hours of household and hired labor.
+Commensurate with the literature, we also include interactions of these controls with improved seed adoption as added controls.
+Interactions with the improved seed variable are meant to further account for household specific variation that adopters may have.
 
@@
 
+To compare the results when using a dichotomous adoption decision with an intensity of adoption variable, we run a similar regression to the above to calculate the comparative advantage of adopting a certain number of times.
+We estimate an equation similar to Eq.
+\ref{eq:GRC_Suri}, but use the number of times that a household adopted an improved maize variety (here, $NAdopt$), not its trajectory.
+
+\begin{align}
+y_{it}&=\sum_{\underline{na}\in\mathcal{NA}\backslash 3}\nu_{\underline{h}}+\Delta_{(0,0,1)}h_{it}+\phi(\nu_{(0,0,1)}-\nu_{(0,0,1)})_{it}1\{NAdopt_{i}=1\}\nonumber\\
+&~~~+\left(\nu_{(1,1,1)}+\phi\left(\nu_{(1,1,1)}-\nu_{(0,0,1)}\right)\right)h_{it}1\{NAdopt_{i}=3\} + X_{it}\beta +\varepsilon_{it},\label{eq:GRC_NAdopt}
+\end{align}
+
+Here instead of a set of trajectories, we have a set of the number of times adopted $\mathcal{NA}$. $\nu$ denotes the average yield from adoption $na$ times. We use adopting once as the base.
+
 \subsection{Endogenous Switching Regression}
 
-The GRC approach uses the variation in the adoption trajectories of each observation in the ESS 1-3 panel as a source of information that allows for an unbiased estimation.
-ESS4 is no longer part of this panel data-set, and so in order to estimate the effect of adoption of improved seeds on yields, an alternative estimation method is needed.
-We propose an endogenous switching regression model which accounts for unobserved heterogeneity in adoption decisions.
-This method imposes restrictive assumptions regarding the distribution of the errors for the selection and output equations.
+The GRC approach uses the variation in the adoption trajectories of each household in the ESS 1-3 panel as a source of information that allows for an unbiased estimation.
+However, because it only includes self-reported use of improved seed, estimates may suffer from missclassification bias.
+We hence use the  DNA-fingerprinting available in ESS4 to explore possible missclassification dynamics that may better inform the results from ESS 1-3.
+We propose an endogenous switching regression model to account for unobserved heterogeneity in adoption decisions.
+This method imposes restrictive assumptions regarding the distribution of the errors for the selection (into adoption) and output equations.
 At the same time, the exclusion restriction needs to be satisfied in order to reduce the bias of the estimation.
@@
 
-Our results use two definitions of yields: measured cropcuts of dry maize and self-reported yields maize yields.
-Each measure of yields has its benefits and drawbacks.
-Cropcuts, for instance, provide a true estimate of yields for a given parcel, but at the cost of a reduced sample size.
-Self-reports have larger sample sizes, but may suffer from measurement error \citep{gollin2021heterogeneity}.
-For controls we use the years of education of the household head, the age of the household head, the sex of the household head, whether there is a title for the land being cultivated, the parcel size in hectares, as well as the hours of household and hired labor.
-Commensurate with the literature, we also include interactions of these controls with improved seed adoption as added controls.
-Interactions with the improved seed variable are meant to further account for household specific variation that adopters may have.
-
 %  self-reported yields are systematically higher than cropcut yields, pointing to the fact that there may be some measurement error at play.
-
-The results of the GRC regression are shown in Table \ref{tbl:unres}.
-The results for self-reported and cropcut yields are largely similar.
-There is some discrepancy in the self-reported yields in the first wave, and specifically with the $(1,0,0)$ trajectory.
-The return to adoption is negative for cropcuts, but positive for self-reported yields.
-The returns to adoption in many cases, are not statistically significant, or they are negative.
-Only in the case of the trajectory $(0,0,1)$: those that adopted in the second and third waves, is there a large, statistically significant effect.
-
-In order to calculate comparative advantage, there must be evidence for the LCA being satisfied.
-The lower panel of  Table \ref{tbl:unres} shows the results for the weak identification test for $\phi$, including each trajectory on its own and a joint test.
-We can see here that for the case of cropcuts with controls, there is a confidence interval for the joint test of $\phi$, giving evidence to the fact that the LCA is satisfied most prominently in that case.
-For the aggregations for this paper, then, we will use that specification.
-
-\input{results/tables/grc}
 
@@
 
-To explore this further, we run a similar regression to calculate the comparative advantage of adopting a certain number of times.
-This is similar to Eq.
-\ref{eq:GRC_Suri}, but uses the number of times that a household adopted hybrid maize (here, $NAdopt$), not its trajectory.
-
-\begin{align}
-y_{it}&=\sum_{\underline{na}\in\mathcal{NA}\backslash 3}\nu_{\underline{h}}+\Delta_{(0,0,1)}h_{it}+\phi(\nu_{(0,0,1)}-\nu_{(0,0,1)})_{it}1\{NAdopt_{i}=1\}\nonumber\\
-&~~~+\left(\nu_{(1,1,1)}+\phi\left(\nu_{(1,1,1)}-\nu_{(0,0,1)}\right)\right)h_{it}1\{NAdopt_{i}=3\} + X_{it}\beta +\varepsilon_{it},\label{eq:GRC_NAdopt}
-\end{align}
-
-In this, instead of a set of trajectories, we have a set of the number of times adopted $\mathcal{NA}$.
-$\nu$ denotes the average yield from adoption $na$ times.
-We use adopting once as the base.
-Figure \ref{fig:n_adopt} shows the results of the regression and calculation of comparative advantage.
+% We then estimate the GRC equation replacing the binary adoption variable by the number of times a farmer adopted an improved variety.
+Figure \ref{fig:n_adopt} shows the results and the calculation of the comparative advantage.
 As expected, adopting zero times leads to negative comparative advantage, but with positive returns to adoption, implying that there is value to adopting to those that have never adopted hybrid maize seed.
@@
 
-\begin{figure}
-    \centering
-    \caption{Aggregation Results: Comparative Advantage and Returns by the Number of Times Adopted (Dry Cropcuts)}
-    \label{fig:n_adopt}
-    \includegraphics[scale=0.75]{results/figures/num_adoption.png}
-    \vspace*{-2em}
-    \begin{table}[H]
-        \centering
-        \begin{tabular}{p{0.75\textwidth}} 
-            \begin{tablenotes}
-                  \small
-                  \item Note: Author's calculations using ESS.
+% \begin{figure}
+%     \centering
+%     \caption{Aggregation Results: Comparative Advantage and Returns by the Number of Times Adopted (Dry Cropcuts)}
+%     \label{fig:n_adopt}
+%     \includegraphics[scale=0.75]{results/figures/num_adoption.png}
+%     \vspace*{-2em}
+%     \begin{table}[H]
+%         \centering
+%         \begin{tabular}{p{0.75\textwidth}} 
+%             \begin{tablenotes}
+%                   \small
+%                   \item Note: Author's calculations using ESS.
 The figures shows the aggregate impact from Table \ref{tbl:unres} using the number of time a person adopted as the basis for the aggregation.
-            \end{tablenotes}
-        \end{tabular}
-    \end{table}
-\end{figure}
+%             \end{tablenotes}
+%         \end{tabular}
+%     \end{table}
+% \end{figure}
 
@@
 To complement the results of the GRC model, we present the Endogenous Switching Regression results.
-The fourth wave of the ESS data allows us to use DNA-fingerprinting results.
-This is advantageous over waves 1-3, because we do not need to rely on self-reported adoption status.
 We first examine a self-reported definition of an improved variety to validate the results from the GRC model.
@@
 Even when households think they are not adopting hybrid maize seed, there is such a large amount of hybrid seed that has been introduced over the years in Ethiopia that farmers may be misclassifying their seed and taking part in sub-optimal growing strategies as a result.
-For instance, if a farmer thinks they are growing traditional seed, but they are actually growing hybrid seed, returns to adoption will be low, since hybrid seeds often require more fertilizer and inputs to work optimally.\footnote{ If they are cultivated like traditional seeds, then their performance may end up being worse.}
+For instance, if a farmer thinks they are growing traditional seed, but they are actually growing hybrid seed, returns to adoption may be low, since hybrid seeds often require more fertilizer and inputs to work optimally.\footnote{If they are cultivated like traditional seeds, then their performance may end up being worse.}
 
-Misclassification of seed is not just a problem in Ethiopia.
-Countries have been subsidizing inputs and seed for decades and many countries might have the same challenges.
 In this section, we will introduce a misclassification-robust approach to the GRC method that will allow us to take misclassification into account when calculating comparative advantage.
-The results from the previous section highlight that there are many definitions to improved seed, which we are able to differentiate in the fourth wave of the ESS given the DNA fingerprinting.
-This gives us the ability to validate how often a hybrid seed is actually used when a farmers self-reports adopting hybrid seed.
+The DNA identification from ESS4 allows us to verify how often an improved seed is actually used when a farmer self-reports adopting improved seed.
 The methodology in this section builds on work that was done in \cite{michuda2021three}, which used confusion matrices from machine learning algorithms as a way to estimate a causal effect that was robust to misclassification.
@@
 
+\appendix
+
+\subsection{GRC Results Table}
+
+The results of the GRC regression are shown in Table \ref{tbl:unres}.
+The results for self-reported and cropcut yields are largely similar.
+There is some discrepancy in the self-reported yields in the first wave, and specifically with the $(1,0,0)$ trajectory.
+The return to adoption is negative for cropcuts, but positive for self-reported yields.
+The returns to adoption in many cases, are not statistically significant, or they are negative.
+Only in the case of the trajectory $(0,0,1)$: those that adopted in the second and third waves, is there a large, statistically significant effect.
+
+In order to calculate comparative advantage, there must be evidence for the LCA being satisfied.
+The lower panel of  Table \ref{tbl:unres} shows the results for the weak identification test for $\phi$, including each trajectory on its own and a joint test.
+We can see here that for the case of cropcuts with controls, there is a confidence interval for the joint test of $\phi$, giving evidence to the fact that the LCA is satisfied most prominently in that case.
+For the aggregations for this paper, then, we will use that specification.
+
+\begin{landscape}
+\input{results/tables/grc}
+\end{landscape}
+
 \end{document}

### 2023-03-23 paper.tex

@@
  Aleksandr Michuda \footnote{Center for Data Science for Enterprise and Society, Cornell University}%
- \and Cristina Chiarella\footnote{UCLouvain}%
+ \and Cristina Chiarella\footnote{Earth and Life Institute, UCLouvain, 1348 Louvain-la-Neuve, Belgium}%
  \and Oscar Barriga-Cabanillas\footnote{The World Bank}%
-  \and Juan Sebastian Correa \footnote{FAO}%
+  \and Juan Sebastian Correa \footnote{Food and Agriculture Organization of the United Nations (FAO), 00153, Rome, Italy}%
   }
+
 %   Assistant Research Professor at the 
@@
 			\noindent  
-			Despite large investments in  developing new  maize varieties, evidence shows that few farmers consistently adopt this new technology over time.
-We study the adoption trajectories of farmers over a three-round panel data set to explain low adoption rates with heterogeneous returns to adoption.
-By relying on self-reported use of improved seeds, we find a puzzling answer for why farmers do not use new seed varieties: adoption leads to seemingly small and negative returns.
-Heterogeneity analysis across different levels of comparative advantage show that farmers who adopt do not have a comparative advantage in adopting improved seed, nor do they exhibit high returns to adopting.
-To reconcile this result with the expected higher yields for improved seeds, we exploit unique information on maize DNA-fingerprinting collected over the same areas in 2018.
-We find that the negative returns to adoption mask heterogeneity in the actual quality of the genetic material of the seeds farmers categorized as improved varieties in the self-reported data.
+			Despite large investments in  developing new  maize varieties, evidence shows that few farmers consistently adopt this new technology over time and that average effects mask considerable hetergeneity in impacts.
+Misclassification of seed is a crucial problem in these contexts, as non-classical measurement error can bias the impacts.
+We take a step further and study misclassification in the context of heterogeneous impacts to adoption.
+We use a three-year panel to estimate heterogeneous returns to adoption from self-reports of hybrid seed use and find peculiar results driven by bias due to the misclassification of seed.
+We exploit unique information on maize DNA-fingerprinting collected over the same areas in 2018.
+We find that misclassification in self-reported hybrid seed adoption mask heterogeneity in the actual quality of the genetic material of the seeds farmers categorized as improved varieties in the self-reported data.
 Positive returns to adoption are found for those farmers that use higher-purity germplasm, drought-tolerant maize, and newly released varieties.
@@
 %%% Why it is important
-Food security is a major challenge in many East African countries, and Ethiopia is no exception, having historically struggled to provide an adequate and reliable food supply \citep{Ramakrishna2002-hv, Jaleta2018-oj}.
+% Food security is a major challenge in many East African countries, and Ethiopia is no exception, having historically struggled to provide an adequate and reliable food supply \citep{Ramakrishna2002-hv, Jaleta2018-oj}.
 Since the drought of 1984, Ethiopia has become increasingly reliant on maize, and as of today, maize constitutes the most widespread crop in the country, with total area cultivated increasing and average yields almost doubling since 1990.\footnote{FAO statistics show the total area cultivated with Maize in 2019 is 25\% larger than the next most widespread crop (sorghum), and 27\% and 139\% larger than the next two following crops, wheat and barley, respectively.}\footnote{Improved security conditions and maize productivity contributed with an important decrease in undernourishment which fell from 47\% in 1990 to 14\% in 2018 (The World Bank, 2021).
 But despite these gains, food insecurity still affects more than half the population of Ethiopia.
-The country has also recently seen increases in food insecurity because of COVID-19, the conflict in Ethiopia’s Tigray region and the desert locust outbreak.} But despite this, progress a large productivity gap persists between the African continent and the rest of the developing world.\footnote{From 2016 to the present, cereal crop yields in Sub-Saharan Africa have consistently been around 70-80\% lower than Europe or North America \citep{worldbankyield}.} The persistence of this gap is often attributed to different adoption rates of green revolution technologies, which include improved seeds.
+The country has also recently seen increases in food insecurity because of COVID-19, the conflict in Ethiopia’s Tigray region and the desert locust outbreak.} But despite this progress a large productivity gap persists between the African continent and the rest of the developing world.\footnote{From 2016 to the present, cereal crop yields in Sub-Saharan Africa have consistently been around 70-80\% lower than Europe or North America \citep{worldbankyield}.} The persistence of this gap is often attributed to different adoption rates of green revolution technologies, which include improved seeds.
 Evidence has shown that the adoption of improved seed leads to increases in agricultural yields \citep{Carter2014-fm}, food security \citep{Shiferaw2014-op}, and poverty reduction \citep{Minten2008-tj}.\footnote{Nevertheless, critics of green revolution technologies have also claimed that these technological packages have not been as profitable for many farmers in Africa as promised.
@@
 
+Low adoption of hybrid seed, despite its benefits, has been a longstanding question in the study of agricultural technologies.
 Several answers have been proposed to explain such low adoption of improved varieties.
 These include lack of liquidity and other economic constraints \citep{Carter2014-fm}, growing conditions such as the type of soil, input use, and farmer characteristics \citep{Munshi2004-og}, imperfections in credit markets \citep{Croppenstedt2003-pq}, ambiguously defined property rights \citep{Place2000-el}, lack of commitment devices \citep{Duflo2009-iv}, high transportation costs that increase the cost of agricultural inputs \citep{Byerlee2013-qk}, and improved seed supply constraints \citep{Bird2020-nt}.
-An alternative mechanism for a lack of improvement has the misclassification of seed and seed dilution.
-Net returns to improved seed is often lower than expected in the long term because a large percentage of farmers use diluted improved seed varieties, or ones that have been recycled over many generations.
+An alternative mechanism for a lack of improvement has been the misclassification of seed and seed dilution (\citealt{wossen2022misperceiving}; \citealt{kosmowski2019varietal}).
+Net returns to improved seed is often lower than expected in the long term because a large percentage of farmers use diluted improved seed varieties, or ones that have been recycled over many generations (\citealt{euler2022because}; \citealt{wossen2019estimating}).
 Evidence suggests that yield improvement and disease resistance wanes over time \citep{Abate2015-rj}.
@@
 
-Misclassification of improved seed has been studied in the case of cassava in Nigeria \citep{wossen2019estimating,macours2019comment}.
-The authors use DNA fingerprinting data to show that there is a 22\% bias in the productivity impact of improved cassava seed.
-In the context of seed adoption, however, there is evidence to suggest that the productivity impacts of seed affects households heterogeneously \citep{Suri2011-oi}, and that knowing average impacts to adoption may not be as informative.
-The observed effect of adoption may be low because households that would have a comparative advantage in adoption of improved seed are constrained by a lack of access or infrastructure.
-Additionally, the non-classical measurement error introducted by misclassification can bias heterogeneity in ways that may attenuate or even reverse productivity impacts.
+Misclassification of improved seed and its implications for measuring productivity impacts has been a relatively new branch of study, with the advent of DNA fingerprinting data.
+For instance, \cite{wossen2019estimating} use DNA fingerprinting data to show that there is a 22\% bias in the productivity impact of improved cassava seed in Nigeria.
+In the context of hybrid seed adoption, however, there is evidence to suggest that productivity impacts of seed affect households heterogeneously \citep{Suri2011-oi}, and that average impacts to adoption may not be as informative.
+The observed effect of adoption may be low because households that would have a comparative advantage in adoption of improved seed are constrained by a lack of access to improved seed or infrastructure.
+Additionally, the non-classical measurement error introduced by misclassification can bias heterogeneity in ways that may attenuate or even reverse productivity impacts.
 This paper provides a way to account for misclassification of seed when additional, error-free data exists, such as DNA fingerprinting data.
-We estimate the comparative advantage of households directly using a Group Random Coefficients (GRC) Model and correct its heterogeneity estimates using misclassification rates from DNA fingerprinting data of the same plots.
-\citep{Suri2011-oi, Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc}.
+We estimate the comparative advantage of households directly using a Group Random Coefficients (GRC) Model and correct its heterogeneity estimates using misclassification rates from DNA fingerprinting data of the same plots \citep{Suri2011-oi, Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc}.
 
@@
 
-Finally, to integrate our results together, we use the misclassification rates from the fourth round of the data and re-estimate the GRC results to be robust to misclassification \cite{michuda2021three}.
+Finally, to integrate our results together, we use the misclassification rates from the fourth round of the data and re-estimate the GRC results to be robust to misclassification (\citealt{michuda2021three}).
 We find that once we account for misclassification, heterogeneity in adoption becomes similar to other contexts: higher comparative advantage is associated with lower returns, pointing to significant barriers to adoption for those that would benefit the most from it.
@@
 Section \ref{sec:strategy} presents the methodological details of our two estimation strategies.
-Section \ref{sec:results} shows the estimations results, and Section \ref{sec:conclusion} presents the conclusion and implications of our analysis.
+Section \ref{sec:results} shows the estimations results, Section \ref{sec:disc} discusses our findings in light of prior evidence, and Section \ref{sec:conclusion} presents the conclusion and implications of our analysis.
 
@@
                   \small
-                  \item Note: Red markers denote the locations of households
+                  \item Note: Red markers denote the locations of households. Coordinates are slightly modified and repeated for confidential reasons.
             \end{tablenotes}
@@
 
-\subsection{Improved seeds}
+\subsection{Improved Seeds}
 
@@
 
-Table \ref{tbl:Trajectories} maps out the frequencies of the trajectories present in the data, sorted by frequency.
+Table \ref{tbl:trajectories} maps out the frequencies of the trajectories present in the data, sorted by frequency.
 Most households do not adopt at all (``never-adopters''), but the second most frequent trajectory are those that adopt in every wave (``always-adopters'').
@@
 
-Once we generate $\theta$ and $\Delta$ for each trajectory, we can examine heterogeneous effects to adoption.
+% Once we generate $\theta$ and $\Delta$ for each trajectory, we can examine heterogeneous effects to adoption.
 Conditional on the LCA being satisfied, that means that there will be eight comparative advantages available for comparison, which is rich in its information, but difficult to interpret.
@@
 
-To compare the results when using a dichotomous adoption decision with an intensity of adoption variable, we run a similar regression to the above to calculate the comparative advantage of adopting a certain number of times.
+% To compare the results when using a dichotomous adoption decision with an intensity of adoption variable, we run a similar regression to the above to calculate the comparative advantage of adopting a certain number of times.
 We estimate an equation similar to Eq.
@@
 
-\begin{align}
-y_{it}&=\sum_{\underline{na}\in\mathcal{NA}\backslash 3}\nu_{\underline{h}}+\Delta_{(0,0,1)}h_{it}+\phi(\nu_{(0,0,1)}-\nu_{(0,0,1)})_{it}1\{NAdopt_{i}=1\}\nonumber\\
-&~~~+\left(\nu_{(1,1,1)}+\phi\left(\nu_{(1,1,1)}-\nu_{(0,0,1)}\right)\right)h_{it}1\{NAdopt_{i}=3\} + X_{it}\beta +\varepsilon_{it},\label{eq:GRC_NAdopt}
-\end{align}
+% \begin{align}
+% y_{it}&=\sum_{\underline{na}\in\mathcal{NA}\backslash 3}\nu_{\underline{h}}+\Delta_{(0,0,1)}h_{it}+\phi(\nu_{(0,0,1)}-\nu_{(0,0,1)})_{it}1\{NAdopt_{i}=1\}\nonumber\\
+% &~~~+\left(\nu_{(1,1,1)}+\phi\left(\nu_{(1,1,1)}-\nu_{(0,0,1)}\right)\right)h_{it}1\{NAdopt_{i}=3\} + X_{it}\beta +\varepsilon_{it},\label{eq:GRC_NAdopt}
+% \end{align}
 
-Here instead of a set of trajectories, we have a set of the number of times adopted $\mathcal{NA}$. $\nu$ denotes the average yield from adoption $na$ times. We use adopting once as the base.
+% Here instead of a set of trajectories, we have a set of the number of times adopted $\mathcal{NA}$. $\nu$ denotes the average yield from adoption $na$ times. We use adopting once as the base.
 
@@
 Note that the observed expectations are those linked to Equations \ref{eq:att_1} and \ref{eq:atu_0} for adopters and non-adopters, while the counterfactual expectations are those presented in Equations \ref{eq:att_0} and \ref{eq:atu_1} respectively.
-In the next section we report these results with bootstrapped standard errors in order to measure the yield benefit of adoption for adopters and non-adopters.
+We report these results with bootstrapped standard errors in order to measure the yield benefit of adoption for adopters and non-adopters.
+
+\subsection{Misclassification-Robust GRC}
+
+As a last step, we introduce a misclassification-robust approach to the GRC method, which allows us to take misclassification into account when calculating comparative advantage.
+Using the DNA identification from ESS4 we are able to verify how often an improved seed is actually used when a farmer self-reports adopting improved seed.
+The methodology in this section builds on work that was done in \cite{michuda2021three}, which used confusion matrices from machine learning algorithms as a way to estimate a causal effect that was robust to misclassification.
+Misclassification introduces non-classical measurement error that may attenuate estimates and bias them towards 0 in the binary case.
+In the multi-regime case, misclassification will tend to make estimates too close to one another, leading to smaller or even wrong-signed estimates.
+
+Let $p_{ij}$ be the $i,j$th cell of a two by two misclassification matrix.
+Suppose that we construct the probability of being in a true trajectory $h = (i,j,k)$, intersected with being in a reported trajectory $\hat{h} = (\hat{i},\hat{j}, \hat{k})$ as:
+
+\begin{equation}
+\label{eq:misclass}
+P(h = (i,j,k) , \hat{h} = (\hat{i},\hat{j}, \hat{k})) = p_{i, \hat{i}}\cdot p_{j, \hat{j}}\cdot p_{k, \hat{k}}
+\end{equation}
+
+We can use Eq. \ref{eq:misclass} to construct the joint distribution, $f(h, \hat{h})$ for each trajectory in the sample.
+
+Once we construct these probabilities, we estimate a Nonlinear-Maximum Likelihood function that is robust to misclassification.
+Essentially, this involves using the likelihood function obtained from the confusion matrices and apply it on the estimation of Eq.
+\ref{eq:GRC} as a conditional distribution of agricultural yield.
+The likelihood for one observation in this case would then be:
+
+\begin{equation}
+\label{eq:mr-grc}
+l(\mu, \kappa, \Delta, \beta, \sigma|y_{t}, \hat{h}) = \sum_{\underbar{$h$}\in\mathcal{H}} \left[ f_{Y_{t}| \mathcal{H}}(y|h=\underbar{$h$})\cdot  \sum_{\bar{h} \in \mathcal{H}} 1\{\hat{h}=\bar{h}\} p_{\underbar{$h$}, \bar{h}} \right]
+\end{equation}
+
+\noindent where $1$ is the indicator function and $f$ is the probability distribution function conditional on true adoption assignment.
+We assume that this distribution is normal and can be written with a mean of $y_t - \sum_{\underline{h}\in\mathcal{H}\backslash (1,1,1)}\mu_{\underline{h}}-\sum_{\underline{h}\in\mathcal{H}_{S}}\Delta_{\underline{h}}h_{it}1\{h_{i}=\underline{h}\} - \kappa_{(1,1,1)}h_{it}1\{\underline{h}=(1,1,1)\}- X_{it}\beta$ and some nuisance variance parameter $\sigma$, which is estimated.
+Put more succinctly, we use Eq.
+\ref{eq:GRC} in the normal distribution.
+We will call this estimator, Misclassification-Robust GRC.\footnote{For more information on the derivation of the original estimator, as well as a link to a python package of its implementation, see \cite{michuda2021three}.}
 
@@
 
-Figure \ref{fig:theta_delta_raw} shows a plot of the comparative advantages and returns for the cropcuts model with controls.
+Figure \ref{fig:theta_delta_raw} shows a plot of the comparative advantages and returns for the cropcuts model with controls (full results for both self-report and cropcuts are shown in the Appendix).
 These results highlight two interesting puzzles: it would seem that for many trajectories, comparative advantage is either very close to 0, implying that there is no difference between farmers that adopt and not adopt.
@@
                   \small
-                  \item Note: Author's calculations using ESS. The figures shows the comparative advantage and the returns to adoption for each adoption trajectory. Dry cropcuts only
+                  \item Note: Author's calculations using ESS.
+The figures shows the comparative advantage and the returns to adoption for each adoption trajectory.
+Log(Dry cropcuts) used as outcome.
+Top figure denotes comparative advantage for each trajectory.
+Bottom figure denotes returns to adoption for each trajectory.
             \end{tablenotes}
@@
 
-To better understand the above puzzling results, we explore similar yield estimations using the different definitions of improved seeds from DNA fingerprinting.
+To better understand the above puzzling results, we explore similar yield estimations using the different definitions of improved seed from DNA fingerprinting.
 The DNA fingerprinting data allows us to be confident in whether a seed that was used, was in fact improved.
@@
 
-In this section, we will introduce a misclassification-robust approach to the GRC method that will allow us to take misclassification into account when calculating comparative advantage.
-The DNA identification from ESS4 allows us to verify how often an improved seed is actually used when a farmer self-reports adopting improved seed.
-The methodology in this section builds on work that was done in \cite{michuda2021three}, which used confusion matrices from machine learning algorithms as a way to estimate a causal effect that was robust to misclassification.
-Misclassification introduces non-classical measurement error that may attenuate estimates and bias them towards 0 in the binary case.
-In the multi-regime case, misclassification will tend to make estimates too close to one another, leading to smaller or even wrong-signed estimates.
 We construct a confusion matrix based on the DNA fingerprinting data in the fourth wave and apply it to the previous three waves.
+Table \ref{tab:misclassification} shows a table of different hybrid seed definitions and the percentage of times that a farmer self-reported traditional or hybrid seed, and whether they in fact were using traditional or hybrid seed, based on the DNA fingerprinting data.
+Table \ref{tab:a} shows that around 86\% of farmers actually use traditional seed when they self-report using traditional seed, and around 70\% of the time, farmers use hybrid seed of at least 95\% purity when they self-report using hybrid.
+There are in fact instances where farmers, however, misclassify their self-reports.
+For instance, in Table \ref{tab:b}, a farmer that self-reports using drought-tolerant maize, only in fact uses drought-tolerant maize 13\% of the time.
+There are similarly large misclassifications for seed released after 2010.
+The previous section highlighted that different definition of maize could actually lead to vastly different treatment impacts from seed adoption.
 
@@
 
-Table \ref{tab:misclassification} shows a table of different hybrid seed definitions and the percentage of times that a farmer self-reported traditional or hybrid seed, and whether they in fact were using traditional or hybrid seed, based on the DNA fingerprinting data.
-Table \ref{tab:a} shows that around 86\% of farmers actually use traditional seed when they self-report using traditional seed, and around 70\% of the time, farmers use hybrid seed of at least 95\% purity when they self-report using hybrid.
-There are in fact instances where farmers, however, misclassify their self-reports.
-For instance, in Table \ref{tab:b}, a farmer that self-reports using drought-tolerant maize, only in fact uses drought-tolerant maize 13\% of the time.
-There are similarly large misclassifications for seed released after 2010.
-The previous section highlighted that different definition of maize could actually lead to vastly different treatment impacts from seed adoption.
-
 In order to apply these misclassification probabilities to the previous three waves, we made assumptions about the nature of misclassification: that the rate of misclassification is constant across time and exhibits no dynamic behavior.
 Although these are strong assumptions, they come from the limitations of the data; the fourth wave of the ESS was the first to include DNA fingerprinting, and so was the first where a table like the one in Table \ref{tab:misclassification} could be calculated.
-Let $p_{ij}$ be the $i,j$th cell of a misclassification table as in Table \ref{tab:misclassification}.
-Suppose that we construct the probability of being in a true trajectory $h = (i,j,k)$, intersected with being in a reported trajectory $\hat{h} = (\hat{i},\hat{j}, \hat{k})$ as:
-
-\begin{equation}
-\label{eq:misclass}
-P(h = (i,j,k) , \hat{h} = (\hat{i},\hat{j}, \hat{k})) = p_{i, \hat{i}}\cdot p_{j, \hat{j}}\cdot p_{k, \hat{k}}
-\end{equation}
-
-We can use Eq.
-\ref{eq:misclass} to construct the joint distribution, $f(h, \hat{h})$ for each trajectory in the sample.
-Since Table \ref{tab:a} provides a general definition of an improved seed, we use it for the main body of the paper.
+Using these missclassification probabilities, we calculate the joint distribution for each trajectory in the sample, using a purity threshold of 95\% or higher as the chosen ``true'' definition of improved seed.
 Table \ref{tab:cm} shows the results of the construction of the joint distribution.
@@
 
-Once we construct these probabilities, we estimate a Nonlinear-Maximum Likelihood function that is robust to misclassification.
-Essentially, this involves using the likelihood function obtained from the confusion matrices and apply it on the estimation of Eq.
-\ref{eq:GRC} as a conditional distribution of agricultural yield.
-The likelihood for one observation in this case would then be:
-
-\begin{equation}
-\label{eq:mr-grc}
-l(\mu, \kappa, \Delta, \beta, \sigma|y_{t}, \hat{h}) = \sum_{\underbar{$h$}\in\mathcal{H}} \left[ f_{Y_{t}| \mathcal{H}}(y|h=\underbar{$h$})\cdot  \sum_{\bar{h} \in \mathcal{H}} 1\{\hat{h}=\bar{h}\} p_{\underbar{$h$}, \bar{h}} \right]
-\end{equation}
-
-\noindent where $1$ is the indicator function and $f$ is the probability distribution function conditional on true adoption assignment.
-We assume that this distribution is normal and can be written with a mean of $y_t - \sum_{\underline{h}\in\mathcal{H}\backslash (1,1,1)}\mu_{\underline{h}}-\sum_{\underline{h}\in\mathcal{H}_{S}}\Delta_{\underline{h}}h_{it}1\{h_{i}=\underline{h}\} - \kappa_{(1,1,1)}h_{it}1\{\underline{h}=(1,1,1)\}- X_{it}\beta$ and some nuisance variance parameter $\sigma$, which is estimated.
-Put more succinctly, we use Eq.
-\ref{eq:GRC} in the normal distribution.
-We will call this estimator, Misclassification-Robust GRC.\footnote{For more information on the derivation of the original estimator, as well as a link to a python package of its implementation, see \cite{michuda2021three}.}
-
 Once we estimate the parameters, we once again calculate comparative advantage and returns to adoption.
 Figure \ref{fig:mr-grc} shows the results of the MR-GRC and previous GRC results side-by-side.
-What's striking is that once misclassification is accounted for, comparative advantage to adoption is no longer negative to trajectories that adopt hybrid.
-In fact, the number of adoptions is positively correlated with comparative advantage.
-This suggests that households are sorting based on how well they can use hybrid.
-What is interesting, however, is that once we account for misclassification, returns to adoption become almost universally negative.
-This is indicative of a negative $phi$ term or that, there is negative selection into hybrid.
-Similar to \cite{Suri2011-oi}, a negative $\phi$ term suggests that there may be institutional factors at play that make the returns to adoption negative and this may be the reason why households do not adopt.
-The only trajectory that has positive returns to adoptions is $(0,0,1)$.
-What's peculiar about that trajectory is that comparative advantage is closest to 0 than any other trajectory.
-This may suggest that all other things being equal, adoption still presents positive returns over traditional seed, especially if that adoption occurs in the third wave.
+Misclassification of seed biases the estimates of the heterogeneous effects, both in terms of the estimation of comparative advantage and the returns to adoption.
+For instance, in the case of $(0,1,1)$, misclassification actually reverses the returns to adoption from being positive to negative.
+In other instances, misclassification attenuates the returns to adoption.
+Once misclassification is accounted for, the comparative advantage to adoption is no longer negative to trajectories that adopt hybrid seed.
+But despite considerable heterogeneity in comparative advantage, returns to adoption to most trajectories is negative, but similar in magnitude.
+This would translate to a $\phi$ term that is close to 0, but slightly negative, meaning that the there is either no sorting of households based on comparative advantage or that there is negative sorting.
+This is similar to the Kenyan context in \cite{Suri2011-oi}, who finds a negative $\phi$ term and negative returns to adoption, meaning  that there are significant barriers to using hybrid seed, and that farmers who are the most poised to gain from adoption are constrained by a lack of access to seed and other complementary inputs.
 
@@
 Misclassification defined using DNA purity data from the fourth wave of ESS.
+Log(Dry cropcuts) used as outcome.
+Top figure denotes comparative advantage for each trajectory.
+Bottom figure denotes returns to adoption for each trajectory.
             \end{tablenotes}
@@
 
-\section{Discussion}
+\section{Discussion}\label{sec:disc}
 
@@
 These findings may also hold for our case, especially taking into consideration the degree of genetic dilution and dispersion of recycled or older-generation improved varieties, which are not as yield responsive to fertilizer application.
-In addition to this, prior evidence also indicates that farmers sub-optimally allocate fertilizer in response to their mislead beliefs about the crop variety (Kosmowski \& Lybbert, 2022; \citealt{wossen2022misperceiving}).
+In addition to this, prior evidence also indicates that farmers sub-optimally allocate fertilizer in response to their mislead beliefs about the crop variety (\citealt{euler2022because}).
 Such wrong beliefs or misclassification has also been shown to be driven by misperception instead of misreporting (\citealt{wossen2022misperceiving}), which is also the case for the farmers in our sample.
@@
 Even moderate droughts have been shown to decrease maize yields of non-drought tolerant varieties by 20\% (\citealt{paul2021heterogeneous}).
-Additionally, even the drought-tolerant improved maize varieties distributed in Ethiopia are only resistant to mild drought but would also fail like any other non-drought tolerant variety in case of severe droughts.
+Additionally, although our results confirm the positive returns to DTM adoption (consistent with \citealt{wossen2017measuring}), even the DTM varieties distributed in Ethiopia are only resistant to mild drought but would also fail like any other non-drought tolerant variety in case of severe droughts.
 
@@
 But the fact that newly bred hybrids that have adapted well to local conditions outperform local varieties and older-generation improved seeds is not surprising given the large investments devoted to improve seed performance and the speed at which new high-yielding varieties are being developed.
-The results from the GRC model highlights how the fact that we imperfectly observe the use of newly improved seeds and the degree of misclassification between self-reported improved seeds and bred hybrids may partially explain why we do not see self-reported improved varieties grant positive returns to adoption.
+The results from the GRC model highlight how the fact that we imperfectly observe the use of newly improved seed and the degree of misclassification between self-reported improved seed and bred hybrids may partially explain why we do not see self-reported improved varieties grant positive returns to adoption.
 Reconciling these findings with the results from the Endogenous Switching Regression, we hypothesize that the negative returns to adoption mask a large heterogeneity in the returns to adoption.
@@
 
-\appendix
+\section{Appendix}
 

### 2023-03-23 paper.tex
Updates from Overleaf
mar 23 changes

@@
 			\noindent  
-			Despite large investments in  developing new  maize varieties, evidence shows that few farmers consistently adopt this new technology over time and that average effects mask considerable hetergeneity in impacts.
-Misclassification of seed is a crucial problem in these contexts, as non-classical measurement error can bias the impacts.
+			Despite large investments in developing new maize varieties, evidence shows that few farmers consistently adopt them.
+Moreover, average positive effects on yields mask considerable heterogeneity, particularly in contexts where seed are not traceable (missclasified as new varieties).
 We take a step further and study misclassification in the context of heterogeneous impacts to adoption.
-We use a three-year panel to estimate heterogeneous returns to adoption from self-reports of hybrid seed use and find peculiar results driven by bias due to the misclassification of seed.
-We exploit unique information on maize DNA-fingerprinting collected over the same areas in 2018.
-We find that misclassification in self-reported hybrid seed adoption mask heterogeneity in the actual quality of the genetic material of the seeds farmers categorized as improved varieties in the self-reported data.
-Positive returns to adoption are found for those farmers that use higher-purity germplasm, drought-tolerant maize, and newly released varieties.
-When we account for misclassification in our heterogeneity analysis, we find that the puzzle is resolved.
-Our findings point to the wide dispersion of older and genetically diluted varieties, for which poorer farmers may be paying a premium.
-The implications of our findings speak to the need for policies to better target context and geography, expand accessibility of improved seeds, and make higher yields varieties more inclusive.
+We use a three-year panel to estimate heterogeneous returns to adoption from self-reports of hybrid seed use and find null returns from adoption.
+We show how this result is driven misclassification of seeds that induce bias by exploiting a unique information on maize DNA-fingerprinting.
+When we account for seed misclassification, we find that heterogeneity in the actual quality of the genetic material of the seeds farmers categorized as improved varieties in the self-reported data masks positive returns to adoption, with the highest returns found for farmers using higher-purity germplasm, drought-tolerant maize, and newly released varieties.
+Our findings point to the wide dispersion of older and genetically diluted varieties, for which poorer farmers may be paying a premium, and therefor the need for policies that expand the accessibility of improved, context and geography-specific seeds, which display higher yields.
 
@@
 Low adoption of hybrid seed, despite its benefits, has been a longstanding question in the study of agricultural technologies.
-Several answers have been proposed to explain such low adoption of improved varieties.
+Given their capacity to provide higher and less variable yields, lack of consistent adoption has received a lot of attention from the economic literature, with several explanations developed in the literature.
 These include lack of liquidity and other economic constraints \citep{Carter2014-fm}, growing conditions such as the type of soil, input use, and farmer characteristics \citep{Munshi2004-og}, imperfections in credit markets \citep{Croppenstedt2003-pq}, ambiguously defined property rights \citep{Place2000-el}, lack of commitment devices \citep{Duflo2009-iv}, high transportation costs that increase the cost of agricultural inputs \citep{Byerlee2013-qk}, and improved seed supply constraints \citep{Bird2020-nt}.
-An alternative mechanism for a lack of improvement has been the misclassification of seed and seed dilution (\citealt{wossen2022misperceiving}; \citealt{kosmowski2019varietal}).
-Net returns to improved seed is often lower than expected in the long term because a large percentage of farmers use diluted improved seed varieties, or ones that have been recycled over many generations (\citealt{euler2022because}; \citealt{wossen2019estimating}).
-Evidence suggests that yield improvement and disease resistance wanes over time \citep{Abate2015-rj}.
-Conversely, given the ubiquity of improved seed varieties over the course of decades of access in Ethiopia, some farmers may use improved seed without knowing it, leading to inferior yields than conventional seed due to a lack of the application of complementary inputs.
+An alternative mechanism for a lack of improvement has been the misclassification of seed and seed genetic dilution (\citealt{wossen2022misperceiving}; \citealt{kosmowski2019varietal}).
+Yield improvement and disease resistance wanes over time as the genetic material is recycled \citep{Abate2015-rj}, making that the net returns from these \textit{improved} seed is lower than expected, and decreases in the long term as farmers continue to use genetically dilute seed varieties, or ones that have been recycled over many generations (\citealt{euler2022because}; \citealt{wossen2019estimating}).
+Moreover, as genes from improved varieties enter the regular seed market its inadvertent use can potentially lead to inferior yields since they then to require the application of complementary inputs.
 
-Misclassification of improved seed and its implications for measuring productivity impacts has been a relatively new branch of study, with the advent of DNA fingerprinting data.
-For instance, \cite{wossen2019estimating} use DNA fingerprinting data to show that there is a 22\% bias in the productivity impact of improved cassava seed in Nigeria.
-In the context of hybrid seed adoption, however, there is evidence to suggest that productivity impacts of seed affect households heterogeneously \citep{Suri2011-oi}, and that average impacts to adoption may not be as informative.
-The observed effect of adoption may be low because households that would have a comparative advantage in adoption of improved seed are constrained by a lack of access to improved seed or infrastructure.
-Additionally, the non-classical measurement error introduced by misclassification can bias heterogeneity in ways that may attenuate or even reverse productivity impacts.
+% Given the ubiquity of improved seed varieties over the course of decades of access in Ethiopia,
+
+The greater availability of DNA fingerprinting data makes possible a new revision of the role of misclassification of improved seed on the adoption of new varieties.
+For instance, \cite{wossen2019estimating} use DNA fingerprinting data to show that there is a 22\% bias in the productivity impact of improved cassava seed in Nigeria, suggesting that average impacts to adoption may not be as informative.
+Additionally, missclassification introduces a non-classical measurement error that can make that the observed productivity impact is attenuated or even reversed.
 This paper provides a way to account for misclassification of seed when additional, error-free data exists, such as DNA fingerprinting data.
@@
 
-Our results contribute to the literature on the drivers of low adoption rates of improved seed varieties.
-The consensus in this literature is that the success of improved varieties depends on a range of factors, and that the widespread adoption of  improved seeds is the result of complementary conditions and inputs that allow farmers to extract the full potential of these varieties.
-Specifically, extension services and on-farm field trials, seed variety characteristics and rainfall have been found to be crucial to the adoption of improved maize in Tanzania \citep{Kaliba2000-jh}.
-In the specific case of Ethiopia, labor, fertilizer use, farmers’ experience with extension packages, rainfall suitability and prices have been shown to increase the opportunity cost of not adopting improved wheat varieties \citep{Wale2006-bv}.
-Market access and the accessibility of extension services are also critical in the adoption of improved chickpea varieties in Ethiopia \citep{Verkaart2019-ol}.
-Other important determinants of heterogeneous effects include risk aversion \citep{Holden2016-vy}, farm size \citep{Ghimire2015-bd}, and credit constraints \citep{Simtowe2008-jn,Balana2020-hx}.
-We are the first to consider how heterogeneous impacts to seed adoption are affected by misclassification of seed.
+Our results contribute to the literature on the drivers of low adoption rates of improved seed varieties as we are the first to consider how heterogeneous impacts to seed adoption are affected by misclassification of seed.
+The consensus in this literature is that the success of improved varieties depends on a range of factors, and that the widespread adoption of improved seeds is the result of complementary conditions and inputs that allow farmers to extract the full potential of these varieties, while other factos such as risk aversion \citep{Holden2016-vy}, farm size \citep{Ghimire2015-bd}, and credit constraints \citep{Simtowe2008-jn,Balana2020-hx} also play a role.\footnote{Complementary inputs include extension services and on-farm field trials, while complementary conditions refer to seed characteristics adapted for the terrain conditions and appropriate rainfall and temperature levels.
+Input availability and appropriate conditions have been found to be crucial to the adoption of improved maize in Tanzania \citep{Kaliba2000-jh}}.
+In the specific case of Ethiopia, labor, fertilizer use, farmers’ experience with extension packages, rainfall suitability and prices have been shown to increase the opportunity cost of not adopting improved wheat varieties \citep{Wale2006-bv}.\footnote{Market access and the accessibility of extension services are also critical in the adoption of improved chickpea varieties in Ethiopia \citep{Verkaart2019-ol}}
 
@@
 \afterpage{{\footnotesize\tabcolsep=1pt  % hold it local
+    \centering
         \input{results/tables/summary}
@@
                 \centering
-        \begin{tabular}{p{0.7\textwidth}} 
+                \begin{tabular}{p{1\textwidth}} 
                     \begin{tablenotes}
                           \small
-                  \item Note: Author's calculations using ESS. Maize farmers only.
+                          \item Note: Author's calculations using ESS.
+Mean and standard deviation for each variable shown.
+Standard deviation in parentheses.
+Maize farmers only.
+Parcel size, yield and distance variables winsorized at the 1\% level.
                     \end{tablenotes}
@@
     \centering
-        \begin{tabular}{p{0.7\textwidth}} 
+        \begin{tabular}{p{1\textwidth}} 
             \begin{tablenotes}
@@
         \centering
-        \begin{tabular}{p{0.7\textwidth}} 
+        \begin{tabular}{p{1\textwidth}} 
             \begin{tablenotes}
@@
 Using the DNA identification from ESS4 we are able to verify how often an improved seed is actually used when a farmer self-reports adopting improved seed.
-The methodology in this section builds on work that was done in \cite{michuda2021three}, which used confusion matrices from machine learning algorithms as a way to estimate a causal effect that was robust to misclassification.
+The methodology in this section builds on work that was done in \cite{michuda2021three}, which used confusion matrices from machine learning algorithms as a way to estimate a causal effect that was robust to misclassification (see Table \ref{tab:misclassification}).
 Misclassification introduces non-classical measurement error that may attenuate estimates and bias them towards 0 in the binary case.
@@
 \input{results/tables/misclass_95purity.tex}
+\vspace*{-2em}
+    \begin{table}[H]
+        \centering
+        \begin{tabular}{p{1\textwidth}} 
+            \begin{tablenotes}
+                  \small
+                  \item Note: Table shows the joint distribution of misclassifying adoption of hybrid maize.
+Adoption is self-reported from the first three data rounds of the ESS data, while misclassification is estimated from the fourth round.
+Joint distribution assumes misclassification is constant and exhibits no dynamic behavior.
+            \end{tablenotes}
+        \end{tabular}
+\end{table}
 

### 2023-03-23 results/tables/misclass_95purity.tex
mar 23 changes

@@
-\begin{table}
+\begin{table}[H]
 \caption{Misclassification Matrix for Seed Adoption based on DNA Fingerprinting}
@@
 \end{tabular}
-\begin{tablenotes}
-    \item \footnotesize Notes: Table shows results of misclassification probabilities for each trajectory in the sample, using probabilities from Table \ref{tab:a}.
-Misclassification probabilities calculated from DNA fingerprinting data from Wave 4 of the ESS and assumes that true adoption of hybrid seed is when DNA purity of hybrid is at least 95\%.
-\end{tablenotes}
 \end{table}

### 2023-03-24 paper.tex
newest version

@@
 \author{%
- Aleksandr Michuda \footnote{Center for Data Science for Enterprise and Society, Cornell University}%
+ \textbf{Corresponding Author}: Aleksandr Michuda \footnote{Center for Data Science for Enterprise and Society, Cornell University, 237 Rhodes Hall, Ithaca, NY, USA, am2497@cornell.edu}%
  \and Cristina Chiarella\footnote{Earth and Life Institute, UCLouvain, 1348 Louvain-la-Neuve, Belgium}%
- \and Oscar Barriga-Cabanillas\footnote{The World Bank}%
+ \and Oscar Barriga-Cabanillas\footnote{The World Bank, 1818 H Street, NW Washington, DC USA}%
   \and Juan Sebastian Correa \footnote{Food and Agriculture Organization of the United Nations (FAO), 00153, Rome, Italy}%
@@
 We use a three-year panel to estimate heterogeneous returns to adoption from self-reports of hybrid seed use and find null returns from adoption.
-We show how this result is driven misclassification of seeds that induce bias by exploiting a unique information on maize DNA-fingerprinting.
-When we account for seed misclassification, we find that heterogeneity in the actual quality of the genetic material of the seeds farmers categorized as improved varieties in the self-reported data masks positive returns to adoption, with the highest returns found for farmers using higher-purity germplasm, drought-tolerant maize, and newly released varieties.
+We leverage unique information on maize DNA-fingerprinting to show this result is the result of bias in the estimation originating from  misclassification of seeds.
+Accounting for seed misclassification, we find that heterogeneity in the actual quality of the genetic material masks positive returns to adoption, with the highest returns found for farmers using higher-purity germplasm, drought-tolerant maize, and newly released varieties.
 Our findings point to the wide dispersion of older and genetically diluted varieties, for which poorer farmers may be paying a premium, and therefor the need for policies that expand the accessibility of improved, context and geography-specific seeds, which display higher yields.
@@
 
-Low adoption of hybrid seed, despite its benefits, has been a longstanding question in the study of agricultural technologies.
-Given their capacity to provide higher and less variable yields, lack of consistent adoption has received a lot of attention from the economic literature, with several explanations developed in the literature.
-These include lack of liquidity and other economic constraints \citep{Carter2014-fm}, growing conditions such as the type of soil, input use, and farmer characteristics \citep{Munshi2004-og}, imperfections in credit markets \citep{Croppenstedt2003-pq}, ambiguously defined property rights \citep{Place2000-el}, lack of commitment devices \citep{Duflo2009-iv}, high transportation costs that increase the cost of agricultural inputs \citep{Byerlee2013-qk}, and improved seed supply constraints \citep{Bird2020-nt}.
-An alternative mechanism for a lack of improvement has been the misclassification of seed and seed genetic dilution (\citealt{wossen2022misperceiving}; \citealt{kosmowski2019varietal}).
+These include economic constraints, such as lack of liquidity and imperfections in credit markets \citep{Croppenstedt2003-pq; Carter2014-fm }, ambiguously defined property rights which limit investments \citep{Place2000-el} as well as supply constraints due to high transportation costs of agricultural inputs \citep{Byerlee2013-qk}, and seeds \citep{Bird2020-nt}.
+Other explanations include mismatches between seed characteristics and agricultural conditions \citep{Munshi2004-og} and lack of commitment devices \citep{Duflo2009-iv}.
+Alternative mechanism, closely related with our methodological approach is the misclassification of seed and seed genetic dilution (\citealt{wossen2022misperceiving}; \citealt{kosmowski2019varietal}).
 Yield improvement and disease resistance wanes over time as the genetic material is recycled \citep{Abate2015-rj}, making that the net returns from these \textit{improved} seed is lower than expected, and decreases in the long term as farmers continue to use genetically dilute seed varieties, or ones that have been recycled over many generations (\citealt{euler2022because}; \citealt{wossen2019estimating}).
-Moreover, as genes from improved varieties enter the regular seed market its inadvertent use can potentially lead to inferior yields since they then to require the application of complementary inputs.
+Moreover, as genes from improved varieties enter the regular seed market its inadvertent use can potentially lead to inferior yields since they require complementary inputs.
 
@@
 
-The greater availability of DNA fingerprinting data makes possible a new revision of the role of misclassification of improved seed on the adoption of new varieties.
+We provide a way to account for misclassification of seed when additional, error-free data exists, such as DNA fingerprinting data.
+Our results contribute to the literature on the drivers of low adoption rates of improved seed varieties as we are the first to consider how heterogeneous impacts to seed adoption are affected by misclassification of seed.
+Considering that  missclassification introduces a non-classical measurement error that attenuates, or even reversed,  observed productivity impacts, the greater availability of DNA fingerprinting data makes possible a new revision of the role of misclassification of improved seed on the adoption of new varieties.
 For instance, \cite{wossen2019estimating} use DNA fingerprinting data to show that there is a 22\% bias in the productivity impact of improved cassava seed in Nigeria, suggesting that average impacts to adoption may not be as informative.
-Additionally, missclassification introduces a non-classical measurement error that can make that the observed productivity impact is attenuated or even reversed.
-This paper provides a way to account for misclassification of seed when additional, error-free data exists, such as DNA fingerprinting data.
 We estimate the comparative advantage of households directly using a Group Random Coefficients (GRC) Model and correct its heterogeneity estimates using misclassification rates from DNA fingerprinting data of the same plots \citep{Suri2011-oi, Tjernstrom_Emilia_Dalia_Ghanem_Oscar_Barriga_Cabanillas_Travis_J_Lybbert_Jeffrey_D_Michler_and_Aleksandr_Michuda2020-bc}.
@@
 
-Our results contribute to the literature on the drivers of low adoption rates of improved seed varieties as we are the first to consider how heterogeneous impacts to seed adoption are affected by misclassification of seed.
 The consensus in this literature is that the success of improved varieties depends on a range of factors, and that the widespread adoption of improved seeds is the result of complementary conditions and inputs that allow farmers to extract the full potential of these varieties, while other factos such as risk aversion \citep{Holden2016-vy}, farm size \citep{Ghimire2015-bd}, and credit constraints \citep{Simtowe2008-jn,Balana2020-hx} also play a role.\footnote{Complementary inputs include extension services and on-farm field trials, while complementary conditions refer to seed characteristics adapted for the terrain conditions and appropriate rainfall and temperature levels.
@@
 These findings may also hold for our case, especially taking into consideration the degree of genetic dilution and dispersion of recycled or older-generation improved varieties, which are not as yield responsive to fertilizer application.
-In addition to this, prior evidence also indicates that farmers sub-optimally allocate fertilizer in response to their mislead beliefs about the crop variety (\citealt{euler2022because}).
+In addition to this, prior evidence also indicates that farmers sub-optimally allocate fertilizer in response to their mislead beliefs about the crop variety leading to lower yield realization and consequent disadotpion  (\citealt{euler2022because}).
 Such wrong beliefs or misclassification has also been shown to be driven by misperception instead of misreporting (\citealt{wossen2022misperceiving}), which is also the case for the farmers in our sample.
-These findings suggest that the seemingly small to negative returns to adoption that we find through the GRC model are driven by both a lower response to fertilizers of the older-generation improved varieties, but also to farmers’ sub-optimal application of fertilizers (and possibly other inputs) due to misperception.
+These findings suggest that the seemingly small to negative returns to adoption that we find through the GRC model are driven by a combination of lower responses to fertilizers of the older-generation improved varieties and farmers’ sub-optimal application of fertilizers (and possibly other inputs) due to misperception.
 
-The suitability of the improved varieties to weather conditions and the weather conditions themselves are two additional factors that may further explain the seemingly negative comparative advantages to adopting the misperceived improved varieties.
+The suitability of the improved varieties to location-specific weather conditions as well as the variation in the weather conditions themselves are two additional factors that may further explain the seemingly negative comparative advantages to adopting the misperceived improved varieties.
 Even moderate droughts have been shown to decrease maize yields of non-drought tolerant varieties by 20\% (\citealt{paul2021heterogeneous}).

### 2023-03-24 paper.tex
make changes

--- 2023-03-23 paper.tex
@@
 			\noindent  
-			Despite large investments in  developing new  maize varieties, evidence shows that few farmers consistently adopt this new technology over time and that average effects mask considerable hetergeneity in impacts.
-Misclassification of seed is a crucial problem in these contexts, as non-classical measurement error can bias the impacts.
-We take a step further and study misclassification in the context of heterogeneous impacts to adoption.
-We use a three-year panel to estimate heterogeneous returns to adoption from self-reports of hybrid seed use and find peculiar results driven by bias due to the misclassification of seed.
-We exploit unique information on maize DNA-fingerprinting collected over the same areas in 2018.
-We find that misclassification in self-reported hybrid seed adoption mask heterogeneity in the actual quality of the genetic material of the seeds farmers categorized as improved varieties in the self-reported data.
-Positive returns to adoption are found for those farmers that use higher-purity germplasm, drought-tolerant maize, and newly released varieties.
-When we account for misclassification in our heterogeneity analysis, we find that the puzzle is resolved.
-Our findings point to the wide dispersion of older and genetically diluted varieties, for which poorer farmers may be paying a premium.
-The implications of our findings speak to the need for policies to better target context and geography, expand accessibility of improved seeds, and make higher yields varieties more inclusive.
+			Despite large investments in developing new maize varieties, evidence shows that few farmers consistently adopt them.
+Average effects on yields mask considerable heterogeneity.
+Misclassification of seed is a crucial problem in these contexts, as non-classical measurement error can lead to biased effects.
+We study misclassification in the context of heterogeneous impacts to adoption.
+We use a three-year panel to estimate heterogeneous returns to adoption from self-reports of hybrid seed use and find results driven by bias due to the seed misclassification.
+We leverage unique information on maize DNA-fingerprinting collected over the same areas in 2018.
+Misclassification in self-reported hybrid seed adoption masks heterogeneity in the quality of the genetic material categorized as improved varieties.
+Accounting for seed misclassification, we find that heterogeneity in the actual quality of the genetic material masks positive returns to adoption, with the highest returns found for farmers using higher-purity germplasm, drought-tolerant maize, and newly released varieties.
+When we account for misclassification in our heterogeneity analysis, we find that farmers that have the largest potential to benefit from adoption are constrained in their ability to do so.
+The wide dispersion of older and genetically diluted varieties, for which poorer farmers may be paying a premium points to the need for policies that expand the accessibility of improved seed and verify its quality.
 
@@
 
+% Given the ubiquity of improved seed varieties over the course of decades of access in Ethiopia,
+
 Misclassification of improved seed and its implications for measuring productivity impacts has been a relatively new branch of study, with the advent of DNA fingerprinting data.
-For instance, \cite{wossen2019estimating} use DNA fingerprinting data to show that there is a 22\% bias in the productivity impact of improved cassava seed in Nigeria.
-In the context of hybrid seed adoption, however, there is evidence to suggest that productivity impacts of seed affect households heterogeneously \citep{Suri2011-oi}, and that average impacts to adoption may not be as informative.
-The observed effect of adoption may be low because households that would have a comparative advantage in adoption of improved seed are constrained by a lack of access to improved seed or infrastructure.
-Additionally, the non-classical measurement error introduced by misclassification can bias heterogeneity in ways that may attenuate or even reverse productivity impacts.
+For instance, \cite{wossen2019estimating} use such data to show that there is a 22\% bias in the productivity impact of improved cassava seed in Nigeria.
+Other literature suggests that hybrid seed adoption is heterogeneous across households \citep{Suri2011-oi}, and that average impacts to adoption may not be as informative.
+The observed effect of adoption may be low because households that would have a comparative advantage in adoption of improved seed are constrained by a lack of access to improved seed or necessary infrastructure.
+Additionally, non-classical measurement error introduced by misclassification can bias heterogeneity in ways that may attenuate or even reverse productivity impacts.
 This paper provides a way to account for misclassification of seed when additional, error-free data exists, such as DNA fingerprinting data.
@@
 
-Our results contribute to the literature on the drivers of low adoption rates of improved seed varieties.
 The consensus in this literature is that the success of improved varieties depends on a range of factors, and that the widespread adoption of  improved seeds is the result of complementary conditions and inputs that allow farmers to extract the full potential of these varieties.
-Specifically, extension services and on-farm field trials, seed variety characteristics and rainfall have been found to be crucial to the adoption of improved maize in Tanzania \citep{Kaliba2000-jh}.
 In the specific case of Ethiopia, labor, fertilizer use, farmers’ experience with extension packages, rainfall suitability and prices have been shown to increase the opportunity cost of not adopting improved wheat varieties \citep{Wale2006-bv}.
@@
 Other important determinants of heterogeneous effects include risk aversion \citep{Holden2016-vy}, farm size \citep{Ghimire2015-bd}, and credit constraints \citep{Simtowe2008-jn,Balana2020-hx}.
-We are the first to consider how heterogeneous impacts to seed adoption are affected by misclassification of seed.
+We are the first to consider how heterogeneous impacts to seed adoption are affected by misclassification of seed.\footnote{In other contexts, extension services and on-farm field trials, seed variety characteristics and rainfall have been found to be crucial to the adoption of improved maize in Tanzania \citep{Kaliba2000-jh}.}
 
@@
 These findings may also hold for our case, especially taking into consideration the degree of genetic dilution and dispersion of recycled or older-generation improved varieties, which are not as yield responsive to fertilizer application.
-In addition to this, prior evidence also indicates that farmers sub-optimally allocate fertilizer in response to their mislead beliefs about the crop variety (\citealt{euler2022because}).
+In addition to this, prior evidence also indicates that farmers sub-optimally allocate fertilizer in response to their mislead beliefs about the crop variety leading to lower yield realization and consequent dis-adoption  (\citealt{euler2022because}).
 Such wrong beliefs or misclassification has also been shown to be driven by misperception instead of misreporting (\citealt{wossen2022misperceiving}), which is also the case for the farmers in our sample.
-These findings suggest that the seemingly small to negative returns to adoption that we find through the GRC model are driven by both a lower response to fertilizers of the older-generation improved varieties, but also to farmers’ sub-optimal application of fertilizers (and possibly other inputs) due to misperception.
+These findings suggest that the seemingly small to negative returns to adoption that we find through the GRC model are driven by a combination of lower responses to fertilizers of the older-generation improved varieties and farmers’ sub-optimal application of fertilizers (and possibly other inputs) due to misperception.
 
-The suitability of the improved varieties to weather conditions and the weather conditions themselves are two additional factors that may further explain the seemingly negative comparative advantages to adopting the misperceived improved varieties.
+The suitability of the improved varieties to location-specific weather conditions as well as the variation in the weather conditions themselves are two additional factors that may further explain the seemingly negative comparative advantages to adopting the misperceived improved varieties.
 Even moderate droughts have been shown to decrease maize yields of non-drought tolerant varieties by 20\% (\citealt{paul2021heterogeneous}).

### 2023-03-30 paper.tex
Version submitted to JDE

@@
 
-    \title{Genetic Dilution erodes productivity: Exploring farmers' low adoption levels of improved maize in  Ethiopia}
+    \title{Genetic Dilution erodes productivity: Exploring farmers' low adoption levels of improved maize in  Ethiopia\footnote{We are grateful for comments on earlier drafts of the paper from Travis Lybbert and James Stevenson, as well as comments from participants in presentations at NEUDC 2022.
+This research was supported by the CGIAR Standing Panel on Impact Assessment (SPIA) under its program of work 2019-2021.
+Thank you to Mai Mahmoud, Mina Karasalo, Karen Macours and the rest of the SPIA Small Grants team.
+All errors are our own.
+Declarations of interest: none.} }
 % This second title is my preferred title at this point (28 Oct 2021). Open to other titles, but it shouldn't be longer than this one!
@@
 			\noindent
-			Despite large investments in developing new maize varieties, evidence shows that few farmers consistently adopt them.
+   Household surveys often rely on self-reports of hybrid seed use, which are prone to measurement error.
+This context exhibits considerable heterogeneity when estimating productivity impacts.
+Using a three-year panel, we estimate heterogeneous returns to adoption from self-reports of hybrid seed use and find results driven by non-classical measurement error due to seed misclassification.
+We leverage information on DNA-fingerprinting collected over the same areas in 2018 and find that heterogeneity in the quality of the genetic material masks positive returns to adoption, with the highest returns found for farmers using higher-purity, drought-tolerant, and newly released varieties.
+When we account for misclassification in our initial heterogeneity analysis, we find that farmers that have the largest potential to benefit from adoption are constrained in their ability to do so.
+Our research points to the need for policies that expand the accessibility of improved seed and the need for mechanisms to verify its quality.
+			% Despite large investments in developing new maize varieties, evidence shows that few farmers consistently adopt them.
 Average effects on yields mask considerable heterogeneity.
@@
         {\small \noindent\emph{JEL Classification}: O12, O13, O33, Q12, C13 \\
-        	\emph{Keywords}: Improved maize, DNA-fingerprinting, Group Random Coefficient, Endogenous Switching Regression, Heterogeneous returns}
+        	\emph{Keywords}: Improved maize, DNA-fingerprinting, Group Random Coefficient, Endogenous Switching Regression, Heterogeneous returns, Ethiopia}
 
